---
title: "NATO convergence and defense spending"
author: "src"
date: "October 19, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
 setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))




write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


**load and subset the data we want**
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
DefNATO_EUspread <- subset(EuropeOmnibus, !is.na(DefSpendDelt_lead))

DefNATO_EUspread_lag0 <- subset(DefNATO_EUspread, !is.na(NATO.EUspread))
DefNATO_EUspread_lag1 <- subset(DefNATO_EUspread, !is.na(NATO.EUspread_lag1))
DefNATO_EUspread_lag2<- subset(DefNATO_EUspread, !is.na(NATO.EUspread_lag2))


pdata <- plm.data(DefNATO_EUspread_lag1, index=c("Country", "Year"))


```



``` {r PublicOpinionLag, echo = TRUE}
NATO.EUspreadResults_model<-list()
NATO.EUspreadResults_name<-c()

NATO.EUspreadResults_model[1] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread, DefNATO_EUspread_lag0) )
NATO.EUspreadResults_name[1]<-"PubOp_ne1"
screenreg(NATO.EUspreadResults_model[1],
          custom.model.names=NATO.EUspreadResults_name[1])
summary(NATO.EUspreadResults_model[[1]])

NATO.EUspreadResults_model[2] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[2]<-"PubOp_ne2"
screenreg(NATO.EUspreadResults_model[c(1:2)],
          custom.model.names=NATO.EUspreadResults_name[c(1:2)])
summary(NATO.EUspreadResults_model[[2]])

NATO.EUspreadResults_model[3] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag2, DefNATO_EUspread_lag2))
NATO.EUspreadResults_name[2]<-"PubOp_ne2"
screenreg(NATO.EUspreadResults_model[c(1:3)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3)])
summary(NATO.EUspreadResults_model[[3]])

#The two year difference in PubOp_ne2 has the largest coefficient and the smallest p-value.

```

After comparing the three options, the three year difference between expenditure and Nato Essential polling appears to have the strongest relationship, even if it no where close to significant.


``` {r SecurityEnvironment, echo = TRUE}


NATO.EUspreadResults_model[4] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + IntAt, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[4]<-"IntAt"
screenreg(NATO.EUspreadResults_model[c(1:3,4)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,4)])
summary(NATO.EUspreadResults_model[[4]])

#Lowers adjusted R^2 intl attack will not be used.

NATO.EUspreadResults_model[5] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + DomAt, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[5]<-"DomAt"
screenreg(NATO.EUspreadResults_model[c(1:3,5)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,5)])
summary(NATO.EUspreadResults_model[[5]])

# No net change in R^2.

NATO.EUspreadResults_model[6] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + CivilWar, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[6]<-"CivilWar"
screenreg(NATO.EUspreadResults_model[c(1:3,6)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,6)])
summary(NATO.EUspreadResults_model[[6]])
summary(DefNATO_EUspread_lag1$CivilWar)

#NATO.EU closer  removes Civil war cases 


NATO.EUspreadResults_model[7] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + IntlCnf, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[7]<-"IntlCnf"
screenreg(NATO.EUspreadResults_model[c(1:3,7)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,7)])
summary(NATO.EUspreadResults_model[[7]])
summary(EuropeOmnibus$IntlCnf)

#NATO.EU closer  removes IntlCnf cases 

```

Minimal gains from the security environment variables, and the signs were all in an unexpected direction.


``` {r MacroEconomics, echo = TRUE}


reg<-8
NATO.EUspreadResults_model[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[reg]<-"GDPpCapDelt"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#No improves p value of lag and signs makes sesne, keeping.



```

Added GDPpCapDelt which had a sign that made sense and improved p value of the lag, even if it doesn't help the R^2

``` {r Parliament, echo = TRUE}


reg<-9
NATO.EUspreadResults_model[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + Cab_left_right, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[reg]<-"Left_Right"
screenreg(NATO.EUspreadResults_model[c(1:3,8,reg)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,8,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Lower adjusted R^2 and doesn't make sense. Dropping Cab_left_right


reg<-10
NATO.EUspreadResults_model[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + Cab_liberty_authority, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[reg]<-"Lib_Auth"
screenreg(NATO.EUspreadResults_model[c(1:3,8,reg)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,8,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Lower adjusted R^2 and doesn't make sense. Dropping Cab_liberty_authority

reg<-11
NATO.EUspreadResults_model[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + Cab_eu_anti_pro, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[reg]<-"EU_anti_pro"
screenreg(NATO.EUspreadResults_model[c(1:3,8,reg)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,8,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Lower adjusted R^2. Dropping Cab_eu_anti_pro


reg<-12
NATO.EUspreadResults_model[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + left_right_ls_spread, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[reg]<-"Left_right_dif"
screenreg(NATO.EUspreadResults_model[c(1:3,8,reg)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,8,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Small gain, but doesn't help p-value, lets compare to the alternatives.

reg<-13
NATO.EUspreadResults_model[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[reg]<-"la_spread"
screenreg(NATO.EUspreadResults_model[c(1:3,8,12,reg)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,8,12,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Weaker value, but actually does more to clarify influence of polling variable.

reg<-14
NATO.EUspreadResults_model[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + eu_anti_pro_ls_spread, DefNATO_EUspread_lag1))
NATO.EUspreadResults_name[reg]<-"eu_spread"
screenreg(NATO.EUspreadResults_model[c(1:3,8,12,13,reg)],
          custom.model.names=NATO.EUspreadResults_name[c(1:3,8,12,13,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Lowers R^2. Not significant.


```

Model #13 is chosen, but is not significant.



**Tests for panel data model.**

``` {r defspreadPoolingTest, echo = TRUE}

NATOEUSpreadResults_model13 <- lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, DefNATO_EUspread_lag1)


DefSpreadPooled <- plm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, data = pdata, model = "pooling")
screenreg(list(NATOEUSpreadResults_model13, DefSpreadPooled),
          custom.model.names = c("OLS", "Pooled"))

DefSpreadBetween <- plm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, data = pdata, model = "between")
summary(DefSpreadBetween)
screenreg(list(NATOEUSpreadResults_model13, DefSpreadPooled, DefSpreadBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

DefSpreadFD <- plm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, data = pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(NATOEUSpreadResults_model13, DefSpreadPooled, DefSpreadBetween, DefSpreadFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


DefSpreadWithin <- plm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, data =  pdata, model = "within")
summary(DefSpreadWithin)
screenreg(list(NATOEUSpreadResults_model13, DefSpreadPooled, DefSpreadBetween, DefSpreadFD, DefSpreadWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


DefSpreadRandom <- plm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, data = pdata, model = "random")
screenreg(list(NATOEUSpreadResults_model13, DefSpreadPooled, DefSpreadBetween, DefSpreadFD, DefSpreadWithin, DefSpreadRandom),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))


##lagrange multiplier (lm) for random effects vs. OLS
plmtest(DefSpreadPooled) #if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"

##fixed effects versus OLS model
pFtest(DefSpreadWithin, DefSpreadPooled)
##p = .168 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
phtest(DefSpreadRandom, DefSpreadWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```
