---
title: "EU Exert Strong Leadership and Investment Spending"
author: "src"
date: "October 20, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---
#Setup
##Load Packages and Support Files
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
#setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


##Load the Data
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()



write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )



```


##Subset the Data
```{r EULeadershipNarrowing, echo = TRUE, results = "hide"}
#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <-subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )





##load data using function
EULeadershipSpread <- subset(EuropeOmnibus, !is.na(EquSpendDelt_lead))


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <-subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))


EULeadershipSpread_gap1 <- subset(EULeadershipSpread, !is.na(EUldrSpread))
EULeadershipSpread_gap2 <- subset(EULeadershipSpread, !is.na(EUldrSpread_lag1))
EULeadershipSpread_gap3 <- subset(EULeadershipSpread, !is.na(EUldrSpread_lag2))



```

#Building the Regression Model
##Choosing the Public Opinion Lag
```{r EULeadershipspreadRegressions, echo = TRUE}
EquiEULeadershipResults1 <- lm(EquSpendDelt_lead ~ EUldrSpread, EULeadershipSpread_gap1) 
summary(EquiEULeadershipResults1)
screenreg(list(EquiEULeadershipResults1))

EquiEULeadershipResults2 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults2)
screenreg(list(EquiEULeadershipResults1, EquiEULeadershipResults2),
              custom.model.names=c("PubOp_0","PubOp_1"))

##lag 1 is not significant

EquiEULeadershipResults3 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag2, EULeadershipSpread_gap3) 
summary(EquiEULeadershipResults3)
screenreg(list(EquiEULeadershipResults1, EquiEULeadershipResults3),
              custom.model.names=c("PubOp_0","PubOp_2"))

##lag2 is not significant

##lag1 has the lowest p value and will be included
```

##Security Environment Controls
``` {r SecurityEnvironment, echo = TRUE}

EULeadershipSpread_gap2_pdata <- plm.data(EULeadershipSpread_gap2, index=c("Country", "Year"))


EquiEULeadershipResults4 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + IntAt, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults4)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults4),
              custom.model.names=c("PubOp_1","IntAt"))



##intAt isn't significant and won't be included

EquiEULeadershipResults5 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults5)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5),
              custom.model.names=c("PubOp_1","UNmilitaryPMil"))

##TBD

EquiEULeadershipResults6 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + IntlCnf, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults6)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults6),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "IntlCnf"))

##IntlCnf is not significant and Won't be included


```


##Macroeconomic Controls

``` {r MacroVariables, echo = TRUE}

EquiEULeadershipResults7 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + GDPpCapLCUdelt, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults7)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults7),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "GDP"))

##GDPpCapLCUdelt isnt significant and will not  be included.
```

##Parliamentary Opinion Controls

``` {r Government, echo = TRUE}


EquiEULeadershipResults8 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_left_right, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults8)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults8),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "cableftright"))

##cableft right is not significant and will not be included

EquiEULeadershipResults9 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults9)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults9),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "cablibertyauthority"))

##cab_liberty_authority is significant at the .1  percent level will be included. 

EquiEULeadershipResults10 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults10)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults9, EquiEULeadershipResults10),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "cablibertyauthority", "cabeuantipro"))

##cab_eu_anti_pro is significant and will be included.

EquiEULeadershipResults11 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro + left_right_ls_spread, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults11)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults9, EquiEULeadershipResults10, EquiEULeadershipResults11),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "cablibertyauthority", "cabeuantipro", "leftrightls"))

##leftrightls is not significant and won't be included

EquiEULeadershipResults12 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro + liberty_authority_ls_spread, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults12)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults9, EquiEULeadershipResults10, EquiEULeadershipResults12),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "cablibertyauthority", "cabeuantipro", "libauthLS"))

##libauth ls isn't significant and wont' be included

EquiEULeadershipResults13 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro + eu_anti_pro_ls_spread, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults13)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults9, EquiEULeadershipResults10, EquiEULeadershipResults13),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "cablibertyauthority", "cabeuantipro", "euantiproLS"))

##euantipro isn't significant and wont' be included.

summary(EquiEULeadershipResults10)

```

**Model 10 will be used. **


#Determining Panel Data Model

``` {r defspreadPoolingTest, echo = TRUE}

EULeadershipResults10 <- lm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro, EULeadershipSpread_gap2) 
summary(EquiEULeadershipResults10)
screenreg(list(EquiEULeadershipResults2, EquiEULeadershipResults5, EquiEULeadershipResults9, EquiEULeadershipResults10),
              custom.model.names=c("PubOp_1","UNmilitaryPMil", "cablibertyauthority", "cabeuantipro"))


EquSpreadPooled <- plm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro, data = EULeadershipSpread_gap2_pdata, model = "pooling")

screenreg(list(EULeadershipResults10, EquSpreadPooled),
          custom.model.names = c("OLS", "Pooled"))

summary(EquSpreadPooled)

EquSpreadBetween <- plm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro, data = EULeadershipSpread_gap2_pdata, model = "between")
summary(EquSpreadBetween)
screenreg(list(EULeadershipResults10, EquSpreadPooled, EquSpreadBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

EquSpreadFD <- plm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro, data = EULeadershipSpread_gap2_pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(EULeadershipResults10, EquSpreadPooled, EquSpreadBetween, EquSpreadFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


EquSpreadWithin <- plm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro, data =  EULeadershipSpread_gap2_pdata, model = "within")
summary(EquSpreadWithin)
screenreg(list(EULeadershipResults10, EquSpreadPooled, EquSpreadBetween, EquSpreadFD, EquSpreadWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


# EquSpreadRandom <- plm(EquSpendDelt_lead ~ EUldrSpread_lag1 + UNmilitaryPMil + Cab_liberty_authority + Cab_eu_anti_pro, data = EULeadershipSpread_gap2_pdata, model = "random")
# screenreg(list(EULeadershipResults10, EquSpreadPooled, EquSpreadBetween, EquSpreadFD, EquSpreadWithin, DefSpreadRandom),
#           custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))
# 

##lagrange multiplier (lm) for random effects vs. OLS
plmtest(EquSpreadPooled) #if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"

##fixed effects versus OLS model
pFtest(EquSpreadWithin, EquSpreadPooled)
##p = .168 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
# phtest(EquSpreadRandom, EquSpreadWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```

- plmtest tells use to use the pooled ols model (p = .341)

- pftest tells us to use pooled because the p value = .669 cannot reject null that the pooling model is best.

**Use pooled.**
