---
title: "Country spending too much or too little on defense data and Operations and Maintenance spending"
author: "Gregory Sanders"
date: "June 1, 2016"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

#Setup
##Load Packages and Support Files
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
 # #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


##Load and Subset the Data
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <- subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


###Prepare the Datasets for Each Lag
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EuropeDefSpread <- subset(EuropeOmnibus, !is.na(OnMspendDelt_lead))



EuropeDefSpread_gap1 <- subset(EuropeDefSpread, !is.na(DefSpread))
EuropeDefSpread_gap2 <- subset(EuropeDefSpread, !is.na(DefSpread_lag1))
EuropeDefSpread_gap3 <- subset(EuropeDefSpread, !is.na(DefSpread_lag2))



```


#Building the Regression Model
##Choosing the Public Opinion Lag
```{r OnMspreadOLS, echo = TRUE}
OnMEUDefSpreadResults1 <- lm(OnMspendDelt ~ DefSpread, EuropeDefSpread_gap1) 
summary(OnMEUDefSpreadResults1)

OnMEUDefSpreadResults2 <- lm(OnMspendDelt ~ DefSpread_lag1, EuropeDefSpread_gap2) 
summary(OnMEUDefSpreadResults2)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2),
          custom.model.names=c("PubOp_df1","PubOp_df2"))

OnMEUDefSpreadResults3 <- lm(OnMspendDelt ~ DefSpread_lag2, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults3)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3"))

##use 2 lag
EuropeDefSpread_gap3_pdata <- plm.data(EuropeDefSpread_gap3, index=c("Country", "Year"))

```

##Security Environment Controls

```{r SecurityEnvironment, echo = TRUE}
OnMEUDefSpreadResults4 <- lm(OnMspendDelt ~ DefSpread_lag2 + IntAt, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults4)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults4),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "IntAt"))

#intAT not significant won't include

OnMEUDefSpreadResults5 <- lm(OnMspendDelt ~ DefSpread_lag2 + CivilWar, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults5)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults5),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "CivilWar"))

##civilwar isn't significant, won't include


OnMEUDefSpreadResults6 <- lm(OnMspendDelt ~ DefSpread_lag2 + IntlCnf, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults6)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults6),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "IntCnf"))

## IntCnf not significant won't include

```

##Macroeconomic Controls

```{r Macroeconomics}

OnMEUDefSpreadResults7 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults7)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDPpc"))

##GDPpCapLCUdelt is significant.


OnMEUDefSpreadResults7a <- lm(OnMspendDelt ~ DefSpread_lag2 + GDP2005usdDelt, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults7)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP2005"))


```


##Parliamentary Opinion Controls
```{r Cabinet}
OnMEUDefSpreadResults8 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt + Cab_left_right, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults8)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults8),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDPpCapLCUdelt", "Cableftright"))


##cableft right not significant will not include



OnMEUDefSpreadResults9 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt + Cab_liberty_authority, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults9)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults9),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3","GDPpCapLCUdelt", "lib_authority"))

##cab liberty authority not significant won't be included

OnMEUDefSpreadResults10 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt + Cab_eu_anti_pro, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults10)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults10),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDPpCapLCUdelt", "euAntiPro"))

##not significant, won't be included eu anti pro

OnMEUDefSpreadResults11 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt +left_right_ls_spread , EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults11)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS"))

##left_right_ls_spread spread is not significant, will not be included 

OnMEUDefSpreadResults12 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt  + liberty_authority_ls_spread, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults12)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11, OnMEUDefSpreadResults12),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS", "libauthLS"))
##libauthLS isn't signficicant won't be included

OnMEUDefSpreadResults13 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt  + eu_anti_pro_ls_spread, EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults13)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11, OnMEUDefSpreadResults13),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS", "euantiproLS"))
##eu anti pro isn't signficicant won't be included

```

**results 7 is final model.**

#Determining Panel Data Model


``` {r defspreadPoolingTest, echo = TRUE}

OnMEUDefSpreadResults7 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt , EuropeDefSpread_gap3) 
summary(OnMEUDefSpreadResults7)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults7),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS"))


OnMspreadPooled <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt , data = EuropeDefSpread_gap3_pdata, model = "pooling")

screenreg(list(OnMEUDefSpreadResults7, OnMspreadPooled),
          custom.model.names = c("OLS", "Pooled"))

OnMspreadBetween <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt , data = EuropeDefSpread_gap3_pdata, model = "between")
summary(OnMspreadBetween)
screenreg(list(OnMEUDefSpreadResults7, OnMspreadPooled, OnMspreadBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

OnMspreadFD <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt , data = EuropeDefSpread_gap3_pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(OnMEUDefSpreadResults7, OnMspreadPooled, OnMspreadBetween, OnMspreadFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


OnMspreadWithin <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt , data =  EuropeDefSpread_gap3_pdata, model = "within")
summary(OnMspreadWithin)
screenreg(list(OnMEUDefSpreadResults7, OnMspreadPooled, OnMspreadBetween, OnMspreadFD, OnMspreadWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


OnMspreadRandom <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapLCUdelt , data = EuropeDefSpread_gap3_pdata, model = "random")
screenreg(list(OnMEUDefSpreadResults7, OnMspreadPooled, OnMspreadBetween, OnMspreadFD, OnMspreadWithin, OnMspreadRandom),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))
summary(OnMspreadRandom)

##lagrange multiplier (lm) for random effects vs. OLS
plmtest(OnMspreadPooled) #if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"

##fixed effects versus OLS model
pFtest(OnMspreadWithin, OnMspreadPooled)
##p = .168 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
phtest(OnMspreadRandom, OnMspreadWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```

- plmtset, p = 4.01e-9, p value small so reject the null hypothesis that pooling model works and potentially random effects is best. 

- pftest p = .009, reject the null that the pooling model is best, which suggest to use the OLS model

- phtest, p = .9099, cannot reject the null hypothesis that random effects is best

**Use random effects.**
