---
title: "Country spending too much or too little on defense data and Operations and Maintenance spending"
author: "src"
date: "October 16, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
 # setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))

##get rird of 2002 and 2003
# EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


**load and subset the data we want**
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EuropeDefSpread <- subset(EuropeOmnibus, !is.na(OnMspendDelt_lead))


# EuropeDefSpread <- subset(EuropeDefSpread, Year>2002)


EuropeDefSpread_lag0 <- subset(EuropeDefSpread, !is.na(DefSpread))
EuropeDefSpread_lag1 <- subset(EuropeDefSpread, !is.na(DefSpread_lag1))
EuropeDefSpread_lag2 <- subset(EuropeDefSpread, !is.na(DefSpread_lag2))

pdata <- plm.data(EuropeDefSpread_lag2, index=c("Country", "Year"))


```


**OLS model adding each variable individually**
```{r OnMspreadOLS, echo = TRUE}
OnMEUDefSpreadResults1 <- lm(OnMspendDelt ~ DefSpread, EuropeDefSpread_lag0) 
summary(OnMEUDefSpreadResults1)

OnMEUDefSpreadResults2 <- lm(OnMspendDelt ~ DefSpread_lag1, EuropeDefSpread_lag1) 
summary(OnMEUDefSpreadResults2)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2),
          custom.model.names=c("PubOp_df1","PubOp_df2"))

OnMEUDefSpreadResults3 <- lm(OnMspendDelt ~ DefSpread_lag2, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults3)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3"))

##use 2 lag

OnMEUDefSpreadResults4 <- lm(OnMspendDelt ~ DefSpread_lag2 + IntAt, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults4)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults4),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "IntAt"))

#intAT not significant won't include

OnMEUDefSpreadResults5 <- lm(OnMspendDelt ~ DefSpread_lag2 + CivilWar, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults5)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults5),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "CivilWar"))

##civilwar isn't significant, won't include


OnMEUDefSpreadResults6 <- lm(OnMspendDelt ~ DefSpread_lag2 + IntlCnf, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults6)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults6),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "IntCnf"))

## IntCnf not significant won't include


OnMEUDefSpreadResults7 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults7)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDPpc"))
```


```{r Macroeconomics}
##GDP doesn't increase r^2 by more htan .02 points and is statistically significant, will be includee 

OnMEUDefSpreadResults8 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + Cab_left_right, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults8)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults8),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDPpCap", "Cableftright"))
```

```{r Cabinet}
##cableft right not significant will not include

OnMEUDefSpreadResults9 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + Cab_liberty_authority, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults9)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults9),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3","GDPpCapDelt", "lib_authority"))

##cab liberty authority not significant won't be included

OnMEUDefSpreadResults10 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + Cab_eu_anti_pro, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults10)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults10),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDPpCap", "euAntiPro"))

##not significant, won't be included eu anti pro

OnMEUDefSpreadResults11 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults11)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS"))

##let_right_ls spread is significant, will be included 

OnMEUDefSpreadResults12 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCap + left_right_ls_spread + liberty_authority_ls_spread, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults12)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11, OnMEUDefSpreadResults12),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS", "libauthLS"))
##libauthLS isn't signficicant won't be included

OnMEUDefSpreadResults13 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCap + left_right_ls_spread + eu_anti_pro_ls_spread, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults13)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11, OnMEUDefSpreadResults13),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS", "euantiproLS"))
##eu anti pro isn't signficicant won't be included

```

##results 11 is final model. 

**interaction with left_right_ls_spread and Defspread**
``` {r OnMspreadInteractiontTest, echo = TRUE}

##center variables

EuropeDefSpread_lag2$left_right_ls_spreadC <- ((EuropeDefSpread_lag2$left_right_ls_spread) - mean(EuropeDefSpread_lag2$left_right_ls_spread))

EuropeDefSpread_lag2$DefSpread_lag2C <- ((EuropeDefSpread_lag2$DefSpread_lag2) - mean(EuropeDefSpread_lag2$DefSpread_lag2))

EuropeDefSpread_lag2$intrleft_right_ls_DefSread <- (EuropeDefSpread_lag2$DefSpread_lag2C * EuropeDefSpread_lag2$left_right_ls_spreadC)

OnMEUDefSpreadResults14 <- lm(OnMspendDelt ~ DefSpread_lag2C + GDPpCapDelt + left_right_ls_spreadC + intrleft_right_ls_DefSread, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults14)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11, OnMEUDefSpreadResults14),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GdP", "leftrightLS", "interaction"))


##interaction term is not significant and will not be used. 

```

**Tests for panel data model.**

``` {r defspreadPoolingTest, echo = TRUE}

OnMEUDefSpreadResults11 <- lm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread, EuropeDefSpread_lag2) 
summary(OnMEUDefSpreadResults11)
screenreg(list(OnMEUDefSpreadResults1,OnMEUDefSpreadResults2, OnMEUDefSpreadResults3, OnMEUDefSpreadResults7, OnMEUDefSpreadResults11),
          custom.model.names=c("PubOp_df1","PubOp_df2", "PubOp_df3", "GDP", "leftrightLS"))


OnMspreadPooled <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread, data = pdata, model = "pooling")

screenreg(list(OnMEUDefSpreadResults11, OnMspreadPooled),
          custom.model.names = c("OLS", "Pooled"))

OnMspreadBetween <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread, data = pdata, model = "between")
summary(OnMspreadBetween)
screenreg(list(OnMEUDefSpreadResults11, OnMspreadPooled, OnMspreadBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

OnMspreadFD <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread, data = pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(OnMEUDefSpreadResults11, OnMspreadPooled, OnMspreadBetween, OnMspreadFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


OnMspreadWithin <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread, data =  pdata, model = "within")
summary(OnMspreadWithin)
screenreg(list(OnMEUDefSpreadResults11, OnMspreadPooled, OnMspreadBetween, OnMspreadFD, OnMspreadWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


OnMspreadRandom <- plm(OnMspendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread, data = pdata, model = "random")
screenreg(list(OnMEUDefSpreadResults11, OnMspreadPooled, OnMspreadBetween, OnMspreadFD, OnMspreadWithin, OnMspreadRandom),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))
summary(OnMspreadRandom)

##lagrange multiplier (lm) for random effects vs. OLS
plmtest(OnMspreadPooled) #if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"

##fixed effects versus OLS model
pFtest(OnMspreadWithin, OnMspreadPooled)
##p = .168 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
phtest(OnMspreadRandom, OnMspreadWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```

- plmtset, p = 4.01e-9, p value small so reject the null hypothesis that pooling model works and potentially random effects is best. 

- pftest p = .009, reject the null that the pooling model is best, which suggest to use the OLS model

- phtest, p = .9099, cannot reject the null hypothesis that random effects is best

##use random effects
