---
title: "NATO Essential and Defense Spending"
author: "src"
date: "October 19, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

#Setup
##Load Packages and Support Files
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
 #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


##Load the Data
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()

write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <- subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))




write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )



```


##Subset the Data
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EuropeNATOessSpread <- subset(EuropeOmnibus, !is.na(DefSpendDelt_lead))

EuropeNATOessSpread_gap1 <- subset(EuropeNATOessSpread, !is.na(NATOessSpread))
EuropeNATOessSpread_gap2 <- subset(EuropeNATOessSpread, !is.na(NATOessSpread_lag1))
EuropeNATOessSpread_gap3 <- subset(EuropeNATOessSpread, !is.na(NATOessSpread_lag2))



```


#Building the Regression Model
##Choosing the Public Opinion Lag
``` {r PublicOpinionLag, echo = TRUE}
NATOessSpreadDef_model<-list()
NATOessSpreadDef_name<-c()

reg<-1
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread, EuropeNATOessSpread_gap1))
NATOessSpreadDef_name[reg]<-"PubOp_ne1"
screenreg(NATOessSpreadDef_model[reg],
          custom.model.name=NATOessSpreadDef_name[reg])
summary(NATOessSpreadDef_model[[reg]])

reg<-2
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag1, EuropeNATOessSpread_gap2))
NATOessSpreadDef_name[reg]<-"PubOp_ne2"
screenreg(NATOessSpreadDef_model[reg],
          custom.model.name=NATOessSpreadDef_name[reg])
summary(NATOessSpreadDef_model[[reg]])

reg<-3
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"PubOp_ne3"
screenreg(NATOessSpreadDef_model[reg],
          custom.model.name=NATOessSpreadDef_name[reg])
summary(NATOessSpreadDef_model[[reg]])

screenreg(NATOessSpreadDef_model[c(1:3)],
          custom.model.name=NATOessSpreadDef_name[c(1:3)])

EuropeNATOessSpread_gap3_pdata <- plm.data(EuropeNATOessSpread_gap3, index=c("Country", "Year"))
```

After comparing the three options, the three year difference between expenditure and Nato Essential polling appears to have the strongest relationship, even if it no where close to significant.

##Security Environment Controls
``` {r SecurityEnvironment, echo = TRUE}
reg<-4
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + IntAt, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"IntAt"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Drop in adjusted R^2, no significance.

reg<-5
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + DomAt, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"DomAt"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

reg<-6
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2 +  CivilWar, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"CivilWar"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

##Removing civil war, not significant,  small increase in R^2 and kills the pvalue of  NATOessSpread
reg<-7
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +IntlCnf, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"IntlCnf"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Small improvement from international conflict, but reduces p-value, so we'll leave it out.


```

Minimal gains from the security environment variables, and the signs were all in an unexpected direction.

##Macroeconomic Controls
``` {r MacroEconomics, echo = TRUE}

reg<-8
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + GDPpCapLCUdelt, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"GDPpCapLCUdelt"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Negative adjusted R^2, leaving out.




```

Not much changed, I substituted in GPDpCap as it wasn't much more powerful than Civil War, but had a better p value and the outcome made more sense. 

##Parliamentary Opinion Controls
``` {r Cabinet, echo = TRUE}
reg<-9
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +Cab_left_right, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"CabLeftRight"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Minimal gains, odd sign.

reg<-10
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +Cab_liberty_authority, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"CabLibertyAuth"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Not significant. More of a gain but still, odd sign and reduces p-value. Rejecting.

reg<-11
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +Cab_eu_anti_pro, EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"CabEUantiPro"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#No improvement in R^2. Does help p value, but not significant.


reg<-12
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2 + left_right_ls_spread  , EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"lr_spread"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Not significant, improvement in r^2 but reduces p-value of stud variable.


reg<-13
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2 +  liberty_authority_ls_spread , EuropeNATOessSpread_gap3))
NATOessSpreadDef_name[reg]<-"LA_spread"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])
#Lowers adj r2 to no benefit. Not significant.

#liberty_authority spread rejected
reg<-14
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2   , EuropeNATOessSpread_gap3))
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#EU anti/pro also rejected.

```


NATOessSpreadDef_model[3] is chosen.



#Determining Panel Data Model

``` {r defspreadPoolingTest, echo = TRUE}

NATOessSpreadDef_model3 <- lm(DefSpendDelt_lead ~ NATOessSpread_lag2, EuropeNATOessSpread_gap3)


NATOessSpreadDEfPooled <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2, data = EuropeNATOessSpread_gap3_pdata, model = "pooling")

summary(NATOessSpreadDEfPooled)

screenreg(list(NATOessSpreadDef_model3, NATOessSpreadDEfPooled),
          custom.model.names = c("OLS", "Pooled"))

NATOessSpreadDEfBetween <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2, data = EuropeNATOessSpread_gap3_pdata, model = "between")
summary(NATOessSpreadDEfBetween)
screenreg(list(NATOessSpreadDef_model3, NATOessSpreadDEfPooled, NATOessSpreadDEfBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

NATOessSpreadDEfFD <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2, data = EuropeNATOessSpread_gap3_pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(NATOessSpreadDef_model3, NATOessSpreadDEfPooled, NATOessSpreadDEfBetween, NATOessSpreadDEfFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


NATOessSpreadDEfWithin <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2, data =  EuropeNATOessSpread_gap3_pdata, model = "within")
summary(NATOessSpreadDEfWithin)
screenreg(list(NATOessSpreadDef_model3, NATOessSpreadDEfPooled, NATOessSpreadDEfBetween, NATOessSpreadDEfFD, NATOessSpreadDEfWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


NATOessSpreadDEfRandom <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2, data = EuropeNATOessSpread_gap3_pdata, model = "random")
screenreg(list(NATOessSpreadDef_model3, NATOessSpreadDEfPooled, NATOessSpreadDEfBetween, NATOessSpreadDEfFD, NATOessSpreadDEfWithin, NATOessSpreadDEfRandom),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))


##lagrange multiplier (lm) for random effects vs. OLS
plmtest(NATOessSpreadDEfPooled)
#p-value 0.1433
#if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"

##fixed effects versus OLS model
pFtest(NATOessSpreadDEfWithin, NATOessSpreadDEfPooled)
##p = 0.07781 
# Possibly can reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
phtest(NATOessSpreadDEfRandom, NATOessSpreadDEfWithin)

##0.8074
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```

-plm test: p=.143, cannot reject null that pooled model is best

-pftest: p = .07, 90% significance to support using within, but since results show within as 0 r^2 and pooled as .01 r^2 we should use pooled

**pooled is final**
