---
title: "Untitled"
author: "src"
date: "October 19, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
#setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))




write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


**load and subset the data we want**
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EuropeNATOessSpread <- subset(EuropeOmnibus, !is.na(DefSpendDelt_lead))

EuropeNATOessSpread_lag0 <- subset(EuropeNATOessSpread, !is.na(NATOessSpread))
EuropeNATOessSpread_lag1 <- subset(EuropeNATOessSpread, !is.na(NATOessSpread_lag1))
EuropeNATOessSpread_lag2<- subset(EuropeNATOessSpread, !is.na(NATOessSpread_lag2))







```



``` {r PublicOpinionLag, echo = TRUE}
NATOessSpreadDef_model<-list()
NATOessSpreadDef_name<-c()

reg<-1
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread, EuropeNATOessSpread_lag0))
NATOessSpreadDef_name[reg]<-"PubOp_ne1"
screenreg(NATOessSpreadDef_model[reg],
          custom.model.name=NATOessSpreadDef_name[reg])
summary(NATOessSpreadDef_model[[reg]])

reg<-2
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag1, EuropeNATOessSpread_lag1))
NATOessSpreadDef_name[reg]<-"PubOp_ne2"
screenreg(NATOessSpreadDef_model[reg],
          custom.model.name=NATOessSpreadDef_name[reg])
summary(NATOessSpreadDef_model[[reg]])

reg<-3
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"PubOp_ne3"
screenreg(NATOessSpreadDef_model[reg],
          custom.model.name=NATOessSpreadDef_name[reg])
summary(NATOessSpreadDef_model[[reg]])

screenreg(NATOessSpreadDef_model[c(1:3)],
          custom.model.name=NATOessSpreadDef_name[c(1:3)])


```

After comparing the three options, the three year difference between expenditure and Nato Essential polling appears to have the strongest relationship, even if it no where close to significant.


``` {r SecurityEnvironment, echo = TRUE}
reg<-4
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + IntAt, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"IntAt"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Drop in adjusted R^2, no significance.

reg<-5
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + DomAt, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"DomAt"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

reg<-6
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2 +  CivilWar, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"CivilWar"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

##Removing civil war, not significant,  small increase in R^2 and kills the pvalue of  NATOessSpread
reg<-7
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +IntlCnf, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"IntlCnf"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Small improvement from international conflict, but reduces p-value, so we'll leave it out.


```

Minimal gains from the security environment variables, and the signs were all in an unexpected direction.


``` {r MacroEconomics, echo = TRUE}

reg<-8
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + GDPpCapDelt, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"GDPpCapDelt"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Negative adjusted R^2, leaving out.




```

Not much changed, I substituted in GPDpCap as it wasn't much more powerful than Civil War, but had a better p value and the outcome made more sense. 


``` {r Cabinet, echo = TRUE}
reg<-9
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +Cab_left_right, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"CabLeftRight"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Minimal gains, odd sign.

reg<-10
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +Cab_liberty_authority, EuropeNATOessSpread_lag2))
NATOessSpreadDef_name[reg]<-"CabLibertyAuth"
screenreg(NATOessSpreadDef_model[c(3,reg)],
          custom.model.name=NATOessSpreadDef_name[c(3,reg)])
summary(NATOessSpreadDef_model[[reg]])

#Not significant. More of a gain but still, odd sign and reduces p-value. Rejecting.

reg<-11
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +Cab_eu_anti_pro, EuropeNATOessSpread_lag2))
screenreg(list(NATOessSpreadDef_model[1],NATOessSpreadDef_model[2],NATOessSpreadDef_model[3],NATOessSpreadDef_model[10],NATOessSpreadDef_model[11]),
           custom.model.names=c("PubOp_df1","PubOp_ne2","PubOp_df3","CabLibAuth",
                            "CabEUantiPro"))
summary(NATOessSpreadDef_model[[reg]])

#Weaker support than Libertarian authority and still weird sign.


reg<-12
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + Cab_liberty_authority+ left_right_ls_spread  , EuropeNATOessSpread_lag2))
screenreg(list(NATOessSpreadDef_model[1],NATOessSpreadDef_model[2],NATOessSpreadDef_model[3],NATOessSpreadDef_model[10],NATOessSpreadDef_model[12]),
           custom.model.names=c("PubOp_df1","PubOp_ne2","PubOp_df3","CabLibAuth",
                            "lr_spread"))
summary(NATOessSpreadDef_model[[reg]])

#Slight improvement, lets compare to alternatives
reg<-13
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  +Cab_liberty_authority + liberty_authority_ls_spread , EuropeNATOessSpread_lag2))
screenreg(list(NATOessSpreadDef_model[1],NATOessSpreadDef_model[2],NATOessSpreadDef_model[3],NATOessSpreadDef_model[10],NATOessSpreadDef_model[12],NATOessSpreadDef_model[13]),
           custom.model.names=c("PubOp_df1","PubOp_ne2","PubOp_df3","CabLibAuth",
                            "lr_spread","la_spread"))
summary(NATOessSpreadDef_model[[reg]])


#liberty_authority spread rejected
reg<-14
NATOessSpreadDef_model[reg] <- list(lm(DefSpendDelt_lead ~ NATOessSpread_lag2  + Cab_liberty_authority+ eu_anti_pro_ls_spread  , EuropeNATOessSpread_lag2))
screenreg(list(NATOessSpreadDef_model[1],NATOessSpreadDef_model[2],NATOessSpreadDef_model[3],NATOessSpreadDef_model[10],NATOessSpreadDef_model[12],NATOessSpreadDef_model[13],NATOessSpreadDef_model[14]),
           custom.model.names=c("PubOp_df1","PubOp_ne2","PubOp_df3","CabLibAuth",
                            "lr_spread","la_spread","eu_spread"))
summary(NATOessSpreadDef_model[[reg]])

#EU anti/pro also rejected.

```


NATOessSpreadDef_model[13] is chosen.



**Tests of within and pooling** we should explain this in the 

``` {r defspreadPoolingTest, echo = TRUE}


screenreg(list(NATOessSpreadDef_model[1],NATOessSpreadDef_model[2],NATOessSpreadDef_model[3],NATOessSpreadDef_model[10],NATOessSpreadDef_model[12]),
           custom.model.names=c("PubOp_df1","PubOp_ne2","PubOp_df3","CabLibAuth",
                            "lr_spread"))
summary(NATOessSpreadDef_model[12])

# VariablesInModel=subset(EuropeNATOessSpread_lag2,select=c(Country,Year,DefSpendDelt_lead , NATOessSpread_lag2 ,  IntlCnf , GDPpCapDelt , Cab_liberty_authority, eu_anti_pro_ls_spread ),Year>2004)
# VariablesInModel$Country<-as.factor(VariablesInModel$Country)
# summary(VariablesInModel$Country)

NATOessPooled <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2  + Cab_liberty_authority+ left_right_ls_spread  , EuropeNATOessSpread_lag2, data = EuropeNATOessSpread_lag2, model = "pooling")
summary(NATOessPooled)
NATOessWithin <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2  + Cab_liberty_authority+ left_right_ls_spread  , EuropeNATOessSpread_lag2, data = EuropeNATOessSpread_lag2,  model = "within")
summary(NATOessWithin)

pFtest(NATOessPooled, NATOessWithin)
##p = .1258 cannot  reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```


**Fixed Effects model** (within)
```{r DefSpreadFixed, echo = TRUE}

NatoEssResultsFixed <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2  + Cab_liberty_authority+ eu_anti_pro_ls_spread, data = VariablesInModel,  model = "within")
summary(NatoEssResultsFixed)

screenreg(list(NATOessSpreadDef_model[14], NatoEssResultsFixed))

#testing for fixed effects
# pFtest(NATOessSpreadDef_modelA, NATOessSpreadDef_model8)

##p = .1358,, cannot reject null hypothesis that there are not significnat effects, use pooled 

```

**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r DefSperadHausmanTest, echo = TRUE}

NATOessSpreadDef_modelWithin <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2  + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATOessSpread_lag2, data = VariablesInModel,  index=c("Country", "Year"), model="within") 

NATOessSpreadDef_modelRandom <- plm(DefSpendDelt_lead ~ NATOessSpread_lag2  + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATOessSpread_lag2, data = VariablesInModel,  index=c("Country", "Year"), model="random")

phtest(NATOessSpreadDef_modelWithin, NATOessSpreadDef_modelRandom)
##p .7333 cannot reject null, which makes sense because it told us before to do pooled. 
```

