---
title: "NATO and EU Convergence and Equipment Spending"
author: "src"
date: "October 19, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
#setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
 setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}


# debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))




write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


**load and subset the data we want**
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EquNATO_EUspread <- subset(EuropeOmnibus, !is.na(EquSpendDelt_lead))

EquNATO_EUspread_lag0 <- subset(EquNATO_EUspread, !is.na(NATO.EUspread))
EquNATO_EUspread_lag1 <- subset(EquNATO_EUspread, !is.na(NATO.EUspread_lag1))
EquNATO_EUspread_lag2<- subset(EquNATO_EUspread, !is.na(NATO.EUspread_lag2))



pdata <- plm.data(EquNATO_EUspread_lag2, index=c("Country", "Year"))

```



``` {r PublicOpinionLag, echo = TRUE}
NATO.EUspreadResults_model<-list()
NATO.EUspreadResults_name<-c()

NATO.EUspreadResults_model[1] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread, EquNATO_EUspread_lag0) )
NATO.EUspreadResults_name[1]<-"PubOp_ne1"
screenreg(NATO.EUspreadResults_model[1],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[1])
summary(NATO.EUspreadResults_model[[1]])
plot(NATO.EUspreadResults_model[[1]])

NATO.EUspreadResults_model[2] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag1, EquNATO_EUspread_lag1))
NATO.EUspreadResults_name[2]<-"PubOp_ne2"
screenreg(NATO.EUspreadResults_model[c(1:2)],
          custom.model.name=NATO.EUspreadResults_name[c(1:2)])
summary(NATO.EUspreadResults_model[[2]])
plot(NATO.EUspreadResults_model[[2]])

NATO.EUspreadResults_model[3] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[2]<-"PubOp_ne2"
screenreg(NATO.EUspreadResults_model[c(1:3)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3)])
summary(NATO.EUspreadResults_model[[3]])
plot(NATO.EUspreadResults_model[[3]])


#The two year difference in PubOp_ne2 has the largest coefficient and the smallest p-value.

```

After comparing the three options, the three year difference between expenditure and Nato Essential polling appears to have the strongest relationship, even if it no where close to significant.


``` {r SecurityEnvironment, echo = TRUE}

reg<-4
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + IntAt, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"IntAt"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Lowers adjusted R^2 intl attack will not be used.

reg<-5
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + DomAt, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"DomAt"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

# Lower domestic attack R^2, will not use.

reg<-6
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + CivilWar, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"CivilWar"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])
summary(EquNATO_EUspread_lag2$CivilWar)

#NATOessential removes Civil war cases 

reg<-7
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + IntlCnf, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"IntlCnf"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])
summary(EuropeOmnibus$IntlCnf)

#NATOessential  closer  removes IntlCnf cases 

```

Minimal gains from the security environment variables, and the signs were all in an unexpected direction.


``` {r MacroEconomics, echo = TRUE}


reg<-8
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"GDPpCapDelt"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])
plot(NATO.EUspreadResults_model[[reg]])
#Shrinks adjusted r^2, not significant. Leaving out.



```

Added GDPpCapDelt which had a sign that made sense and improved p value of the lag, even if it doesn't help the R^2

``` {r Parliament, echo = TRUE}


reg<-9
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + Cab_left_right, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"Left_Right"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Shrinks adjusted r^2, not significant. Leaving out  Dropping Cab_left_right


reg<-10
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + Cab_liberty_authority, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"Lib_Auth"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Lower adjusted R^2, not significant and doesn't make sense. Dropping Cab_liberty_authority

reg<-11
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + Cab_eu_anti_pro, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"EU_anti_pro"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Lower adjusted R^2. Dropping Cab_eu_anti_pro


reg<-12
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + left_right_ls_spread, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"Left_right_dif"
screenreg(NATO.EUspreadResults_model[c(1:3,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#Small gain, drop in adjusted R^2, small p-value gain. Leaving out

reg<-13
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + liberty_authority_ls_spread, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"la_spread"
screenreg(NATO.EUspreadResults_model[c(1:3,12,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,reg)])
summary(NATO.EUspreadResults_model[[reg]])

#No real change to adjusted R-squared but decreases pvalue of study variable. Leaving out.

reg<-14
NATO.EUspreadResults_model[reg] <- list(lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread, EquNATO_EUspread_lag2))
NATO.EUspreadResults_name[reg]<-"eu_spread"
screenreg(NATO.EUspreadResults_model[c(1:3,12,13,reg)],
          custom.NATO.EUspreadResults_name=NATO.EUspreadResults_name[c(1:3,12,13,reg)])
summary(NATO.EUspreadResults_model[[reg]])
plot(NATO.EUspreadResults_model[[reg]])
#Adjusted R^2 steady and lowers the p value on study variable. Seems worth keeping


```

Model #14 is chosen, but is not significant.



**Tests for panel data model.**

``` {r defspreadPoolingTest, echo = TRUE}

NATOEUSpreadResults_model14 <- lm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread, EquNATO_EUspread_lag2)


EquSpreadPooled <- plm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread, data = pdata, model = "pooling")
screenreg(list(NATOEUSpreadResults_model14, EquSpreadPooled),
          custom.model.names = c("OLS", "Pooled"))

EquSpreadBetween <- plm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread, data = pdata, model = "between")
summary(EquSpreadBetween)
screenreg(list(NATOEUSpreadResults_model14, EquSpreadPooled, EquSpreadBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

EquSpreadFD <- plm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread, data = pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(NATOEUSpreadResults_model14, EquSpreadPooled, EquSpreadBetween, EquSpreadFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


EquSpreadWithin <- plm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread, data =  pdata, model = "within")
summary(EquSpreadWithin)
screenreg(list(NATOEUSpreadResults_model14, EquSpreadPooled, EquSpreadBetween, EquSpreadFD, EquSpreadWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


# EquDefSpreadRandom <- plm(EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread, data = pdata, model = "random")
# screenreg(list(NATOEUSpreadResults_model14, DefSpreadPooled, DefSpreadBetween, DefSpreadFD, DefSpreadWithin, DefSpreadRandom),
#           custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))


##lagrange multiplier (lm) for random effects vs. OLS
plmtest(EquSpreadPooled) 
#if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"
# p- 0.4439

##fixed effects versus OLS model
pFtest(EquSpreadWithin, EquSpreadPooled)
##p = 0.6879 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
# phtest(EquSpreadRandom, EquSpreadWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```


P values are huge, sticking with pooled.