---
title: "NATO Essential and Equipment Spending"
author: "src"
date: "October 19, 2015"
output:
  html_document:
    number_sections: yes
    toc: yes
---

#Setup ##Load Packages and Support Files
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
#setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
 #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


##Load the Data
```{r Input, echo = FALSE, results = "hide"}


# debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <- subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))




write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


##Subset the Data
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EquNATOessSpread <- subset(EuropeOmnibus, !is.na(EquSpendDelt_lead))

EquNATOessSpread_gap1 <- subset(EquNATOessSpread, !is.na(NATOessSpread))
EquNATOessSpread_gap2 <- subset(EquNATOessSpread, !is.na(NATOessSpread_lag1))
EquNATOessSpread_gap3<- subset(EquNATOessSpread, !is.na(NATOessSpread_lag2))


```

#Building the Regression Model
##Choosing the Public Opinion Lag


``` {r PublicOpinionLag, echo = TRUE}
NATOessEquSpread_model<-list()
NATOessEquSpread_name<-c()

NATOessEquSpread_model[1] <- list(lm(EquSpendDelt_lead ~ NATOessSpread, EquNATOessSpread_gap1) )
NATOessEquSpread_name[1]<-"PubOp_ne1"
screenreg(NATOessEquSpread_model[1],
          custom.model.name=NATOessEquSpread_name[1])
summary(NATOessEquSpread_model[[1]])
plot(NATOessEquSpread_model[[1]])

NATOessEquSpread_model[2] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1, EquNATOessSpread_gap2))
NATOessEquSpread_name[2]<-"PubOp_ne2"
screenreg(NATOessEquSpread_model[c(1:2)],
          custom.model.name=NATOessEquSpread_name[c(1:2)])
summary(NATOessEquSpread_model[[2]])
plot(NATOessEquSpread_model[[2]])

NATOessEquSpread_model[3] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag2, EquNATOessSpread_gap3))
NATOessEquSpread_name[3]<-"PubOp_ne3"
screenreg(NATOessEquSpread_model[c(1:3)],
          custom.model.name=NATOessEquSpread_name[c(1:3)])
summary(NATOessEquSpread_model[[3]])
plot(NATOessEquSpread_model[[3]])

EquNATOessSpread_gap2_pdata <- plm.data(EquNATOessSpread_gap2, index=c("Country", "Year"))

#The two year difference in PubOp_ne2 has the largest coefficient and the smallest p-value.

```

After comparing the three options, the three year difference between expenditure and Nato Essential polling appears to have the strongest relationship, even if it no where close to significant.

##Security Environment Controls
``` {r SecurityEnvironment, echo = TRUE}


NATOessEquSpread_model[4] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + IntAt, EquNATOessSpread_gap2))
NATOessEquSpread_name[4]<-"IntAt"
screenreg(NATOessEquSpread_model[c(1:3,4)],
          custom.model.name=NATOessEquSpread_name[c(1:3,4)])
summary(NATOessEquSpread_model[[4]])

#Lowers adjusted R^2 intl attack will not be used.

NATOessEquSpread_model[5] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + DomAt, EquNATOessSpread_gap2))
NATOessEquSpread_name[5]<-"DomAt"
screenreg(NATOessEquSpread_model[c(1:3,5)],
          custom.model.name=NATOessEquSpread_name[c(1:3,5)])
summary(NATOessEquSpread_model[[5]])

# Lower domestic attack R^2, will not use.

NATOessEquSpread_model[6] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar, EquNATOessSpread_gap2))
NATOessEquSpread_name[6]<-"CivilWar"
screenreg(NATOessEquSpread_model[c(1:3,6)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6)])
summary(NATOessEquSpread_model[[6]])
summary(EquNATOessSpread_gap2$CivilWar)

#Civil war increases adj R^2 by a tiny amount and and improve p-value, leaving in.


NATOessEquSpread_model[7] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +IntlCnf, EquNATOessSpread_gap2))
NATOessEquSpread_name[7]<-"IntlCnf"
screenreg(NATOessEquSpread_model[c(1:3,6,7)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,7)])
summary(NATOessEquSpread_model[[7]])
summary(EuropeOmnibus$IntlCnf)

#Not significant. Decreases Adjusted R^2

```

Minimal gains from the security environment variables, and the signs were all in an unexpected direction.

##Macroeconomic Controls
``` {r MacroEconomics, echo = TRUE}


reg<-8
NATOessEquSpread_model[reg] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +GDPpCapUSDdelt, EquNATOessSpread_gap2))
NATOessEquSpread_name[reg]<-"GDPpCapUSDdelt"
screenreg(NATOessEquSpread_model[c(1:3,6,reg)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,reg)])
summary(NATOessEquSpread_model[[reg]])

#Shrinks adjusted r^2, not significant. Leaving out.



```

Tempted to add GDPpCapUSDdelt which had a sign that made sense and improved p value of the lag, but reduces the adjusted R^2.

##Parliamentary Opinion Controls
``` {r Parliament, echo = TRUE}


reg<-9
NATOessEquSpread_model[reg] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +Cab_left_right, EquNATOessSpread_gap2))
NATOessEquSpread_name[reg]<-"Left_Right"
screenreg(NATOessEquSpread_model[c(1:3,6,reg)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,reg)])
summary(NATOessEquSpread_model[[reg]])

#Shrinks adjusted r^2, not significant. Leaving out  Dropping Cab_left_right


reg<-10
NATOessEquSpread_model[reg] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +Cab_liberty_authority, EquNATOessSpread_gap2))
NATOessEquSpread_name[reg]<-"Lib_Auth"
screenreg(NATOessEquSpread_model[c(1:3,6,reg)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,reg)])
summary(NATOessEquSpread_model[[reg]])

#Lower adjusted R^2, not significant and doesn't make sense. Dropping Cab_liberty_authority

reg<-11
NATOessEquSpread_model[reg] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +Cab_eu_anti_pro, EquNATOessSpread_gap2))
NATOessEquSpread_name[reg]<-"EU_anti_pro"
screenreg(NATOessEquSpread_model[c(1:3,6,reg)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,reg)])
summary(NATOessEquSpread_model[[reg]])
plot(NATOessEquSpread_model[[reg]])
#Slightly lowers p-value, but doesn't raise adjusted R^2, and doesn't help with residuals vs. leverage.


reg<-12
NATOessEquSpread_model[reg] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +left_right_ls_spread, EquNATOessSpread_gap2))
NATOessEquSpread_name[reg]<-"Left_right_dif"
screenreg(NATOessEquSpread_model[c(1:3,6,reg)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,reg)])
summary(NATOessEquSpread_model[[reg]])
plot(NATOessEquSpread_model[[reg]])
#Not significant, lowers adjusted R^2, small p-value gain. Leaving out.

reg<-13
NATOessEquSpread_model[reg] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +liberty_authority_ls_spread, EquNATOessSpread_gap2))
NATOessEquSpread_name[reg]<-"la_spread"
screenreg(NATOessEquSpread_model[c(1:3,6,reg)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,reg)])
summary(NATOessEquSpread_model[[reg]])

#Not significant, Reduces  adjusted R-squared and  decreases pvalue of study variable. Leaving out.

reg<-14
NATOessEquSpread_model[reg] <- list(lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar +eu_anti_pro_ls_spread, EquNATOessSpread_gap2))
NATOessEquSpread_name[reg]<-"eu_spread"
screenreg(NATOessEquSpread_model[c(1:3,6,12,13,reg)],
          custom.model.name=NATOessEquSpread_name[c(1:3,6,8,13,reg)])
summary(NATOessEquSpread_model[[reg]])
plot(NATOessEquSpread_model[[3]])
#Not significant, Reduces  adjusted R-squared and  decreases pvalue of study variable. Leaving out.


```

Model #6 is chosen, but is not significant. Not checking interactions because there's no significant variables to check with.


#Determining Panel Data Model

``` {r defspreadPoolingTest, echo = TRUE}

NATOessSpreadEqu_model6 <- lm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar, EquNATOessSpread_gap2)


NATOessSpreadEquPooled <- plm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar, data = EquNATOessSpread_gap2_pdata, model = "pooling")

screenreg(list(NATOessSpreadEqu_model6, NATOessSpreadEquPooled),
          custom.model.names = c("OLS", "Pooled"))
summary(NATOessSpreadEquPooled)

NATOessSpreadEquBetween <- plm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar, data = EquNATOessSpread_gap2_pdata, model = "between")
summary(NATOessSpreadEquBetween)
screenreg(list(NATOessSpreadEqu_model6, NATOessSpreadEquPooled, NATOessSpreadEquBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

summary(NATOessSpreadEquBetween)

NATOessSpreadEquFD <- plm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar, data = EquNATOessSpread_gap2_pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(NATOessSpreadEqu_model6, NATOessSpreadEquPooled, NATOessSpreadEquBetween, NATOessSpreadEquFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


NATOessSpreadEquWithin <- plm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar, data =  EquNATOessSpread_gap2_pdata, model = "within")
summary(NATOessSpreadEquWithin)
screenreg(list(NATOessSpreadEqu_model6, NATOessSpreadEquPooled, NATOessSpreadEquBetween, NATOessSpreadEquFD, NATOessSpreadEquWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


# NATOessSpreadEquRandom <- plm(EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar, data = EquNATOessSpread_gap2_pdata, model = "random")
# screenreg(list(NATOessSpreadEqu_model6, NATOessSpreadEquPooled, NATOessSpreadEquBetween, NATOessSpreadEquFD, NATOessSpreadEquWithin, NATOessSpreadEquRandom),
#           custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))


##lagrange multiplier (lm) for random effects vs. OLS
plmtest(NATOessSpreadEquPooled) 
#if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"
# p-value = 0.2924

##fixed effects versus OLS model
pFtest(NATOessSpreadEquWithin, NATOessSpreadEquPooled)
##p = 0.7616 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
# phtest(NATOessSpreadEquRandom, NATOessSpreadEquWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```

- plm test p = .294, cannot reject null thus used pooled

-pftest: .76, cannot reject null thus use pooled.

**Pooled model will be final**