---
title: "EU Exert Strong Leadership and Defense Spending"
author: "src"
date: "October 19, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

#Setup 
##Load Packages and Support Files
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
 #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


##Load the Data
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <- subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom", "Italy",
                                                      "Netherlands", "Poland", "Portugal","Slovakia","Spain","Turkey"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


##Subset the Data
```{r EULeadershipNarrowing, echo = TRUE, results = "hide"}

##load data using function
EULeadershipSpread <- subset(EuropeOmnibus, !is.na(DefSpendDelt_lead))


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <-subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))

EULeadershipSpread_gap1 <- subset(EULeadershipSpread, !is.na(EUldrSpread))
EULeadershipSpread_gap2 <- subset(EULeadershipSpread, !is.na(EUldrSpread_lag1))
EULeadershipSpread_gap3 <- subset(EULeadershipSpread, !is.na(EUldrSpread_lag2))


```

#Building the Regression Model
##Choosing the Public Opinion Lag
```{r EULeadershipspreadRegressions, echo = TRUE}
EULeadershipResults1 <- lm(DefSpendDelt ~ EUldrSpread, EULeadershipSpread_gap1) 
summary(EULeadershipResults1)
screenreg(list(EULeadershipResults1))

#lag0 on EUldrSpread is significant at the .05 level.

EULeadershipResults2 <- lm(DefSpendDelt ~ EUldrSpread_lag1, EULeadershipSpread_gap2) 
summary(EULeadershipResults2)
screenreg(list(EULeadershipResults1, EULeadershipResults2),
              custom.model.names=c("PubOp_df2","PubOp_df2"))

EULeadershipResults3 <- lm(DefSpendDelt ~ EUldrSpread_lag2, EULeadershipSpread_gap3) 
summary(EULeadershipResults3)
screenreg(list(EULeadershipResults1, EULeadershipResults3),
              custom.model.names=c("PubOp_df2","PubOp_df3"))

##lag 2 and three aren't significant
```

##Security Environment Controls
``` {r SecurityEnvironment, echo = TRUE}

EULeadershipResults4 <- lm(DefSpendDelt ~ EUldrSpread + IntAt, EULeadershipSpread_gap1) 
summary(EULeadershipResults4)
screenreg(list(EULeadershipResults1, EULeadershipResults4),
              custom.model.names=c("PubOp_df2","IntAt"))

EULeadershipSpread_gap1_pdata <- plm.data(EULeadershipSpread_gap1, index=c("Country", "Year"))

##not significant, IntAt won't be included.


EULeadershipResults5 <- lm(DefSpendDelt ~ EUldrSpread + CivilWar, EULeadershipSpread_gap1) 
summary(EULeadershipResults5)
screenreg(list(EULeadershipResults1, EULeadershipResults5),
              custom.model.names=c("PubOp_df2","CivilWar"))

##not si gnificant, won't be included

EULeadershipResults6 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf, EULeadershipSpread_gap1) 
summary(EULeadershipResults6)
screenreg(list(EULeadershipResults1, EULeadershipResults6),
              custom.model.names=c("PubOp_df2","IntCnf"))

##intl cnf is significant and will be included


```

##Macroeconomic Controls
```{r Macroeconomics}

EULeadershipResults0 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt, EULeadershipSpread_gap1) 
summary(EULeadershipResults0)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0),
              custom.model.names=c("PubOp_df2","IntlCnf", "GDPpc"))
#GDPpc is significant and will be included 

``` 

##Parliamentary Opinion Controls

```{r Parliamentary}
EULeadershipResults7 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + Cab_left_right, EULeadershipSpread_gap1) 
summary(EULeadershipResults7)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults7),
              custom.model.names=c("PubOp_df2", "IntlCnf", "GDPpc", "CabLeftRight"))

##cableftright not significant, won't be included

EULeadershipResults8 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + Cab_liberty_authority, EULeadershipSpread_gap1) 
summary(EULeadershipResults8)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults8),
              custom.model.names=c("PubOp_df2", "IntlCnf", "GDPpc","CabLiberty_authority"))

##not significant and won't be included, cab_liberty_authority

EULeadershipResults9 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + Cab_eu_anti_pro, EULeadershipSpread_gap1) 
summary(EULeadershipResults9)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults9),
              custom.model.names=c("PubOp_df2","IntlCnf", "GDPpc", "cab_eu_anti_pro"))

##cab_eu_anti_pro is not significant and wont be included

EULeadershipResults10 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + left_right_ls_spread, EULeadershipSpread_gap1) 
summary(EULeadershipResults10)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults10),
              custom.model.names=c("PubOp_df2", "IntlCnf", "GDPpc", "leftrightlsspread"))

##leftrightlsspread is not significant and won't be included

EULeadershipResults11 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + liberty_authority_ls_spread, EULeadershipSpread_gap1) 
summary(EULeadershipResults11)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults10),
              custom.model.names=c("PubOp_df2", "IntlCnf", "GDPpc", "libertyauthoritylsspread"))

##libertyauthority ls spread isn't significant and wont' be lincluded 

EULeadershipResults12 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, EULeadershipSpread_gap1) 
summary(EULeadershipResults12)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults12),
              custom.model.names=c("PubOp_df2", "IntlCnf", "GDPpc", "eu_anti_prolsspread"))

##eu_anti/pro is significant and will be included.


```

**model 12 is final**

###Test Interactions
Interaction with eu_anti_pro_ls and Defspread.
``` {r EquSpreadInteractiontTest, echo = TRUE}

##center variables

EULeadershipSpread_gap1$eu_anti_pro_ls_spreadC <- ((EULeadershipSpread_gap1$eu_anti_pro_ls_spread) - mean(EULeadershipSpread$eu_anti_pro_ls_spread,na.rm=TRUE))

EULeadershipSpread_gap1$EUldrSpreadC <- ((EULeadershipSpread_gap1$EUldrSpread) - mean(EULeadershipSpread_gap1$EUldrSpread,na.rm=TRUE))

EULeadershipSpread_gap1$intreu_anti_pro_ls_EUldrSread <- (EULeadershipSpread_gap1$EUldrSpreadC * EULeadershipSpread_gap1$eu_anti_pro_ls_spreadC)

EquEUDefSpreadResults14 <- lm(DefSpendDelt ~ EUldrSpreadC + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spreadC + intreu_anti_pro_ls_EUldrSread, EULeadershipSpread_gap1) 
summary(EquEUDefSpreadResults14)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults12,EquEUDefSpreadResults14),
              custom.model.names=c("PubOp_df2", "IntlCnf", "GDPpc", "eu_anti_prolsspread","interaction"))

##interaction not significant will not be included


```

##Experiment with multiple polling variables
```{r Experiments}

EULeadershipResultsDefSpread <- lm(DefSpendDelt ~ EUldrSpread + DefSpread_lag1 + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, subset(EULeadershipSpread_gap1, !is.na(DefSpread_lag1) ))
summary(EULeadershipResultsDefSpread)

cor.test(EULeadershipSpread_gap1$EUldrSpread_lag1,EULeadershipSpread_gap1$DefSpread_lag1)

#If used together, EUldrSpread is stronger abd DefSpread_lag1 is no longer significant. But they might be correlated so it's not a model we can really use.

```

#Determining Panel Data Model

``` {r defspreadPoolingTest, echo = TRUE}

EULeadershipResults12 <- lm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, EULeadershipSpread_gap1) 
summary(EULeadershipResults12)
screenreg(list(EULeadershipResults1, EULeadershipResults6, EULeadershipResults0, EULeadershipResults12),
              custom.model.names=c("PubOp_df2", "IntlCnf", "GDPpc", "eu_anti_prolsspread"))


DefSpreadPooled <- plm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, data = EULeadershipSpread_gap1_pdata, model = "pooling")

screenreg(list(EULeadershipResults12, DefSpreadPooled),
          custom.model.names = c("OLS", "Pooled"))
summary(DefSpreadPooled)

DefSpreadBetween <- plm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, data = EULeadershipSpread_gap1_pdata, model = "between")
summary(DefSpreadBetween)
screenreg(list(EULeadershipResults12, DefSpreadPooled, DefSpreadBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

DefSpreadFD <- plm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, data = EULeadershipSpread_gap1_pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(EULeadershipResults12, DefSpreadPooled, DefSpreadBetween, DefSpreadFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


DefSpreadWithin <- plm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, data =  EULeadershipSpread_gap1_pdata, model = "within")
summary(DefSpreadWithin)
screenreg(list(EULeadershipResults12, DefSpreadPooled, DefSpreadBetween, DefSpreadFD, DefSpreadWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


DefSpreadRandom <- plm(DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt + eu_anti_pro_ls_spread, data = EULeadershipSpread_gap1_pdata, model = "random")
screenreg(list(EULeadershipResults12, DefSpreadPooled, DefSpreadBetween, DefSpreadFD, DefSpreadWithin, DefSpreadRandom),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))


##lagrange multiplier (lm) for random effects vs. OLS
plmtest(DefSpreadPooled) #if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"

##fixed effects versus OLS model
pFtest(DefSpreadWithin, DefSpreadPooled)
##p = .168 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
phtest(DefSpreadRandom, DefSpreadWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```

- For lagrange multiplier, the p value is .5912 indicating that we cannot reject the null hypothesis that the pooled model is best.

- for pooled and fixed effects, p value - .168, cannot reject hypothesis that pooled is best.

- p value is .002, which means fixed effects are more efficient than random effects. 

**Use pooled.**

