---
title: "Untitled"
author: "src"
date: "October 19, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
#setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))




write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


**load and subset the data we want**
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EuropeNATO_EUspread <- subset(EuropeOmnibus, !is.na(DefSpendDelt_lead))

EuropeNATO_EUspread_lag0 <- subset(EuropeNATO_EUspread, !is.na(NATO.EUspread))
EuropeNATO_EUspread_lag1 <- subset(EuropeNATO_EUspread, !is.na(NATO.EUspread_lag1))
EuropeNATO_EUspread_lag2<- subset(EuropeNATO_EUspread, !is.na(NATO.EUspread_lag2))







```



``` {r PublicOpinionLag, echo = TRUE}
NATO.EUspreadResults<-list()
model.names<-c()

NATO.EUspreadResults[1] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread, EuropeNATO_EUspread_lag0) )
model.names[1]<-"PubOp_ne1"
screenreg(NATO.EUspreadResults[1],
          custom.model.names=model.names[1])
summary(NATO.EUspreadResults[[1]])

NATO.EUspreadResults[2] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1, EuropeNATO_EUspread_lag1))
model.names[2]<-"PubOp_ne2"
screenreg(NATO.EUspreadResults[c(1:2)],
          custom.model.names=model.names[c(1:2)])
summary(NATO.EUspreadResults[[2]])

NATO.EUspreadResults[3] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag2, EuropeNATO_EUspread_lag2))
model.names[2]<-"PubOp_ne2"
screenreg(NATO.EUspreadResults[c(1:3)],
          custom.model.names=model.names[c(1:3)])
summary(NATO.EUspreadResults[[3]])

#The two year difference in PubOp_ne2 has the largest coefficient and the smallest p-value.

```

After comparing the three options, the three year difference between expenditure and Nato Essential polling appears to have the strongest relationship, even if it no where close to significant.


``` {r SecurityEnvironment, echo = TRUE}


NATO.EUspreadResults[4] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + IntAt, EuropeNATO_EUspread_lag1))
model.names[4]<-"IntAt"
screenreg(NATO.EUspreadResults[c(1:3,4)],
          custom.model.names=model.names[c(1:3,4)])
summary(NATO.EUspreadResults[[4]])

#Lowers adjusted R^2 intl attack will not be used.

NATO.EUspreadResults[5] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + DomAt, EuropeNATO_EUspread_lag1))
model.names[5]<-"DomAt"
screenreg(NATO.EUspreadResults[c(1:3,5)],
          custom.model.names=model.names[c(1:3,5)])
summary(NATO.EUspreadResults[[5]])

# No net change in R^2.

NATO.EUspreadResults[6] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + CivilWar, EuropeNATO_EUspread_lag1))
model.names[6]<-"CivilWar"
screenreg(NATO.EUspreadResults[c(1:3,6)],
          custom.model.names=model.names[c(1:3,6)])
summary(NATO.EUspreadResults[[6]])
summary(EuropeNATO_EUspread_lag1$CivilWar)

#NATO.EU closer  removes Civil war cases 


NATO.EUspreadResults[7] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + IntlCnf, EuropeNATO_EUspread_lag1))
model.names[7]<-"IntlCnf"
screenreg(NATO.EUspreadResults[c(1:3,7)],
          custom.model.names=model.names[c(1:3,7)])
summary(NATO.EUspreadResults[[7]])
summary(EuropeOmnibus$IntlCnf)

#NATO.EU closer  removes IntlCnf cases 

```

Minimal gains from the security environment variables, and the signs were all in an unexpected direction.


``` {r MacroEconomics, echo = TRUE}


reg<-8
NATO.EUspreadResults[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt, EuropeNATO_EUspread_lag1))
model.names[reg]<-"GDPpCapDelt"
screenreg(NATO.EUspreadResults[c(1:3,reg)],
          custom.model.names=model.names[c(1:3,reg)])
summary(NATO.EUspreadResults[[reg]])

#No improves p value of lag and signs makes sesne, keeping.



```

Added GDPpCapDelt which had a sign that made sense and improved p value of the lag, even if it doesn't help the R^2

``` {r Parliament, echo = TRUE}


reg<-9
NATO.EUspreadResults[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + Cab_left_right, EuropeNATO_EUspread_lag1))
model.names[reg]<-"Left_Right"
screenreg(NATO.EUspreadResults[c(1:3,8,reg)],
          custom.model.names=model.names[c(1:3,8,reg)])
summary(NATO.EUspreadResults[[reg]])

#Lower adjusted R^2 and doesn't make sense. Dropping Cab_left_right


reg<-10
NATO.EUspreadResults[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + Cab_liberty_authority, EuropeNATO_EUspread_lag1))
model.names[reg]<-"Lib_Auth"
screenreg(NATO.EUspreadResults[c(1:3,8,reg)],
          custom.model.names=model.names[c(1:3,8,reg)])
summary(NATO.EUspreadResults[[reg]])

#Lower adjusted R^2 and doesn't make sense. Dropping Cab_liberty_authority

reg<-11
NATO.EUspreadResults[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + Cab_eu_anti_pro, EuropeNATO_EUspread_lag1))
model.names[reg]<-"EU_anti_pro"
screenreg(NATO.EUspreadResults[c(1:3,8,reg)],
          custom.model.names=model.names[c(1:3,8,reg)])
summary(NATO.EUspreadResults[[reg]])

#Lower adjusted R^2. Dropping Cab_eu_anti_pro


reg<-12
NATO.EUspreadResults[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + left_right_ls_spread, EuropeNATO_EUspread_lag1))
model.names[reg]<-"Left_right_dif"
screenreg(NATO.EUspreadResults[c(1:3,8,reg)],
          custom.model.names=model.names[c(1:3,8,reg)])
summary(NATO.EUspreadResults[[reg]])

#Small gain, but doesn't help p-value, lets compare to the alternatives.

reg<-13
NATO.EUspreadResults[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + liberty_authority_ls_spread, EuropeNATO_EUspread_lag1))
model.names[reg]<-"la_spread"
screenreg(NATO.EUspreadResults[c(1:3,8,12,reg)],
          custom.model.names=model.names[c(1:3,8,12,reg)])
summary(NATO.EUspreadResults[[reg]])

#Weaker value, but actually does more to clarify influence of polling variable.

reg<-14
NATO.EUspreadResults[reg] <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag1  + GDPpCapDelt + eu_anti_pro_ls_spread, EuropeNATO_EUspread_lag1))
model.names[reg]<-"eu_spread"
screenreg(NATO.EUspreadResults[c(1:3,8,12,13,reg)],
          custom.model.names=model.names[c(1:3,8,12,13,reg)])
summary(NATO.EUspreadResults[[reg]])

#Lowers R^2. Not significant.


```

Model #13 is chosen, but is not significant.




**Tests of within and pooling** we should explain this in the 

``` {r defspreadPoolingTest, echo = TRUE}
# 
# EUDefSpreadResults_EUspread <- list(lm(DefSpendDelt_lead ~ NATO.EUspread_lag2 +  IntlCnf + GDPpCapDelt + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATO.EUspread_lag2 , EuropeNATO_EUspread_lag2)
# 
# screenreg(list(NATO.EUspreadResults_pubop0,NATO.EUspreadResults_pubop1,NATO.EUspreadResults_pubop2,EUDefSpreadResults_CivilWar,EUDefSpreadResults_IntlCnf,EUDefSpreadResults_GDPpCapDelt,EUDefSpreadResults_GDPpCapDelt_noCW,EUDefSpreadResults_cablibauth,EUDefSpreadResults_LRspread,EUDefSpreadResults_LAspread),
#            custom.model.names=c("PubOp_df1","PubOp_ne2","PubOp_df3","CivilWar","IntlCnf","GDPpCapDelt","GnoCW","CabLibAuth",
#                             "lr_spread","la_spread"))
# summary(EUDefSpreadResults_LAspread)
# 
# VariablesInModel=subset(EuropeNATO_EUspread_lag2,select=c(Country,Year,DefSpendDelt_lead , NATO.EUspread_lag2 ,  IntlCnf , GDPpCapDelt , Cab_liberty_authority, eu_anti_pro_ls_spread ),Year>2004)
# VariablesInModel$Country<-as.factor(VariablesInModel$Country)
# summary(VariablesInModel$Country)
# 
# NATOcnvPooled <- list(plm(DefSpendDelt_lead ~ NATO.EUspread_lag2 +  IntlCnf + GDPpCapDelt + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATO.EUspread_lag2, data = VariablesInModel, model = "pooling")
# summary(NATOcnvPooled)
# NATOcnvWithin <- list(plm(DefSpendDelt_lead ~ NATO.EUspread_lag2 +  IntlCnf + GDPpCapDelt + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATO.EUspread_lag2, data = VariablesInModel,  model = "within")
# summary(NATOcnvWithin)
# 
# pFtest(NATOcnvPooled, NATOcnvWithin)
# ##p = .1258 cannot  reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```


**Fixed Effects model** (within)
```{r DefSpreadFixed, echo = TRUE}
# 
# NATOcnvResultsFixed <- list(plm(DefSpendDelt_lead ~ NATO.EUspread_lag2 +  IntlCnf + GDPpCapDelt + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATO.EUspread_lag2, data = VariablesInModel,  model = "within"))
# summary(NATOcnvResultsFixed)
# 
# screenreg(list(EUDefSpreadResults_EUspread, NATOcnvResultsFixed))
# 
# #testing for fixed effects
# # pFtest(EUDefSpreadResultsA, EUDefSpreadResults8)
# 
# ##p = .1358,, cannot reject null hypothesis that there are not significnat effects, use pooled 

```

**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r DefSperadHausmanTest, echo = TRUE}
# 
# NATO.EUspreadResultsWithin <- list(plm(DefSpendDelt_lead ~ NATO.EUspread_lag2 +  IntlCnf + GDPpCapDelt + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATO.EUspread_lag2, data = VariablesInModel,  index=c("Country", "Year"), model="within"))
# 
# NATO.EUspreadResultsRandom <- list(plm(DefSpendDelt_lead ~ NATO.EUspread_lag2 +  IntlCnf + GDPpCapDelt + Cab_liberty_authority+ eu_anti_pro_ls_spread + eu_anti_pro_ls_spread:NATO.EUspread_lag2, data = VariablesInModel,  index=c("Country", "Year"), model="random"))
# 
# phtest(NATO.EUspreadResultsWithin, NATO.EUspreadResultsRandom)
# ##p .7333 cannot reject null, which makes sense because it told us before to do pooled. 
```

