---
title: "EU Exert Strong Leadership and Operations and Maintenance Spending"
author: "src"
date: "October 20, 2015"
output: html_document
---

#Setup 
##Load Packages and Support Files
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")

```


##Load the Data
```{r Input, echo = FALSE, results = "hide"}


#debug(CompilePubOpDataOmnibus)
##load data using function
#undebug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()



write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <-subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )




```


##Subset the Data
```{r EULeadershipNarrowing, echo = TRUE, results = "hide"}

##load data using function
EULeadershipSpread <- subset(EuropeOmnibus, !is.na(OnMspendDelt_lead))


#Subset to countries with more complete time series (excludes Greece and Serbia)
EuropeOmnibus <-subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))

EULeadershipSpread_gap1 <- subset(EULeadershipSpread, !is.na(EUldrSpread))
EULeadershipSpread_gap2 <- subset(EULeadershipSpread, !is.na(EUldrSpread_lag1))
EULeadershipSpread_gap3 <- subset(EULeadershipSpread, !is.na(EUldrSpread_lag2))


```

#Building the Regression Model
##Choosing the Public Opinion Lag
```{r EULeadershipspreadRegressions, echo = TRUE}
OnMEULeadershipResults1 <- lm(OnMspendDelt_lead ~ EUldrSpread, EULeadershipSpread_gap1) 
summary(OnMEULeadershipResults1)
screenreg(list(OnMEULeadershipResults1))

OnMEULeadershipResults2 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag1, EULeadershipSpread_gap2) 
summary(OnMEULeadershipResults2)
screenreg(list(OnMEULeadershipResults1, OnMEULeadershipResults2),
              custom.model.names=c("PubOp_0","PubOp_1"))

##lag 1 is not significant

OnMEULeadershipResults3 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults3)
screenreg(list(OnMEULeadershipResults1, OnMEULeadershipResults3),
              custom.model.names=c("PubOp_0","PubOp_2"))

##lag2 is not significant
EULeadershipSpread_gap2_pdata <- plm.data(EULeadershipSpread_gap2, index=c("Country", "Year"))


##lag2 has the lowest p value and will be included
```

##Security Environment Controls
``` {r SecurityEnvironment, echo = TRUE}
OnMEULeadershipResults4 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2 + IntAt, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults4)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults4),
              custom.model.names=c("PubOp_1","IntAt"))

##intAt isn't significant and won't be included

OnMEULeadershipResults5 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2 + CivilWar, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults5)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5),
              custom.model.names=c("PubOp_1","CivilWar"))

##CivilWar isn't significant and won't be included

OnMEULeadershipResults6 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2  + IntlCnf, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults6)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults6),
              custom.model.names=c("PubOp_1","CivilWar", "IntlCnf"))

##IntlCnf is not significant and Won't be included


```

##Macroeconomic Controls


``` {r MacroVariables, echo = TRUE}

OnMEULeadershipResults7 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2  + GDPpCapLCUdelt, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults7)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults7),
              custom.model.names=c("PubOp_1","CivilWar", "GDP"))

##GDPpCapLCUdelt isnt significant and will not  be included.
```

##Parliamentary Opinion Controls


``` {r Government, echo = TRUE}


OnMEULeadershipResults8 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2  + Cab_left_right, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults8)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults8),
              custom.model.names=c("PubOp_1","CivilWar", "cableftright"))

##cableft right is not significant and will not be included

OnMEULeadershipResults9 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2  + Cab_liberty_authority, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults9)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults9),
              custom.model.names=c("PubOp_1","CivilWar", "cablibertyauthority"))

##cab_liberty_authority is not significant and will not be included

OnMEULeadershipResults2 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2  +  Cab_eu_anti_pro, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults2)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults9, OnMEULeadershipResults2),
              custom.model.names=c("PubOp_1","CivilWar", "cablibertyauthority", "cabeuantipro"))

##cab_eu_anti_pro is significant and will be included.

OnMEULeadershipResults11 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2   + left_right_ls_spread, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults11)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults9, OnMEULeadershipResults2, OnMEULeadershipResults11),
              custom.model.names=c("PubOp_1","CivilWar", "cablibertyauthority", "cabeuantipro", "leftrightls"))

##leftrightls is not significant and won't be included

OnMEULeadershipResults12 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2   + liberty_authority_ls_spread, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults12)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults9, OnMEULeadershipResults2, OnMEULeadershipResults12),
              custom.model.names=c("PubOp_1","CivilWar", "cablibertyauthority", "cabeuantipro", "libauthLS"))

##libauth ls isn't significant and wont' be included

OnMEULeadershipResults13 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2   + eu_anti_pro_ls_spread, EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults13)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults9, OnMEULeadershipResults2, OnMEULeadershipResults13),
              custom.model.names=c("PubOp_1","CivilWar", "cablibertyauthority", "cabeuantipro", "euantiproLS"))

##euantipro isn't significant and wont' be included.

summary(OnMEULeadershipResults2)

```


**Model 2 will be used.**


#Determining Panel Data Model

``` {r defspreadPoolingTest, echo = TRUE}

EULeadershipResults2 <- lm(OnMspendDelt_lead ~ EUldrSpread_lag2  , EULeadershipSpread_gap3) 
summary(OnMEULeadershipResults2)
screenreg(list(OnMEULeadershipResults2, OnMEULeadershipResults5, OnMEULeadershipResults9, OnMEULeadershipResults2),
              custom.model.names=c("PubOp_1","CivilWar", "cablibertyauthority", "cabeuantipro"))


OnMspreadPooled <- plm(OnMspendDelt_lead ~ EUldrSpread_lag2  , data = EULeadershipSpread_gap2_pdata, model = "pooling")

screenreg(list(EULeadershipResults2, OnMspreadPooled),
          custom.model.names = c("OLS", "Pooled"))

summary(OnMspreadPooled)

OnMspreadBetween <- plm(OnMspendDelt_lead ~ EUldrSpread_lag2  , data = EULeadershipSpread_gap2_pdata, model = "between")
summary(OnMspreadBetween)
screenreg(list(EULeadershipResults2, OnMspreadPooled, OnMspreadBetween),
          custom.model.names = c("OLS", "Pooled", "Between"))

OnMspreadFD <- plm(OnMspendDelt_lead ~ EUldrSpread_lag2  , data = EULeadershipSpread_gap2_pdata, model = "fd") ##when variables don't vary from one year to the next, the variable is dropped because the first difference takes difference of the model averaged over time from the original model and for these variables the value would be 0.
screenreg(list(EULeadershipResults2, OnMspreadPooled, OnMspreadBetween, OnMspreadFD),
          custom.model.names = c("OLS", "Pooled", "Between", "FD"))


OnMspreadWithin <- plm(OnMspendDelt_lead ~ EUldrSpread_lag2  , data =  EULeadershipSpread_gap2_pdata, model = "within")
summary(OnMspreadWithin)
screenreg(list(EULeadershipResults2, OnMspreadPooled, OnMspreadBetween, OnMspreadFD, OnMspreadWithin),
          custom.model.names = c("OLS", "Pooled", "Between", "FD", "within"))


# OnMspreadRandom <- plm(OnMspendDelt_lead ~ EUldrSpread_lag2  , data = EULeadershipSpread_gap2_pdata, model = "random")
# screenreg(list(EULeadershipResults2, OnMspreadPooled, OnMspreadBetween, OnMspreadFD, OnMspreadWithin, DefSpreadRandom),
#           custom.model.names = c("OLS", "Pooled", "Between", "FD", "within", "Random"))
# 

##lagrange multiplier (lm) for random effects vs. OLS
plmtest(OnMspreadPooled) #if pvalue small, go ahead and estimate a random effects model. "alternative hypothesis, significant effects"

##fixed effects versus OLS model
pFtest(OnMspreadWithin, OnMspreadPooled)
##p = .168 cannot reject the null that the pooling model works (all coefficients for each country should be the same ever year) if p value low, supports the within model(accept alternative hypothesis: significant effects)

##hausman test for fixed versus random effects model:compares how close the coefficients of the model are:
# phtest(OnMspreadRandom, OnMspreadWithin)
##small p value means one model is inconsistent, means use fixed effects model b/c that gives you consistent estimates.
```

- plmtest tells use to use the pooled ols model (p = .341)

- pftest tells us to use pooled because the p value = .669 cannot reject null that the pooling model is best.

**Use pooled.**
