---
title: "European Defense and Public Opinion: Selected Models"
author: "Greg Sanders"
date: "October 19, 2015"
output: 
html_document: 
keep_md: yes
toc: yes
---


```{r Setup, echo = FALSE, results = "hide", warning=FALSE, message=FALSE}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")
```


##Load the Data
```{r Input, echo = FALSE, results = "hide", warning=FALSE, message=FALSE}

EuropeOmnibus <- CompilePubOpDataOmnibus()
EuropeOmnibus <-subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom","Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))

EuropeOmnibus <-subset(EuropeOmnibus, select=-c(Unit.Currency))


panelmodels<-function(selected.formula,
                      source.data,
                      regression.name,
                      dependent.name=NA,
                      include.random=TRUE){
    
    model.df<-data.frame(name=regression.name,
                         dependent.name=dependent.name,
                         ols=I(list(lm(as.formula(selected.formula), 
                                       data = source.data))),
                         pooling=I(list(plm(as.formula(selected.formula), 
                                            data = source.data, model = "pooling"))),
                         between=I(list(plm(as.formula(selected.formula), 
                                            data = source.data, model = "between"))),
                         fd=I(list(plm(as.formula(selected.formula), 
                                       data = source.data, model = "fd"))),
                         within=I(list(plm(as.formula(selected.formula), 
                                           data = source.data, model = "within"))),
                         random=I(ifelse(include.random,
                                         list(plm(as.formula(selected.formula), 
                                                  data = source.data, model = "random")),
                                         NA))
                         
    )
    
    model.df
    
}

```




##Subset the Data
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function

EuropeOmnibus<-plm.data(EuropeOmnibus, index=c("Country", "Year"))

EquSpread <- subset(EuropeOmnibus, !is.na(EquSpendDelt_lead))
DefSpread <- subset(EuropeOmnibus, !is.na(DefSpendDelt_lead))

Europe_model<-list()
EuropeDef_name<-c()
Europe_model<-list()
EuropeEqu_name<-c()

```


#Hypothesis 1: Public Support for Defense Spending

Hypothesis 1 a: Net public support for more defense spending results in an increase of defense spending.
Hypothesis 1 b: Net public support for more defense spending results in an increase of investment spending.

``` {r Hyp1_Too_Much_Little_Defense, echo = TRUE}
summary(is.na(DefSpread$DefSpread_lag1))

#Model #1 is Defense Spending and Defense Spread
Europe_model<-panelmodels(selected.formula="DefSpendDelt_lead ~ DefSpread_lag1 + GDPpCapUSDdelt",
                          source.data=subset(DefSpread,!is.na(DefSpread_lag1)),
                          regression.name="Too Much/Little & Change in Total Defense",
                          include.random=FALSE)

# summary(Europe_model$ols[[1]])
summary(Europe_model$pooling[[1]])
# summary(Europe_model$between[[1]])
# summary(Europe_model$fd[[1]])
# summary(Europe_model$within[[1]])
# summary(Europe_model$random[[1]])
# plot(Europe_model$ols[[1]])
#190, #86, #194,81


#Mode2 #1 is Equipment Spending and Defense Spread
Europe_model<-rbind(
    Europe_model[1,],
    panelmodels(selected.formula="EquSpendDelt ~ DefSpread_lag2 + GDPpCapUSDdelt + left_right_ls_spread",
                source.data=subset(EquSpread,!is.na(DefSpread_lag2)),
                regression.name="Too Much/Little & Change in Def. Equipment")
)

# summary(Europe_model$ols[[2]])
# summary(Europe_model$pooling[[2]])
# summary(Europe_model$between[[2]])
# summary(Europe_model$fd[[2]])
# summary(Europe_model$within[[2]])
summary(Europe_model$random[[2]])
# plot(Europe_model$ols[[2]])
htmlreg(list(Europe_model$pooling[[1]],Europe_model$random[[2]]),
        file="Output\\Hypothesis1.html",
        custom.model.name=c(as.character(Europe_model$name[1:2])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
        custom.coef.names = c(NA, #(Intercept)
                              "Too Much/Little Def. Spending (year T-2)",#DefSpread_lag1
                              "Change in GDP per Capita (year T-1)", #GDPpCapUSDdelt
                              "Too Much/Little Def. Spending (year T-3)",#DefSpread_lag2
                              "Left-Right Polarization (year T-1)"),#left_right_ls_spread 
        reorder.coef=c(2,4,3,5,1),
        groups = list("Polling" = 1:2, "MacroEconomics" = 3,"Parliamentary" = 4,"Intercept" = 5),
        custom.note="%stars. Note: The model of total defense spending used random effects. The defense equipment used pooled regression."
)



```

#Hypothesis 2: Public Support for Active European Foreign Policy

Hypothesis 2a: Net public support of individual European countries for a greater presence of the EU in international affairs results in an increase of investment defense spending.
Hypothesis 2b: Net public support of individual European countries for a greater presence of the EU in international affairs results in an increase of investment spending.


``` {r Hyp2_EU_Leadership, echo = TRUE}

#Model #3 is Defense Spending and European Leadership
Europe_model<-rbind(
    Europe_model[1:2,],
    
    panelmodels(selected.formula="DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapUSDdelt +
                                 eu_anti_pro_ls_spread",
                source.data=subset(DefSpread,!is.na(EUldrSpread)),
                regression.name="EU leadership & Change in Total Defense",
                dependent.name="Change in Defense",
                include.random=TRUE)
)

# summary(Europe_model$ols[[3]])
summary(Europe_model$pooling[[3]])
# summary(Europe_model$between[[3]])
# summary(Europe_model$fd[[3]])
# summary(Europe_model$within[[3]])
# summary(Europe_model$random[[3]])
# plot(Europe_model$ols[[3]])
Europe_model<-rbind(
    Europe_model[1:3,],
    panelmodels(selected.formula="EquSpendDelt ~ EUldrSpread_lag1 + CivilWar + Cab_liberty_authority + Cab_eu_anti_pro",
                source.data=subset(EquSpread,!is.na(EUldrSpread_lag1)),
                regression.name="EU leadership & Change in Def. Equipment",
                dependent.name = "Change in Def. Equip.",
                include.random=FALSE)
)


#Mode2 #4 is Equipment Spending and European Leadership    
# summary(Europe_model$ols[[4]])
summary(Europe_model$pooling[[4]])
# summary(Europe_model$between[[4]])
# summary(Europe_model$fd[[4]])
# summary(Europe_model$within[[4]])
# summary(Europe_model$random[[4]])
# plot(Europe_model$ols[[4]])





htmlreg(list(Europe_model$pooling[[3]],Europe_model$pooling[[4]]),
        file="Output\\Hypothesis2.html",
        custom.model.name=c(as.character(Europe_model$name[3:4])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
        custom.coef.names = c(NA,  # (Intercept)  
                              "EU Leadership (year T-1)", #EUldrSpread   
                              "Militarized Interstate Disputes (year T-1)", # IntlCnf 
                              "Change in GDP per Capita (year T-1)", #GDPpCapUSDdelt
                              "EU-Anti-Pro Polarization (year T-1)", #eu_anti_pro_ls_spread 
                              "EU Leadership (year T-2)", #EUldrSpread_lag1
                              "Civil War (year T-21 to T-1)", #CivilWar 
                              "Cabinet Liberty-Authority (year T-1)", #Cab_liberty_authority
                              "Cabinet EU-Anti-Pro (year T-1)"), # Cab_eu_anti_pro   
        reorder.coef=c(2,6,3,7,4,8,9,5,1),
        groups = list("Polling Net Support for" = 1:2, "Security"=3:4, "Macroeconomic" = 5, "Parliamentary"=6:8,"Constant" = 9),
        custom.note="%stars. Note: Both models used pooled regression.",
        bold=0.05
)

```

#Hypothesis 3: Public Support for NATO

Hypothesis 3 a: Net public support for believing that NATO is essential to your country's security will increase defense spending. 
Hypothesis 3 b: Net public support for believing that NATO is essential to your country's security will increase defense spending. 
Hypothesis 3 cb: Net public support for believing that NATO is essential to your country's security will increase investment spending.

``` {r Hyp3_NATO, echo = TRUE}


#Model #5 is Defense Spending and NATO EU convergence
Europe_model<-rbind(
    Europe_model[1:4,],
    panelmodels(selected.formula="DefSpendDelt_lead ~ NATO.EUspread_lag1   + GDPpCapUSDdelt + liberty_authority_ls_spread",
                source.data=subset(DefSpread,!is.na(NATO.EUspread_lag1)&
                                                 (!Country %in% c(   "Turkey"))),
                regression.name="U.S.-EU Partnership & Change in Total Defense",
                include.random=FALSE)
)

# summary(Europe_model$ols[[5]])
summary(Europe_model$pooling[[5]])
# summary(Europe_model$between[[5]])
# summary(Europe_model$fd[[5]])
# summary(Europe_model$within[[5]])
# summary(Europe_model$random[[5]])
# plot(Europe_model$ols[[5]])

#Model #6 is Defense Spending and NATO essential
Europe_model<-rbind(
    Europe_model[1:5,],
    panelmodels(selected.formula="DefSpendDelt_lead ~ NATOessSpread_lag2",
                source.data=subset(EquSpread,!is.na(NATOessSpread_lag2)),
                regression.name="NATO Essential & Change in Total Defense",
                include.random=TRUE)
)

# summary(Europe_model$ols[[6]])
summary(Europe_model$pooling[[6]])
# summary(Europe_model$between[[6]])
# summary(Europe_model$fd[[6]])
# summary(Europe_model$within[[6]])
# summary(Europe_model$random[[6]])

# plot(Europe_model$ols[[6]])


#Model #7 is Equipment Spending and EU NATO  convergence
Europe_model<-rbind(
    Europe_model[1:6,],
    panelmodels(selected.formula="EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapUSDdelt + eu_anti_pro_ls_spread",
                source.data=subset(EquSpread,!is.na(NATO.EUspread_lag2)&
                                                 (!Country %in% c(   "Turkey"))),
                regression.name="U.S.-EU Partnership & Change in Def. Equipment",
                include.random=FALSE)
)


# summary(Europe_model$ols[[7]])
summary(Europe_model$pooling[[7]])
# summary(Europe_model$between[[7]])
# summary(Europe_model$fd[[7]])
# summary(Europe_model$within[[7]])
# summary(Europe_model$random[[7]])
# plot(Europe_model$ols[[7]])



#Model #8 is Equipment Spending and NATO essential
Europe_model<-rbind(
    Europe_model[1:7,],
    panelmodels(selected.formula="EquSpendDelt_lead  ~ NATOessSpread_lag1 + CivilWar" ,
                source.data=subset(EquSpread,!is.na(NATOessSpread_lag1)),
                regression.name="NATO Essential & Change in Def. Equipment",
                include.random=FALSE)
)


# summary(Europe_model$ols[[8]])
summary(Europe_model$pooling[[8]])
# summary(Europe_model$between[[8]])
# summary(Europe_model$fd[[8]])
# summary(Europe_model$within[[8]])
# summary(Europe_model$random[[8]])
# plot(Europe_model$ols[[8]])



htmlreg(list(Europe_model$pooling[[5]],Europe_model$within[[6]],Europe_model$pooling[[7]],Europe_model$pooling[[8]]),
        file="Output\\Hypothesis3.html",
        custom.model.name=c(as.character(Europe_model$name[5:8])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
        custom.coef.names = c(NA,#(Intercept)
                              "U.S.-EU Partnership (Year T-2)",#NATO.EUspread_lag1
                              "Change in GDP per Capita (year T-1)",#GDPpCapUSDdelt
                              "Liberty-Authority Polarization (year T-1)",#liberty_authority_ls_spread
                              "NATO as Essential to Security (Year T-3)",#NATOessSpread_lag2
                              "U.S.-EU Partnership (Year T-3)",#NATO.EUspread_lag2
                              "EU-Anti-Pro Polarization (Year T-1)",#eu_anti_pro_ls_spread
                              "NATO as Essential to Security (Year T-2)",#NATOessSpread_lag1
                              "Civil War (year T-21 to T-1)"),#CivilWar),
        reorder.coef=c(2,6,8,5,9,3,4,7,1),
        groups = list("Polling Net Support for" = 1:4, "Security"=5,  "Macroeconomic" =6, "Parliamentary"=7:8,"Intercept" = 9),
        custom.note="%stars. Note: All four models used pooled regression.",
        bold=0.05
)


```

#Dependent Variable cross-analyses

##All Tests Using Top-Line Defense Spending as a Dependent Variable

```{r DefenseDependent}
screenreg(list(Europe_model$pooling[[1]],Europe_model$pooling[[3]],Europe_model$pooling[[5]],Europe_model$within[[6]]),
          custom.model.name=c(as.character(Europe_model$name[c(1,3,5,6)])),
          digits=4,
          stars=c(0.001,0.01,0.05,0.1),
          reorder.coef=c(1,7,2,4,9,5,3,8,6),
          groups = list("Intercept" = 1,"Polling" = 2:5, "Security"=6,"Macroeconomic" = 7, "Parliamentary"=8:9)
)


```

##All Tests Using Defense Investment Spending as a Dependent Variable

```{r EquipmentDependent}
screenreg(list(Europe_model$random[[2]],Europe_model$pooling[[4]],Europe_model$pooling[[7]],Europe_model$pooling[[8]]),
          custom.model.name=c(as.character(Europe_model$name[c(2,4,7,8)])),
          digits=4,
          stars=c(0.001,0.01,0.05,0.1),
          reorder.coef=c(1,2,5,9,11,6,3,4,7,10,8),
          groups = list("Intercept" = 1,"Polling" = 2:5, "Security"=6,"Macroeconomic" = 7, "Parliamentary"=8:11)
)

```