---
title: "European Defense and Public Opinion: Selected Models"
author: "Greg Sanders"
date: "October 19, 2015"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---


```{r Setup, echo = FALSE, results = "hide", warning=FALSE, message=FALSE}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")
```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide", warning=FALSE, message=FALSE}

EuropeOmnibus <- CompilePubOpDataOmnibus()
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))




panelmodels<-function(selected.formula,
                      source.data,
                      regression.name,
                      include.random=TRUE){
    
    model.df<-data.frame(name=regression.name,
        ols=I(list(lm(as.formula(selected.formula), 
                              data = source.data))),
        pooling=I(list(plm(as.formula(selected.formula), 
                              data = source.data, model = "pooling"))),
                         between=I(list(plm(as.formula(selected.formula), 
                              data = source.data, model = "between"))),
                         fd=I(list(plm(as.formula(selected.formula), 
                              data = source.data, model = "fd"))),
                         within=I(list(plm(as.formula(selected.formula), 
                              data = source.data, model = "within"))),
                         random=I(ifelse(include.random,
                                       list(plm(as.formula(selected.formula), 
                                                  data = source.data, model = "random")),
                                       NA))
                                       
    )
    
    model.df
    
}

```




**load and subset the data we want**
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EquSpread <- subset(EuropeOmnibus, !is.na(EquSpendDelt_lead))
DefSpread <- subset(EuropeOmnibus, !is.na(DefSpendDelt_lead))

Europe_model<-list()
EuropeDef_name<-c()
Europe_model<-list()
EuropeEqu_name<-c()

```


#Hypothesis 1: Public Support for Defense Spending

Hypothesis 1 a: Net public support for more defense spending results in an increase of defense spending.
Hypothesis 1 b: Net public support for more defense spending results in an increase of investment spending.

``` {r Hyp1_Too_Much_Little_Defense, echo = TRUE}

Europe_model<-panelmodels(selected.formula="DefSpendDelt_lead ~ DefSpread_lag1 + GDPpCapDelt",
                              source.data=DefSpread,
                              regression.name="Too Much/Little & Def",
                             include.random=FALSE)

# summary(Europe_model$ols[[1]])
summary(Europe_model$pooling[[1]])
# summary(Europe_model$between[[1]])
# summary(Europe_model$fd[[1]])
# summary(Europe_model$within[[1]])
# summary(Europe_model$random[[1]])

Europe_model<-rbind(
    Europe_model[1,],
    panelmodels(selected.formula="EquSpendDelt ~ DefSpread_lag2 + GDPpCapDelt + left_right_ls_spread",
                              source.data=DefSpread,
                              regression.name="Too Much/Little & Equ")
)
    
# summary(Europe_model$ols[[2]])
# summary(Europe_model$pooling[[2]])
# summary(Europe_model$between[[2]])
# summary(Europe_model$fd[[2]])
# summary(Europe_model$within[[2]])
summary(Europe_model$random[[2]])

screenreg(list(Europe_model$pooling[[1]],Europe_model$random[[2]]),
          custom.model.name=c(as.character(Europe_model$name[1:2])),
          digits=3,
          stars=c(0.01,0.05,0.1),
          reorder.coef=c(1,2,4,3,5),
          groups = list("Intercept" = 1,"Polling" = 2:3, "MacroEconomics" = 4,"Parliamentary" = 5)
          )



```

#Hypothesis 2: Public Support for Active European Foreign Policy

Hypothesis 2a: Net public support of individual European countries for a greater presence of the EU in international affairs results in an increase of investment defense spending.
Hypothesis 2b: Net public support of individual European countries for a greater presence of the EU in international affairs results in an increase of investment spending.


``` {r Hyp2_EU_Leadership, echo = TRUE}

Europe_model<-rbind(
    Europe_model[1:2,],
    panelmodels(selected.formula="DefSpendDelt ~ EUldrSpread + IntlCnf + GDPpCapDelt +
                                 eu_anti_pro_ls_spread",
                source.data=DefSpread,
                regression.name="EU leader. & Def",
                include.random=TRUE)
)

# summary(Europe_model$ols[[3]])
summary(Europe_model$pooling[[3]])
# summary(Europe_model$between[[3]])
# summary(Europe_model$fd[[3]])
# summary(Europe_model$within[[3]])
# summary(Europe_model$random[[3]])

Europe_model<-rbind(
    Europe_model[1:3,],
    panelmodels(selected.formula="EquSpendDelt ~ EUldrSpread_lag1 + CivilWar + Cab_liberty_authority + Cab_eu_anti_pro",
                source.data=DefSpread,
                regression.name="EU leader. & Equ",
                include.random=FALSE)
)
    
    
# summary(Europe_model$ols[[4]])
summary(Europe_model$pooling[[4]])
# summary(Europe_model$between[[4]])
# summary(Europe_model$fd[[4]])
# summary(Europe_model$within[[4]])
# summary(Europe_model$random[[4]])

screenreg(list(Europe_model$pooling[[3]],Europe_model$pooling[[4]]),
          custom.model.name=c(as.character(Europe_model$name[3:4])),
          digits=3,
          stars=c(0.01,0.05,0.1),
          reorder.coef=c(1,2,6,3,7,4,8,9,5),
          groups = list("Intercept" = 1,"Polling" = 2:3, "Security"=4:5, "Macroeconomic" = 6, "Parliamentary"=7:9)
          )

```

#Hypothesis 3: Public Support for NATO

Hypothesis 3 a: Net public support for believing that NATO is essential to your country's security will increase defense spending. 
Hypothesis 3 b: Net public support for believing that NATO is essential to your country's security will increase defense spending. 
Hypothesis 3 cb: Net public support for believing that NATO is essential to your country's security will increase investment spending.

``` {r Hyp3A, echo = TRUE}

Europe_model<-rbind(
    Europe_model[1:4,],
    panelmodels(selected.formula="DefSpendDelt_lead ~ NATOessSpread_lag2",
                source.data=DefSpread,
                regression.name="NATO-EU Conv. & Def",
                include.random=TRUE)
)

# summary(Europe_model$ols[[5]])
# summary(Europe_model$pooling[[5]])
# summary(Europe_model$between[[5]])
summary(Europe_model$fd[[5]])
# summary(Europe_model$within[[5]])
# summary(Europe_model$random[[5]])


Europe_model<-rbind(
    Europe_model[1:5,],
    panelmodels(selected.formula="DefSpendDelt_lead ~ NATOessSpread_lag2",
                source.data=DefSpread,
                regression.name="NATO Ess. & Def",
                include.random=FALSE)
)

# summary(Europe_model$ols[[6]])
# summary(Europe_model$pooling[[6]])
# summary(Europe_model$between[[6]])
summary(Europe_model$fd[[6]])
# summary(Europe_model$within[[6]])
# summary(Europe_model$random[[6]])



Europe_model<-rbind(
    Europe_model[1:6,],
    panelmodels(selected.formula="EquSpendDelt_lead ~ NATO.EUspread_lag2  + GDPpCapDelt + eu_anti_pro_ls_spread",
                source.data=DefSpread,
                regression.name="NATO-EU Conv. & Equ",
                include.random=FALSE)
)
    
    
# summary(Europe_model$ols[[7]])
# summary(Europe_model$pooling[[7]])
# summary(Europe_model$between[[7]])
summary(Europe_model$fd[[7]])
# summary(Europe_model$within[[7]])
# summary(Europe_model$random[[3]])


Europe_model<-rbind(
    Europe_model[1:7,],
    panelmodels(selected.formula="EquSpendDelt_lead ~ NATOessSpread_lag1  + CivilWar",
                source.data=DefSpread,
                regression.name="NATO Ess. & Equ",
                include.random=FALSE)
)

# summary(Europe_model$ols[[8]])
# summary(Europe_model$pooling[[8]])
# summary(Europe_model$between[[8]])
summary(Europe_model$fd[[8]])
# summary(Europe_model$within[[8]])
# summary(Europe_model$random[[8]])


screenreg(list(Europe_model$fd[[5]],Europe_model$fd[[6]],Europe_model$fd[[7]],Europe_model$fd[[8]]),
          custom.model.name=c(as.character(Europe_model$name[5:8])),
          digits=4,
          stars=c(0.01,0.05,0.1),
          reorder.coef=c(1,2,3,6,4,5),
          groups = list("Intercept" = 1,"Polling" = 2:4, "Macroeconomic" = 5, "Parliamentary"=6)

          )



```

#Dependent Variable cross-analyses

##All Tests Using Top-Line Defense Spending as a Dependent Variable

```{r DefenseDependent}
screenreg(list(Europe_model$pooling[[1]],Europe_model$pooling[[3]],Europe_model$fd[[5]],Europe_model$fd[[6]]),
          custom.model.name=c(as.character(Europe_model$name[c(1,3,5,6)])),
          digits=4,
          stars=c(0.01,0.05,0.1),
          reorder.coef=c(1,7,2,4,8,5,3,6),
          groups = list("Intercept" = 1:2,"Polling" = 3:5, "Security"=6,"Macroeconomic" = 7, "Parliamentary"=8)
          )


```

##All Tests Using Defense Investment Spending as a Dependent Variable

```{r EquipmentDependent}
screenreg(list(Europe_model$random[[2]],Europe_model$pooling[[4]],Europe_model$fd[[7]],Europe_model$fd[[8]]),
          custom.model.name=c(as.character(Europe_model$name[c(2,4,7,8)])),
          digits=4,
          stars=c(0.01,0.05,0.1),
          reorder.coef=c(1,9,2,5,10,12,6,3,4,7,11,8),
          groups = list("Intercept" = 1:2,"Polling" = 3:6, "Security"=7,"Macroeconomic" = 8, "Parliamentary"=9:12)
          )

```