---
title: "Europe Regression"
author: "src"
date: "Wednesday, April 08, 2015"
output:
  html_document:
    keep_md: yes
---




**Load packages and support files**
```{r Setup, echo = TRUE }
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)

# setwd("C:/Users/scohen/My Documents/Europe/Europe/Europe/") #Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# setwd("C:/Users/scohen/My Documents/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = TRUE}
debug(CompilePubOpDataOmnibus)
##load data using function
EuropeOmnibus <- CompilePubOpDataOmnibus()
debug(StandardizeCountries)
write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))

##get rird of 2002 and 2003
#EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002", "2003"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


##lead

# shift<-function(x,shift_by){
#     stopifnot(is.numeric(shift_by))
#     stopifnot(is.numeric(x))
#  
#     if (length(shift_by)>1)
#         return(sapply(shift_by,shift, x=x))
#  
#     out<-NULL
#     abs_shift_by=abs(shift_by)
#     if (shift_by > 0 )
#         out<-c(tail(x,-abs_shift_by),rep(NA,abs_shift_by))
#     else if (shift_by < 0 )
#         out<-c(rep(NA,abs_shift_by), head(x,-abs_shift_by))
#     else
#         out<-x
#     out
# }
# 
# EuropeOmnibus$defSpnd_lead <- shift(EuropeOmnibus$DefSpnd, 1)

##create data frame to test for correlation between variables
EuropeOmnibus<-data.frame(
  Dspend  = EuropeOmnibus$DefSpnd,
  DefSpnd_lead  = EuropeOmnibus$DefSpnd_lead,
    = EuropeOmnibus$ThreatRatio,
  IntAt = EuropeOmnibus$IntAt,
  DomAt = EuropeOmnibus$DomAt,
  CivilWar = EuropeOmnibus$CivilWar,
  IntlCnf=  EuropeOmnibus$IntlCnf,
  Population =  EuropeOmnibus$Populationulation,
  GDPpCap = EuropeOmnibus$GDPpCapap,
  Dem = EuropeOmnibus$democ,
  NATO = EuropeOmnibus$NATOally,
  USldrSpread =EuropeOmnibus$USldrSpread,
  DefSpread =EuropeOmnibus$DefSpread,
  Country = EuropeOmnibus$Country,
  Year = EuropeOmnibus$Year
  
)


```

**create data frame to test correlation between variables and test them**
```{r USldrSpread, Correlation, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EuropeOmnibus,select=-c(Country,Year))))

```

**Summary statistics of data**
```{r USldrSpreadSummary, echo = TRUE}

summary(EuropeOmnibus)

```

###US leadership 

**OLS models, adding each variable individually**
```{r USldrSpreadBuilding, echo = TRUE}
USleadResults1 <- lm(log(DefSpnd_lead) ~ USldrSpread, EuropeOmnibus) 
USleadResults2 <- lm(log(DefSpnd_lead) ~ USldrSpread, EuropeOmnibus)
USleadResults3 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt, EuropeOmnibus)
USleadResults4 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt, EuropeOmnibus)
USleadResults5 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar, EuropeOmnibus)
USleadResults6 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf, EuropeOmnibus)
USleadResults7 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population), EuropeOmnibus)
USleadResults8 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EuropeOmnibus)
USleadResults9 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem, EuropeOmnibus)
# USleadResults10 <- lm(log(DefSpnd_lead) ~
#     USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
screenreg(list(USleadResults1, USleadResults2, USleadResults3, USleadResults4, USleadResults5, USleadResults6, USleadResults7, USleadResults8, USleadResults9))

USleadResultChosen<-USleadResults9

plot(USleadResultChosen)

```


**Hypothesis testing**

2. F-tests
``` {r USleadSpreadFtest, echo = TRUE}
##For IntAt and GDPpCap
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults1 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeOmnibus)
var.test(USleadResultChosen, Hresults1, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivilWar and DomAt
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults2 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
var.test(USleadResultChosen, Hresults2, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivilWar and IntlCnf
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults3 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
var.test(USleadResultChosen, Hresults3, ratio = .000001, conf.level = 0.95)
##Cannot reject the null hypothesis that both of these coefficients are equal to zero.  
#because of this, IntlCnf might not merit inclusion in the model becasue it's statistics show that by itself, it's not statistically significant at the 95% level, it only increases r^2 by .01 when added and it doesn't affect the other variables.  (except a decrease in   by .05 but this is not a huge impact on the model)
#IntlCnf will also be tested with Dem so that f-test will determine it's merit in the model.
##Civil war is already significant with other variables.

##For CivilWar and Dem
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults4 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeOmnibus)
var.test(USleadResultChosen, Hresults4, ratio = .000001, conf.level = 0.95)
##Reject the null hypothesis that both of these coefficients are equal to zero

##For IntlCnf and Dem
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults5 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeOmnibus)
var.test(USleadResultChosen, Hresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that  both of these coefficients are equal to zero
##This means that the above uncertainty of IntlCnf's significance is resolved and IntlCnf should stay in the model

##For GDPpCap and CivilWar
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults6 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + IntlCnf + log(Population) + Dem , EuropeOmnibus)
var.test(USleadResultChosen, Hresults6, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##For Dem and USldrSpread
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults7 <- lm(log(DefSpnd_lead) ~   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) , EuropeOmnibus)
var.test(USleadResultChosen, Hresults7, ratio = .000001, conf.level = .95)
## reject the null hypothesis that both of these coefficients are equal to zero

##For CivilWar and USldrSpread
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults8 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
var.test(USleadResultChosen, Hresults8, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##for USldrSpread and Dem
USleadResultChosen <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
Hresults9 <- lm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeOmnibus)
var.test(USleadResultChosen, Hresults9, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero


```


**Fixed effects model**
```{r USleaderFixedEffects, echo = TRUE}

Bresults1 <- plm(log(DefSpnd_lead) ~ USldrSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , data=EuropeOmnibus, index=c("Country", "Year"), model="within")
summary(Bresults1)
#plot(Bresults1)

#Comparing OLS to fixed effects model and testing for fixed effects
screenreg(list(USleadResultChosen, Bresults1))

pFtest(Bresults1, USleadResultChosen)


```


###Country spending too much or too little on defense data --> defense spending

**load and subset the data we want**
```{r DefSpreadNarrowing, echo = TRUE}

##load data using function
EuropeDefSpread <- subset(EuropeOmnibus, !is.na(DefSpread))

#EuropeDefSpread <- subset(EuropeDefSpread,!Country %in% c("EU 10","EU 7","USA","EU 11","EU 9","EU 12","EU 8"))

##need to make this for years greater than 2004
#EuropeDefSpread <- subset(EuropeDefSpread, !Year %in% c("2002", "2003"))

##get rid of sweden
EuropeDefSpread <- subset(EuropeDefSpread, !Country %in% c("Sweden", "Serbia", "Bulgaria", "Romania"))


```

**summary statistics of data**
```{r DefSpreadSummary, echo = TRUE}

##Summary statistics of data
summary(EuropeDefSpread)

```

**create data frame to test correlation between variables and test them**
```{r DefSpreadCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EuropeDefSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r DefSpreadOLS, echo = TRUE}
EUDefSpreadResults1 <- lm(log(DefSpnd_lead) ~ DefSpread, EuropeDefSpread) 

EUDefSpreadResults3 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt, EuropeDefSpread)

EUDefSpreadResults4 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, EuropeDefSpread)

EUDefSpreadResults5 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, EuropeDefSpread)

EUDefSpreadResults6 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, EuropeDefSpread)

EUDefSpreadResults7 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), EuropeDefSpread)

EUDefSpreadResults8 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EuropeDefSpread)

EUDefSpreadResults9 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EuropeDefSpread)

#EUDefSpreadResults10 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)

screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults4, EUDefSpreadResults5, EUDefSpreadResults6, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults9))

plot(EUDefSpreadResults10)

```


**Hypothesis testing**
``` {r DefSpreadTesting, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r DefSpreadFixed, echo = TRUE}
# Non- Conformable Arguments
EUDefSpreadResultsA <- plm(log(DefSpnd_lead) ~ DefSpread +  IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), data=subset(EuropeDefSpread,Year>2004&Year<2013), index=c("Country", "Year"), model="within") #Took out ThreatR
summary(EUDefSpreadResultsA)

screenreg(list(EUDefSpreadResults9, EUDefSpreadResultsA))

pFtest(EUDefSpreadResultsA, EUDefSpreadResults9)

```




###Country spending too much or too little on defense data --> Investment spending

**load and subset the data we want**
```{r DefSpreadNarrowing, echo = TRUE}

##load data using function
EuropeDefSpread <- subset(EuropeOmnibus, !is.na(DefSpread))

#EuropeDefSpread <- subset(EuropeDefSpread,!Country %in% c("EU 10","EU 7","USA","EU 11","EU 9","EU 12","EU 8"))

##need to make this for years greater than 2004
#EuropeDefSpread <- subset(EuropeDefSpread, !Year %in% c("2002", "2003"))

##get rid of sweden
EuropeDefSpread <- subset(EuropeDefSpread, !Country %in% c("Sweden", "Serbia", "Bulgaria", "Romania"))


```

**summary statistics of data**
```{r DefSpreadSummary, echo = TRUE}

##Summary statistics of data
summary(EuropeDefSpread)

```

**create data frame to test correlation between variables and test them**
```{r DefSpreadCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EuropeDefSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r DefSpreadOLS, echo = TRUE}
EUDefSpreadResults1 <- lm(log(EquSpnd_lead) ~ DefSpread, EuropeDefSpread) 

EUDefSpreadResults3 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt, EuropeDefSpread)

EUDefSpreadResults4 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt, EuropeDefSpread)

EUDefSpreadResults5 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, EuropeDefSpread)

EUDefSpreadResults6 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, EuropeDefSpread)

EUDefSpreadResults7 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), EuropeDefSpread)

EUDefSpreadResults8 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EuropeDefSpread)

EUDefSpreadResults9 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EuropeDefSpread)

EUDefSpreadResults10 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EuropeDefSpread)

screenreg(list(EUDefSpreadResults1, EUDefSpreadResults2, EUDefSpreadResults3, EUDefSpreadResults4, EUDefSpreadResults5, EUDefSpreadResults6, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults9, EUDefSpreadResults10))

plot(EUDefSpreadResults10)

```


**Hypothesis testing**
``` {r DefSpreadTesting, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r DefSpreadFixed, echo = TRUE}
# Non- Conformable Arguments
EUDefSpreadResultsA <- plm(log(EquSpnd_lead) ~ DefSpread +  IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), data=subset(EuropeDefSpread,Year>2004&Year<2013), index=c("Country", "Year"), model="within") 
summary(EUDefSpreadResultsA)

screenreg(list(EUDefSpreadResults10, EUDefSpreadResultsA))

pFtest(EUDefSpreadResultsA, EUDefSpreadResults10)

```





###How desirable is it that the EU exert strong leadership in world affairs? Defense Spending

**load and subset the data we want**
```{r EUexertLeadershipNarrowing, echo = TRUE}

##load data using function
EULeadershipSpread <- subset(EuropeOmnibus, !is.na(EUldrSpread))


##get rid of Serbia
EULeadershipSpread <- subset(EULeadershipspread, !Country %in% c("Serbia"))



```

**summary statistics of data**
```{r EULeadershipSpreadSummary, echo = TRUE}

##Summary statistics of data
summary(EULeadershipSpread)

```

**create data frame to test correlation between variables and test them**
```{r EULeadershipSpreadCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EULeadershipSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r EULeadershipspreadRegressions, echo = TRUE}
EULeadershipResults1 <- lm(log(DefSpnd_lead) ~ DefSpread, EULeadershipSpread) 

EULeadershipResults2 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, EULeadershipSpread)

EULeadershipResults3 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt, EULeadershipSpread)

EULeadershipResults4 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, EULeadershipSpread)

EULeadershipResults5 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, EULeadershipSpread)

EULeadershipResults6 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, EuropeDefSpread)

EULeadershipResults7 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), EULeadershipSpread)

EULeadershipResults8 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EULeadershipSpread)

EULeadershipResults9 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem, EULeadershipSpread)

##edit government variable
EULeadershipResults10 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EULeadershipSpread)

screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults3, EULeadershipResults4, EULeadershipResults5, EULeadershipResults6, EULeadershipResults7, EULeadershipResults8, EULeadershipResults9, EULeadershipResults10))

plot(EULeadershipResults10)

```


**Hypothesis testing**
``` {r EULeadershipspreadTesting, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r EULeadershipspreadFixed, echo = TRUE}

EULeadershipResultsA <- plm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EULeadershipSpread, index=c("Country", "Year"), model="within") 
summary(DefSpreadresultsA)

screenreg(list(EULeadershipResults10, EULeadershipResultsA))

pFtest(EULeadershipResultsA, EULeadershipResults10)

```




###How desirable is it that the EU exert strong leadership in world affairs? Investment SPending

**load and subset the data we want**
```{r EUexertLeadershipNarrowing, echo = TRUE}

##load data using function
EULeadershipSpread <- subset(EuropeOmnibus, !is.na(EUldrSpread))


##get rid of Serbia
EULeadershipSpread <- subset(EULeadershipspread, !Country %in% c("Serbia"))



```

**summary statistics of data**
```{r EULeadershipSpreadSummary, echo = TRUE}

##Summary statistics of data
summary(EULeadershipSpread)

```

**create data frame to test correlation between variables and test them**
```{r EULeadershipSpreadCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EULeadershipSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r EULeadershipspreadRegressions, echo = TRUE}
EULeadershipResults1 <- lm(log(EquSpnd_lead) ~ DefSpread, EULeadershipSpread) 

EULeadershipResults2 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt, EULeadershipSpread)

EULeadershipResults3 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt, EULeadershipSpread)

EULeadershipResults4 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt, EULeadershipSpread)

EULeadershipResults5 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, EULeadershipSpread)

EULeadershipResults6 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, EuropeDefSpread)

EULeadershipResults7 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), EULeadershipSpread)

EULeadershipResults8 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EULeadershipSpread)

EULeadershipResults9 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EULeadershipSpread)

##edit government variable
EULeadershipResults10 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EULeadershipSpread)

screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults3, EULeadershipResults4, EULeadershipResults5, EULeadershipResults6, EULeadershipResults7, EULeadershipResults8, EULeadershipResults9, EULeadershipResults10))

plot(EULeadershipResults10)

```


**Hypothesis testing**
``` {r EULeadershipspreadTesting, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r EULeadershipspreadFixed, echo = TRUE}

EULeadershipResultsA <- plm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EULeadershipSpread, index=c("Country", "Year"), model="within") 
summary(DefSpreadresultsA)

screenreg(list(EULeadershipResults10, EULeadershipResultsA))

pFtest(EULeadershipResultsA, EULeadershipResults10)

```




###Is NATO still essential to your country's secuirty? and Defense Spending

**load and subset the data we want**
```{r NATOEssentialNarrowing, echo = TRUE}

##load data using function
NATOEssentialSpread <- subset(EuropeOmnibus, !is.na(NATOessSpread))


##get rid of 2014
NATOEssentialSpread <- subset(NATOEssentialSpread, !Year %in% c("2014"))



```

**summary statistics of data**
```{r NATOEssentialSummary, echo = TRUE}

##Summary statistics of data
summary(NATOEssentialSpread)

```

**create data frame to test correlation between variables and test them**
```{r NATOEssentialCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(NATOEssentialSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r NATOEssentialRegressions, echo = TRUE}
NATOEssentialResults1 <- lm(log(DefSpnd_lead) ~ DefSpread, NATOEssentialSpread) 

NATOEssentialResults2 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, NATOEssentialSpread)

NATOEssentialResults3 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt, NATOEssentialSpread)

NATOEssentialResults4 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, NATOEssentialSpread)

NATOEssentialResults5 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, NATOEssentialSpread)

NATOEssentialResults6 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, NATOEssentialSpread)

NATOEssentialResults7 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), NATOEssentialSpread)

NATOEssentialResults8 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread)

NATOEssentialResults9 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread)

##edit government variable
NATOEssentialResults10 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread)

screenreg(list(NATOEssentialResults1, NATOEssentialResults2, NATOEssentialResults3, NATOEssentialResults4, NATOEssentialResults5, NATOEssentialResults6, NATOEssentialResults7, NATOEssentialResults8, NATOEssentialResults9, NATOEssentialResults10))

plot(NATOEssentialResults10)

```


**Hypothesis testing**
``` {r NATOEssentialTesting, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r NATOEssentialFixed, echo = TRUE}

NATOEssentialResultsA <- plm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread, index=c("Country", "Year"), model="within") 
summary(DefSpreadresultsA)

screenreg(list(NATOEssentialResults10, NATOEssentialResultssA))

pFtest(NATOEssentialResultsA, NATOEssentialResults10)

```




###Is NATO still essential to your country's secuirty? and Investment Spending

**load and subset the data we want**
```{r NATOEssentialNarrowing_EquSpnd_lead, echo = TRUE}

##load data using function
NATOEssentialSpread <- subset(EuropeOmnibus, !is.na(NATOessSpread))


##get rid of 2014
NATOEssentialSpread <- subset(NATOEssentialSpread, !Year %in% c("2014"))



```

**summary statistics of data**
```{r NATOEssentialSummary_EquSpnd_lead, echo = TRUE}

##Summary statistics of data
summary(NATOEssentialSpread)

```

**create data frame to test correlation between variables and test them**
```{r NATOEssentialCorr_EquSpnd_lead, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(NATOEssentialSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r NATOEssentialRegressions_EquSpnd_lead, echo = TRUE}
NATOEssentialResults1 <- lm(log(EquSpnd_lead) ~ DefSpread, NATOEssentialSpread) 

NATOEssentialResults2 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt, NATOEssentialSpread)

NATOEssentialResults3 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt, NATOEssentialSpread)

NATOEssentialResults4 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt, NATOEssentialSpread)

NATOEssentialResults5 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, NATOEssentialSpread)

NATOEssentialResults6 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, NATOEssentialSpread)

NATOEssentialResults7 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), NATOEssentialSpread)

NATOEssentialResults8 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread)

NATOEssentialResults9 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread)

##edit government variable
NATOEssentialResults10 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread)

screenreg(list(NATOEssentialResults1, NATOEssentialResults2, NATOEssentialResults3, NATOEssentialResults4, NATOEssentialResults5, NATOEssentialResults6, NATOEssentialResults7, NATOEssentialResults8, NATOEssentialResults9, NATOEssentialResults10))

plot(NATOEssentialResults10)

```


**Hypothesis testing**
``` {r NATOEssentialTesting_EquSpnd_lead, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r NATOEssentialFixedEquSpnd_lead, echo = TRUE}

NATOEssentialResultsA <- plm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), NATOEssentialSpread, index=c("Country", "Year"), model="within") 
summary(DefSpreadresultsA)

screenreg(list(NATOEssentialResults10, NATOEssentialResultssA))

pFtest(NATOEssentialResultsA, NATOEssentialResults10)

```




##EU and NATO should be closer or more independent?


**load and subset the data we want**
```{r EUnNATOCloseNarrowing, echo = TRUE}

##load data using function
EUnNATOCloseSpread <- subset(EuropeOmnibus, !is.na(NATO.EUspread))


##get rid of Serbia and Turkey
EUnNATOCloseSpread <- subset(EUnNATOCloseSpread, !Country %in% c("Serbia", "Turkey"))

#get rid of 2014
EUnNATOCloseSpread <- subset(EUnNATOCloseSpread, !Year %in% c("2014"))



```

**summary statistics of data**
```{r EUnNATOCloseSummary, echo = TRUE}

##Summary statistics of data
summary(EUnNATOCloseSpread)

```

**create data frame to test correlation between variables and test them**
```{r EUnNATOCloseCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EUnNATOCloseSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r EUnNATOCloseRegressions, echo = TRUE}
EUnNATOCloseSpreadResultsResults1 <- lm(log(DefSpnd_lead) ~ DefSpread, EUnNATOCloseSpread) 

EUnNATOCloseSpreadResultsResults2 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, EUnNATOCloseSpread)

EUnNATOCloseSpreadResultsResults3 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt, EUnNATOCloseSpread)

EUnNATOCloseSpreadResults4 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, EUnNATOCloseSpread)

EUnNATOCloseSpreadResults5 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, EUnNATOCloseSpread)

EUnNATOCloseSpreadResults6 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, EUnNATOCloseSpread)

EUnNATOCloseSpreadResults7 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), EUnNATOCloseSpread)

EUnNATOCloseSpreadResults8 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUnNATOCloseSpread)

EUnNATOCloseSpreadResults9 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUnNATOCloseSpread)

##edit government variable
EUnNATOCloseSpreadResults10 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUnNATOCloseSpread)

screenreg(list(EUnNATOCloseSpreadResults1, EUnNATOCloseSpreadResults2, EUnNATOCloseSpreadResults3, EUnNATOCloseSpreadResults4, EUnNATOCloseSpreadResults5, EUnNATOCloseSpreadResults6, EUnNATOCloseSpreadResults7, EUnNATOCloseSpreadResults8, EUnNATOCloseSpreadResults9, EUnNATOCloseSpread1Results10))

plot(EUnNATOCloseSpreadResults10)

```


**Hypothesis testing**
``` {r EUnNATOCloseTesting, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r EUnNATOCloseFixed, echo = TRUE}

EUnNATOCloseSpreadResultsA <- plm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUnNATOCloseSpread, index=c("Country", "Year"), model="within") 
summary(EUnNATOCloseSpreadResultsA)

screenreg(list(EUnNATOCloseSpreadResults10, EUnNATOCloseSpreadResultssA))

pFtest(EUnNATOCloseSpreadResults10, EUnNATOCloseSpreadResultsA)

```




### EU Favorability


**load and subset the data we want**
```{r EUfvrSpreadNarrowing, echo = TRUE}

##load data using function
EUfvrSpread <- subset(EuropeOmnibus, !is.na(EUfvrSpread))


##get rid of Serbia 
EUfvrSpread <- subset(EUfvrSpread, !Country %in% c("Serbia"))

#get rid of 2014
EUfvrSpread <- subset(EUfvrSpread, !Year %in% c("2014"))



```

**summary statistics of data**
```{r EUfvrSpreadSpreadSummary, echo = TRUE}

##Summary statistics of data
summary(EUfvrSpread)

```

**create data frame to test correlation between variables and test them**
```{r EUfvrSpreadSpreadCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EUfvrSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r EUfvrSpreadRegressions, echo = TRUE}
EUfvrSpreadResults1 <- lm(log(DefSpnd_lead) ~ DefSpread, EUfvrSpread) 

EUfvrSpreadResults2 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, EUfvrSpread)

EUfvrSpreadResults3 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt, EUfvrSpread)

EUfvrSpreadResults4 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt, EUfvrSpread)

EUfvrSpreadResults5 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar, EUfvrSpread)

EUfvrSpreadResults6 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf, EUfvrSpread)

EUfvrSpreadResults7 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population), EUfvrSpread)

EUfvrSpreadResults8 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUfvrSpread)

EUfvrSpreadResults9 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUfvrSpread)

##edit government variable
EUfvrSpreadResults10 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUfvrSpread)

screenreg(list(EUfvrSpreadResults1, EUfvrSpreadResults2, EUfvrSpreadResults3, EUfvrSpreadResults4, EUfvrSpreadResults5, EUfvrSpreadResults6, EUfvrSpreadResults7, EUfvrSpreadResults8, EUfvrSpreadResults9, EUfvrSpreadResults10))

plot(EUfvrSpreadResultsResults10)

```


**Hypothesis testing**
``` {r EUfvrSpreadTesting, echo = TRUE}

##  and GDPpCap
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults1 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivilWar
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults2 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and IntlCnf
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults3 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivilWar and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults4 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + IntlCnf + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults5 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + log(Population) + log(GDPpCap) , EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Population and Dem
Cresults10 <- lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
Gresults6 <-  lm(log(DefSpnd_lead) ~ DefSpread +   + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap) + Dem , EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r EUfvrSpreadFixed, echo = TRUE}

EUfvrSpreadResultsResultsA <- plm(log(DefSpnd_lead) ~ DefSpread + IntAt + DomAt + CivilWar + IntlCnf + log(Population) + log(GDPpCap), EUfvrSpreadResults, index=c("Country", "Year"), model="within") 
summary(EUfvrSpreadResultsA)

screenreg(list(EUfvrSpreadResults10, EUfvrSpreadResultsA))

pFtest(EUfvrSpreadResultsA, EUfvrSpreadResults10)

```

