---
title: "European Defense and Public Opinion: Selected Models"
author: "Greg Sanders"
date: "October 19, 2015"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---


```{r Setup, echo = FALSE, results = "hide", warning=FALSE, message=FALSE}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(reshape2)
require(plyr)
require(quantmod)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# #setwd("C:/Users/scohen/My Documents/Europe/Git/Europe")

source("EuropeInput.R")
```


##Load the Data
```{r Input, echo = FALSE, results = "hide", warning=FALSE, message=FALSE}



EuropeOmnibus <- CompilePubOpDataOmnibus(path="Data//")
# EuropeOmnibus <-subset(EuropeOmnibus, select=-c(Euro))
EuropeOmnibus<-plm.data(EuropeOmnibus, index=c("Country", "Year"))

panelmodels<-function(selected.formula,
                      source.data,
                      regression.name,
                      dependent.name=NA,
                      independent.name=NA,
                      include.random=TRUE){
    
    model.df<-data.frame(name=regression.name,
                         dependent.name=dependent.name,
                         independent.name=independent.name,
                         ols=I(list(lm(as.formula(selected.formula), 
                                       data = source.data))),
                         fixed=I(list(plm(as.formula(selected.formula),
                                          data = source.data,
                                          model = "within"))),
                         fixed.time=I(list(plm(as.formula(paste(selected.formula,"+ factor(Year)")), 
                                               data =  source.data,
                                               model = "within"))),
                         random=I(ifelse(include.random,
                                         list(plm(as.formula(selected.formula), 
                                                  data =  source.data,
                                                  model = "random")),
                                         NA))
                         
    )
    
    model.df
}
```




##Subset the Data
```{r DatasetNarrowing, echo = FALSE, results = "hide"}
EuropeOmnibus <-subset(EuropeOmnibus, Country %in% c("France","Germany","United Kingdom",
                                                     "Italy","Netherlands","Poland",
                                                     "Portugal","Slovakia","Spain","Turkey"))


#Limit data to those cases where both the dependent and independent variables are present
EuropeMoreDef <- subset(EuropeOmnibus, !is.na(DefSpend_lead) & !is.na(EquSpend_lead) & 
                            !is.na(OnMspend_lead) & !is.na(DefSpread_lag1))
EuropeEUleadership <- subset(EuropeOmnibus, !is.na(DefSpend_lead) &
                                 !is.na(EquSpend_lead) & !is.na(OnMspend_lead) &
                                 !is.na(EUldrSpread_lag1) & Country != "Turkey")
EuropeNATOess <- subset(EuropeOmnibus,  !is.na(DefSpend_lead) &
                                 !is.na(EquSpend_lead) & !is.na(OnMspend_lead) &
                                !is.na(NATOessSpread_lag1))
EuropeEUcloserUS <- subset(EuropeOmnibus, !is.na(DefSpend_lead) &
                                 !is.na(EquSpend_lead) & !is.na(OnMspend_lead) &
                                 !is.na(EUcloserUSspread_lag1) & Country != "Turkey")
EuropeOmnibus<- subset(EuropeOmnibus, !is.na(DefSpend_lead) & (!is.na(DefSpread_lag1)|
                                                               !is.na(EUldrSpread_lag1)|
                                                                   !is.na(NATOessSpread_lag1)|
                                                                   !is.na(EUcloserUSspread_lag1)))



EuropeModel<-list()

Controls<-"+ log(NGDP_eu2014)+
                            EUdebt_NGDP+Dfc_NGDP+
                             +IntAt+log(GCivilWarBRD)+
                             Cab_liberty_authority+liberty_authority_ls_spread+
                            Cab_eu_anti_pro"


ControlsNoTimeEffect<-"+ log(NGDP_eu2014)+
                            Dfc_NGDP+
                             +IntAt+UNmilitaryPMil+
                             Cab_liberty_authority+liberty_authority_ls_spread+
                            Cab_eu_anti_pro"

```
Debt_NGDP+log(Population)+log(RussiaDefSpend)+log(GCivilWarBRD)+UNmilitaryPMil+Dfc_NGDP

#Hypothesis 1: Public Support for Defense Spending

Hypothesis 1 a: Net public support for more defense spending results in an increase of defense spending.
Hypothesis 1 b: Net public support for more defense spending results in an increase of investment spending.
Hypothesis 1 b: Net public support for more defense spending results in an increase of operations and maintenance spending.

``` {r Hyp1_Too_Much_Little_Defense, echo = TRUE}


#Model 1: Total Defense and More/Less Defense Spending 
EuropeModel<-panelmodels(selected.formula=paste("log(DefSpend_lead) ~ DefSpread_lag1",Controls),
                          source.data=EuropeMoreDef,
                          regression.name="Too Much/Little Def. Spending & Total Defense",
                          dependent.name="Total Defense",
                          independent.name="Defense More/Less",
                          include.random=FALSE)


summary(EuropeModel$fixed[[1]])
# plot(EuropeModel$ols[[1]])
#190, #86, #194,81


#Model #2 is Equipment Spending and Defense Spread
EuropeModel<-rbind(
    EuropeModel[1,],
    EuropeModel<-panelmodels(selected.formula=paste("log(EquSpend_lead) ~ DefSpread_lag1",Controls),
                          source.data=EuropeMoreDef,
                          regression.name="Too Much/Little Def. Spending & Def. Equipment",
                          dependent.name="Defense Equipment",
                          independent.name="Defense More/Less",
                          include.random=FALSE)
)

summary(EuropeModel$fixed[[2]])
# plot(EuropeModel$ols[[1]])
#190, #86, #194,81



#Model #3 is Equipment Spending and Defense Spread
EuropeModel<-rbind(
    EuropeModel[1:2,],
    EuropeModel<-panelmodels(selected.formula=paste("log(OnMspend_lead) ~ DefSpread_lag1",Controls),
                          source.data=EuropeMoreDef,
                          regression.name="Too Much/Little Def. Spending & Def. O&M/Other",
                          dependent.name="Defense O&M/Other",
                          independent.name="Defense More/Less",
                          include.random=FALSE)
)

summary(EuropeModel$fixed[[3]])
# plot(EuropeModel$ols[[1]])
#190, #86, #194,81




htmlreg(list(EuropeModel$fixed[[1]],EuropeModel$fixed[[2]],EuropeModel$fixed[[3]]),
        file="Output\\Hypothesis1.html",
        custom.model.name=c(as.character(EuropeModel$name[1:3])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
        custom.coef.names = c("More Def. Spending",#Dlog(NGDPPC_eu2014)
                              "Log(GDP), (2014 Euros)", #log(NGDP_eu2014)
                              "EU27-wide Debt, (Percent of EU 27 GDP)",# EUdebt_NGDP
                              "Deficit, (Percent of GDP)",# Dfc_NGDP
                              "International Terrorism, (Number of Incidents)", #IntAt
                              "Log(Global Civil War Battle-Related Deaths)", # log(GCivilWarBRD)
                              "Cabinet Liberty/Authority, (GAL=0, TAN=10)", #Cab_liberty_authority
                              "Liberty/Authority Polarization", #liberty_authority_ls_spread
                              "Cabinet Anti/Pro EU, (Anti=0, Pro=10)"), #Cab_liberty_authority)
        groups = list("Public Net Support for (year T-2)" = 1,"Macroeconomics (year T-1)" = 2:4,"Security (year T-1)" = 5:6,"Parliamentary (year T-1)" = 7:9),
        custom.note="%stars.",
            bold=0.05
)

# "Debt as a percent of GDP",# Debt_NGDP
#                               "UN Mission supporting troops and experts, per million",# UNmilitaryPMil


```

#Hypothesis 2: Public Support for Active European Foreign Policy

Hypothesis 2a: Net public support of individual European countries for a greater presence of the EU in international affairs results in an increase of investment defense spending.
Hypothesis 2b: Net public support of individual European countries for a greater presence of the EU in international affairs results in an increase of investment spending.
Hypothesis 2c: Net public support of individual European countries for a greater presence of the EU in international affairs results in an increase of operations & maintenance spending.


``` {r Hyp2_EU_Leadership, echo = TRUE}
#Model 4: Total Defense and EU Leadership
EuropeModel<-rbind(
    EuropeModel[1:3,],
    panelmodels(selected.formula=paste("log(DefSpend_lead) ~ EUldrSpread_lag1",Controls),
                source.data=EuropeEUleadership,
                regression.name="Desire for EU Leadership & Total Defense",
                dependent.name="Total Defense",
                independent.name="EU Leadership",
                include.random=FALSE)
)


summary(EuropeModel$fixed[[4]])
# plot(EuropeModel$ols[[4]])
# summary(EuropeModel$ols[[4]])


#Model #5 is Equipment Spending and European Leadership    
EuropeModel<-rbind(
    EuropeModel[1:4,],
    EuropeModel<-panelmodels(selected.formula=paste("log(EquSpend_lead) ~ EUldrSpread_lag1",Controls),
                          source.data=EuropeEUleadership,
                          regression.name="Desire for EU Leadership & Def. Equipment",
                          dependent.name="Defense Equipment",
                          independent.name="EU leadership",
                          include.random=FALSE)
)




#Model #5 is Equipment Spending and European Leadership    
summary(EuropeModel$fixed[[5]])

# plot(EuropeModel$ols[[4]])


#Model #6 is O&M/Other Spending and Defense Spread
EuropeModel<-rbind(
    EuropeModel[1:5,],
    EuropeModel<-panelmodels(selected.formula=paste("log(OnMspend_lead) ~ EUldrSpread_lag1",Controls),
                          source.data=EuropeEUleadership,
                          regression.name="Desire for EU Leadership & Def. O&M/Other",
                          dependent.name="Defense O&M/Other",
                          independent.name="EU leadership",
                          include.random=FALSE)
)

summary(EuropeModel$fixed[[6]])
# plot(EuropeModel$ols[[1]])
#190, #86, #194,81




htmlreg(list(EuropeModel$fixed[[4]],EuropeModel$fixed[[5]],EuropeModel$fixed[[6]]),
        file="Output\\Hypothesis2.html",
        custom.model.name=c(as.character(EuropeModel$name[4:6])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
         custom.coef.names = c("Strong EU Leadership ",#Dlog(NGDPPC_eu2014)
                              "Log(GDP), (2014 Euros)", #log(NGDP_eu2014)
                              "EU27-wide Debt, (Percent of EU27 GDP)",# EUdebt_NGDP
                              "Deficit, (Percent of GDP)",# Dfc_NGDP
                              "International Terrorism, (Number of Incidents)", #IntAt
                              "Log(Global Civil War Battle-Related Deaths)", # log(GCivilWarBRD)
                              "Cabinet Liberty/Authority, (GAL=0, TAN=10)", #Cab_liberty_authority
                              "Liberty/Authority Polarization", #liberty_authority_ls_spread
                              "Cabinet Anti/Pro EU, (Anti=0, Pro=10)"), #Cab_liberty_authority)
        groups = list("Polling Net Desire for (year T-2)" = 1,"Macroeconomics (year T-1)" = 2:4,"Security (year T-1)" = 5:6,"Parliamentary (year T-1)" = 7:9),          custom.note="%stars.",
        bold=0.05
)

```

#Hypothesis 3: Public Support for NATO

Hypothesis 3 a: Net public support for transatlantic partnership will increase defense spending. 
Hypothesis 3 b: Net public support for transatlantic partnership will decrease defense spending. 
Hypothesis 3 c: Net public support for transatlantic partnership will increase investment spending.
Hypothesis 3 d: Net public support for transatlantic partnership will increase operations & maintenance spending.

``` {r Hyp3_Transatlantic, echo = TRUE}
#Model 7: Total Defense and NATO Essential
EuropeModel<-rbind(
    EuropeModel[1:6,],
    panelmodels(selected.formula=paste("log(DefSpend_lead) ~ NATOessSpread_lag1",Controls),
                source.data=EuropeNATOess,
                regression.name="NATO Essential & Total Defense",
                dependent.name="Total Defense",
                independent.name="NATO Essential",
                include.random=FALSE)
)


summary(EuropeModel$fixed[[7]])
# plot(EuropeModel$ols[[7]])


#Model 8: Defense Equipment and NATO Essential
EuropeModel<-rbind(
    EuropeModel[1:7,],
    panelmodels(selected.formula=paste("log(EquSpend_lead) ~ NATOessSpread_lag1",Controls),
                source.data=EuropeNATOess,
                regression.name="NATO Essential & Defense Equipment",
                dependent.name="Defense Equipment",
                independent.name="NATO Essential",
                include.random=TRUE)
)

summary(EuropeModel$fixed[[8]])
# plot(EuropeModel$ols[[8]])

#Model #    9 is Defense O&M/Other and NATO essential
EuropeModel<-rbind(
    EuropeModel[1:8,],
    EuropeModel<-panelmodels(selected.formula=paste("log(OnMspend_lead) ~ NATOessSpread_lag1",Controls),
                          source.data=EuropeNATOess,
                          regression.name="NATO Essential & Def. O&M/Other",
                          dependent.name="Defense O&M/Other",
                          independent.name="NATO Essential",
                          include.random=FALSE)
)

summary(EuropeModel$fixed[[9]])
# plot(EuropeModel$ols[[9]])




#Model 10: Total Defense and EU closer to US
EuropeModel<-rbind(
    EuropeModel[1:9,],
    panelmodels(selected.formula=paste("log(DefSpend_lead) ~ EUcloserUSspread_lag1",
                                       ControlsNoTimeEffect),
                source.data=EuropeEUcloserUS,
                regression.name="EU closer to US & Total Defense",
                dependent.name="Defense O&M/Other",
                independent.name="EU closer to US",
                include.random=FALSE)
)


summary(EuropeModel$fixed.time[[10]])
# plot(EuropeModel$ols[[10]])




#Model 11: Defense Equipment and EU closer to US
EuropeModel<-rbind(
    EuropeModel[1:10,],
    panelmodels(selected.formula=paste("log(EquSpend_lead) ~ EUcloserUSspread_lag1",
                                       ControlsNoTimeEffect),
                source.data=EuropeEUcloserUS,
                regression.name="EU closer to US & Defense Equipment",
                dependent.name="Defense Equipment",
                independent.name="EU closer to US",
                include.random=FALSE)
)

summary(EuropeModel$fixed.time[[11]])
# plot(EuropeModel$ols[[11]])

#Model # 12 Defense O&M/Other and EU closer to US
EuropeModel<-rbind(
    EuropeModel[1:11,],
    EuropeModel<-panelmodels(selected.formula=paste("log(OnMspend_lead) ~ EUcloserUSspread_lag1",
                                                    ControlsNoTimeEffect),
                          source.data=EuropeEUcloserUS,
                          regression.name="EU closer to US & Def. O&M/Other",
                          dependent.name="Defense O&M/Other",
                          independent.name="EU closer to US",
                          include.random=FALSE)
)

summary(EuropeModel$fixed.time[[12]])
# plot(EuropeModel$ols[[12]])

htmlreg(list(EuropeModel$fixed[[7]],EuropeModel$fixed[[8]],EuropeModel$fixed[[9]],
             EuropeModel$fixed.time[[10]],EuropeModel$fixed.time[[11]],EuropeModel$fixed.time[[12]]),
        file="Output\\Hypothesis3.html",
        custom.model.name=c(as.character(EuropeModel$name[7:12])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
        reorder.coef=c(1,10,2,3,4,5,6,11,7,8,9,12:18),
        custom.coef.names = c("NATO as Essential to Security",#Dlog(NGDPPC_eu2014)#   NATOessSpread_lag1
                              "Log(GDP), (2014 Euros)", #log(NGDP_eu2014)
                              "EU27-wide Debt, (Percent of EU27 GDP)",# EUdebt_NGDP
                              "Deficit, (Percent of GDP)",# Dfc_NGDP
                              "International Terrorism, (Number of Incidents)", #IntAt
                              "Log(Global Civil War Battle-Related Deaths)", # log(GCivilWarBRD)
                              "Cabinet Liberty/Authority, (GAL=0, TAN=10)", #Cab_liberty_authority
                              "Liberty/Authority Polarization", #liberty_authority_ls_spread
                              "Cabinet Anti/Pro EU, (Anti=0, Pro=10)", #Cab_liberty_authority)
                              "Closer EU-U.S. Partnership",#Dlog(NGDPPC_eu2014)#EUcloserUSspread_lag1
                              "Deployed UN Troops, (Per Million in Population)", # UNmilitaryPMil
                              "Year=2007", #factor(Year)2006
                              "Year=2008", #factor(Year)2007
                              "Year=2010", # factor(Year)2009
                              "Year=2011", # factor(Year)2010
                              "Year=2012", # factor(Year)2011
                              "Year=2013", # factor(Year)2012
                              "Year=2014"), # factor(Year)2013
        groups = list("Polling Net Support for (year T-2)" = 1:2,"Macroeconomics (year T-1)" = 3:5,"Security (year T-1)" = 6:8,"Parliamentary (year T-1)" = 9:11,"Annual Effects (year T)"= 12:18),
        custom.note="%stars. EU-U.S. Year=2009 is omitted because polling was not available two years prior.",
        bold=0.05
)
#Note that because of leads, the dummy refers to the year+1 for the dependent variable.

htmlreg(list(EuropeModel$fixed[[7]],EuropeModel$fixed[[8]],EuropeModel$fixed[[9]]),
        file="Output\\Hypothesis3NATOess.html",
        custom.model.name=c(as.character(EuropeModel$name[7:9])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
        custom.coef.names = c("NATO as Essential to Security",#Dlog(NGDPPC_eu2014)#   NATOessSpread_lag1
                              "Log(GDP), (2014 Euros)", #log(NGDP_eu2014)
                              "EU27-wide Debt, (Percent of EU27 GDP)",# EUdebt_NGDP
                              "Deficit, (Percent of GDP)",# Dfc_NGDP
                              "International Terrorism, (Number of Incidents)", #IntAt
                              "Log(Global Civil War Battle-Related Deaths)", # log(GCivilWarBRD)
                              "Cabinet Liberty/Authority, (GAL=0, TAN=10)", #Cab_liberty_authority
                              "Liberty/Authority Polarization", #liberty_authority_ls_spread
                              "Cabinet Anti/Pro EU, (Anti=0, Pro=10)"), #Cab_liberty_authority)
        groups = list("Polling Net Support for (year T-2)" = 1,"Macroeconomics (year T-1)" = 2:3,"Security (year T-1)" = 4:5,"Parliamentary (year T-1)" = 7:9),
        custom.note="%stars.",
        bold=0.05
)


htmlreg(list(EuropeModel$fixed.time[[10]],EuropeModel$fixed.time[[11]],EuropeModel$fixed.time[[12]]),
        file="Output\\Hypothesis3EUcloserUS.html",
        custom.model.name=c(as.character(EuropeModel$name[10:12])),
        digits=3,
        stars=c(0.001,0.01,0.05,0.1),
        custom.coef.names = c("Closer EU-U.S. Partnership",#Dlog(NGDPPC_eu2014)#EUcloserUSspread_lag1
                              "Log(GDP), (2014 Euros)", #log(NGDP_eu2014)
                              "Deficit, (Percent of GDP)",# Dfc_NGDP
                              "International Terrorism, (Number of Incidents)", #IntAt
                              "Deployed UN Troops, (Per Million in Population)", # UNmilitaryPMil
                              "Cabinet Liberty/Authority, (GAL=0, TAN=10)", #Cab_liberty_authority
                              "Liberty/Authority Polarization", #liberty_authority_ls_spread
                              "Cabinet Anti/Pro EU, (Anti=0, Pro=10)", #Cab_liberty_authority)
                              "Year=2007", #factor(Year)2006
                              "Year=2008", #factor(Year)2007
                              "Year=2010", # factor(Year)2009
                              "Year=2011", # factor(Year)2010
                              "Year=2012", # factor(Year)2011
                              "Year=2013", # factor(Year)2012
                              "Year=2014"), # factor(Year)2013
        groups = list("Polling Net Support for (year T-2)" = 1,"Macroeconomics (year T-1)" = 2:3,"Security (year T-1)" = 4:5,"Parliamentary (year T-1)" = 6:8,"Annual Effects (year T)"= 9:15),
        custom.note="%stars. EU-U.S. Year=2009 is omitted because polling was not available two years prior.",
        bold=0.05
)
```

#Dependent Variable cross-analyses

##All Tests Using Top-Line Defense Spending as a Dependent Variable

```{r DefenseDependent}
screenreg(list(EuropeModel$fixed[[1]],EuropeModel$fixed[[4]],EuropeModel$fixed[[7]],EuropeModel$fixed.time[[10]]),
          custom.model.name=c(as.character(EuropeModel$independent.name[c(1,4,7,10)])),
          digits=4,
          stars=c(0.001,0.01,0.05,0.1),
          reorder.coef=c(1,10,11,12,2,3,4,5,6,13,7,8,9,14:20),
          groups = list("Polling" = 1:4, "Macroeconomic" = 5:7, "Security" = 8:10, 
                        "Parliamentary"=11:13,"Annual Effects (year T=0)"= 14:20),
          custom.note="%stars. EU-U.S. Year=2009 is omitted because polling was not available two years prior.",
        bold=0.05
)


```

##All Tests Using Defense Equipment Spending as a Dependent Variable

```{r EquipmentDependent}
screenreg(list(EuropeModel$fixed[[2]],EuropeModel$fixed[[5]],EuropeModel$fixed[[8]],EuropeModel$fixed.time[[11]]),
          custom.model.name=c(as.character(EuropeModel$independent.name[c(2,5,8,11)])),
          digits=4,
          stars=c(0.001,0.01,0.05,0.1),
          reorder.coef=c(1,10,11,12,2,3,4,5,6,13,7,8,9,14:20),
          groups = list("Polling" = 1:4, "Macroeconomic" = 5:7, "Security" = 8:10, 
                        "Parliamentary"=11:13,"Annual Effects (year T=0)"= 14:20),
          custom.note="%stars. EU-U.S. Year=2009 is omitted because polling was not available two years prior.",
        bold=0.05
)

```

##All Tests Using Defense Operations, Maintenance, and Other Spending as a Dependent Variable

```{r OnMdependent}
screenreg(list(EuropeModel$fixed[[3]],EuropeModel$fixed[[6]],EuropeModel$fixed[[9]],EuropeModel$fixed.time[[12]]),
          custom.model.name=c(as.character(EuropeModel$independent.name[c(3,6,9,12)])),
          digits=4,
          stars=c(0.001,0.01,0.05,0.1),
          reorder.coef=c(1,10,11,12,2,3,4,5,6,13,7,8,9,14:20),
          groups = list("Polling" = 1:4, "Macroeconomic" = 5:7, "Security" = 8:10, 
                        "Parliamentary"=11:13,"Annual Effects (year T=0)"= 14:20)
)

```