---
title: "NATOEssential and Investment Spending"
author: "src"
date: "October 16, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
setwd("H:/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}
debug(CompilePubOpDataOmnibus)
##load data using function
EuropeOmnibus <- CompilePubOpDataOmnibus()
debug(StandardizeCountries)
write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))

##get rird of 2002 and 2003
#EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002", "2003"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors

EuropeOmnibus$Cab_left_right <- as.character(EuropeOmnibus$Cab_left_right)
EuropeOmnibus$Cab_left_right <- as.numeric(EuropeOmnibus$Cab_left_right)

EuropeOmnibus$Cab_liberty_authority <- as.character(EuropeOmnibus$Cab_liberty_authority)
EuropeOmnibus$Cab_liberty_authority <- as.numeric(EuropeOmnibus$Cab_liberty_authority)

EuropeOmnibus$Cab_eu_anti_pro <- as.character(EuropeOmnibus$Cab_eu_anti_pro)
EuropeOmnibus$Cab_eu_anti_pro <- as.numeric(EuropeOmnibus$Cab_eu_anti_pro)

EuropeOmnibus$left_right_ls_spread <- as.character(EuropeOmnibus$left_right_ls_spread)
EuropeOmnibus$left_right_ls_spread <- as.numeric(EuropeOmnibus$left_right_ls_spread)

EuropeOmnibus$liberty_authority_ls_spread <- as.character(EuropeOmnibus$liberty_authority_ls_spread)
EuropeOmnibus$liberty_authority_ls_spread <- as.numeric(EuropeOmnibus$liberty_authority_ls_spread)

EuropeOmnibus$eu_anti_pro_ls_spread <- as.character(EuropeOmnibus$eu_anti_pro_ls_spread)
EuropeOmnibus$eu_anti_pro_ls_spread <- as.numeric(EuropeOmnibus$eu_anti_pro_ls_spread)

```



###Is NATO still essential to your country's secuirty? and Investment Spending

**load and subset the data we want**
```{r NATOEssentialNarrowing_EquSpnd_lead, echo = TRUE, results = "hide"}

##load data using function
NATOEssentialSpread <- subset(EuropeOmnibus, !is.na(NATOessSpread))


##get rid of 2014
NATOEssentialSpread <- subset(NATOEssentialSpread, !Year %in% c("2014"))


```



**OLS model adding each variable individually**
```{r NATOEssentialRegressions_EquSpnd_lead, echo = TRUE}
InvNATOEssResults1 <- lm(log(EquSpnd_lead) ~ NATOessSpread, NATOEssentialSpread)
summary(InvNATOEssResults1)
screenreg(list(InvNATOEssResults1))

InvNATOEssResults2 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt, NATOEssentialSpread)
summary(InvNATOEssResults2)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2))

##significnat, in creases r^2, will be included

InvNATOEssResults3 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt +DomAt, NATOEssentialSpread)
summary(InvNATOEssResults3)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults3))

##no impact, doesn't increase r^2, will not be included

InvNATOEssResults4 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + CivilWar, NATOEssentialSpread)
summary(InvNATOEssResults4)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults4))

##not significant, doesn't affect other variables greatly, will not be included


InvNATOEssResults5 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + IntlCnf, NATOEssentialSpread)
summary(InvNATOEssResults5)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults5))

##doesn't change adjusted r^2, is not significant, will not be included


InvNATOEssResults6 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population), NATOEssentialSpread)
summary(InvNATOEssResults6)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6))

##significant, increases r^2, will be included


InvNATOEssResults7 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap), NATOEssentialSpread)
summary(InvNATOEssResults7)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7))

##increases significance, is significant, increases r^2, will be included


InvNATOEssResults8 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_left_right, NATOEssentialSpread)
summary(InvNATOEssentialResults8)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults8))

##Cab_left_right is significant, does increase r^2 by .01, will be included

InvNATOEssResults9 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, NATOEssentialSpread)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults8, InvNATOEssResults9))

##Cab_liberty_authority is more significant, will be inlcued in the model and not cab_left_right

InvNATOEssResults10 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + Cab_eu_anti_pro, NATOEssentialSpread)
summary(InvNATOEssResults10)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults9, InvNATOEssResults10))

##cab_eu_anti_pro isn't significant, won't be included in the model

InvNATOEssResults11 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + left_right_ls_spread, NATOEssentialSpread)
summary(InvNATOEssResults11)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults9, InvNATOEssResults11))

##left_right_ls is not significant, wont be included

InvNATOEssResults12 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spread, NATOEssentialSpread)
summary(InvNATOEssResults12)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults9, InvNATOEssResults12))

##liberty_authority_ls_spread is significant and will be included

InvNATOEssResults13 <- lm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spread + eu_anti_pro_ls_spread, NATOEssentialSpread)
summary(InvNATOEssResults13)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults9, InvNATOEssResults12, InvNATOEssResults13))

##eu_anti_pro is not significant, won't be included in the model

```


``` {r InvNATOEssentialInteractions, echo = TRUE}


##center lib_auth_ls_spread

NATOEssentialSpread$liberty_authority_ls_spreadC <- ((NATOEssentialSpread$liberty_authority_ls_spread) - mean(NATOEssentialSpread$liberty_authority_ls_spread))

##center NATOessSpread

NATOEssentialSpread$NATOessSpreadC <- ((NATOEssentialSpread$NATOessSpread) - mean(NATOEssentialSpread$NATOessSpread))

Inter_libauthls_NATOessSpr <- ((NATOEssentialSpread$liberty_authority_ls_spreadC) * (NATOEssentialSpread$NATOessSpreadC))


##include interaction in model

InvNATOEssResults14 <- lm(log(EquSpnd_lead) ~ NATOessSpreadC + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spreadC + Inter_libauthls_NATOessSpr, NATOEssentialSpread)
summary(InvNATOEssResults14)
screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults9, InvNATOEssResults12, InvNATOEssResults14))

##the interaction term is significant, will be included

```

``` {r InvestmentSpendingNATOEssentialpooledTest, echo = TRUE}

InvNATOEssResultsPooled <- plm(log(EquSpnd_lead) ~ NATOessSpreadC + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spreadC + Inter_libauthls_NATOessSpr, data = NATOEssentialSpread, model = "pooling")

InvNATOEssResultsWithin <- plm(log(EquSpnd_lead) ~ NATOessSpreadC + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spreadC + Inter_libauthls_NATOessSpr, data =  NATOEssentialSpread, model = "within")
summary(InvNATOEssResultsWithin)

pFtest(InvNATOEssResultsWithin, InvNATOEssResultsPooled)
##p <2.2e^-16 thus reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```


**Fixed Effects model**
```{r InvestmentSpendingNATOEssentialFixedEquSpnd_lead, echo = TRUE}

InvNATOEssentialResultsA <- plm(log(EquSpnd_lead) ~ NATOessSpreadC  + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spreadC + Inter_libauthls_NATOessSpr, NATOEssentialSpread, index=c("Country", "Year"), model="within") 
summary(InvNATOEssentialResultsA)

screenreg(list(InvNATOEssResults7, InvNATOEssentialResultsA))

pFtest(InvNATOEssentialResultsA, InvNATOEssResults7)

```


**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r InvestmentSpendingNATOEssentialHausmanTest, echo = TRUE}

InvNATOEssResultsWithin <- plm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spreadC + Inter_libauthls_NATOessSpr, data = NATOEssentialSpread, index=c("Country", "Year"), model="within") 

InvNATOEssResultsRandom <- plm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spreadC + Inter_libauthls_NATOessSpr, data = NATOEssentialSpread, index=c("Country", "Year"), model="random")

screenreg(list(InvNATOEssResultsWithin, InvNATOEssResultsRandom))

phtest(InvNATOEssResultsWithin, InvNATOEssResultsRandom)
##p = 1.014e^-11, possibly accept null and use random effects because random effects yields better results..

InvNATOEssResultsWithin1 <- plm(log(EquSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spread + Inter_libauthls_NATOessSpr, data = NATOEssentialSpread, index=c("Country", "Year"), model="within") 

screenreg(list(InvNATOEssResults1, InvNATOEssResults2, InvNATOEssResults6, InvNATOEssResults7, InvNATOEssResults9, InvNATOEssResults12, InvNATOEssResultsWithin1))

##need to discuss the panel data models..not sure what to do. 

```
