---
title: "Country spending too much or too little on defense data and Investment spending"
author: "src"
date: "October 16, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
setwd("H:/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}
debug(CompilePubOpDataOmnibus)
##load data using function
EuropeOmnibus <- CompilePubOpDataOmnibus()
debug(StandardizeCountries)
write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))

##get rird of 2002 and 2003
#EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002", "2003"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors

EuropeOmnibus$Cab_left_right <- as.character(EuropeOmnibus$Cab_left_right)
EuropeOmnibus$Cab_left_right <- as.numeric(EuropeOmnibus$Cab_left_right)

EuropeOmnibus$Cab_liberty_authority <- as.character(EuropeOmnibus$Cab_liberty_authority)
EuropeOmnibus$Cab_liberty_authority <- as.numeric(EuropeOmnibus$Cab_liberty_authority)

EuropeOmnibus$Cab_eu_anti_pro <- as.character(EuropeOmnibus$Cab_eu_anti_pro)
EuropeOmnibus$Cab_eu_anti_pro <- as.numeric(EuropeOmnibus$Cab_eu_anti_pro)

EuropeOmnibus$left_right_ls_spread <- as.character(EuropeOmnibus$left_right_ls_spread)
EuropeOmnibus$left_right_ls_spread <- as.numeric(EuropeOmnibus$left_right_ls_spread)

EuropeOmnibus$liberty_authority_ls_spread <- as.character(EuropeOmnibus$liberty_authority_ls_spread)
EuropeOmnibus$liberty_authority_ls_spread <- as.numeric(EuropeOmnibus$liberty_authority_ls_spread)

EuropeOmnibus$eu_anti_pro_ls_spread <- as.character(EuropeOmnibus$eu_anti_pro_ls_spread)
EuropeOmnibus$eu_anti_pro_ls_spread <- as.numeric(EuropeOmnibus$eu_anti_pro_ls_spread)

```


###Country spending too much or too little on defense data --> Investment spending

**load and subset the data we want**
```{r EquSpreadNarrowing, echo = TRUE, results = "hide"}

##load data using function
EuropeDefSpread <- subset(EuropeOmnibus, !is.na(DefSpread))


##get rid of sweden
EuropeDefSpread <- subset(EuropeDefSpread, !Country %in% c("Sweden", "Serbia", "Bulgaria", "Romania"))


```


**OLS model adding each variable individually**
```{r EquSpreadOLS, echo = TRUE}
EquEUDefSpreadResults1 <- lm(log(EquSpnd_lead) ~ DefSpread, EuropeDefSpread) 
summary(EquEUDefSpreadResults1)

EquEUDefSpreadResults3 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt, EuropeDefSpread)
summary(EquEUDefSpreadResults3)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3))

#IntAt is significant, increases r^2 --> will be inlcuded in model

EquEUDefSpreadResults4 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + DomAt, EuropeDefSpread)
summary(EquEUDefSpreadResults4)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults4))

##DomAt has no effect, is not significant nad doesn't change r^2 -->will not be included

EquEUDefSpreadResults5 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + CivilWar, EuropeDefSpread)
summary(EquEUDefSpreadResults5)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults5))

##CivWar doesn't change other coeficients, increases r^2 by .01, and is not significant thus will not be included in the model. 

EquEUDefSpreadResults6 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + IntlCnf, EuropeDefSpread)
summary(EquEUDefSpreadResults6)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults6))

##IntlCnf does not change anything and is not significant, will not be included

EquEUDefSpreadResults7 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population), EuropeDefSpread)
summary(EquEUDefSpreadResults7)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7))

##log(Population) is significant, increases significance of inAt, changes direction of DefSpread and increases r^2 by .6, it will be included

EquEUDefSpreadResults8 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap), EuropeDefSpread)
summary(EquEUDefSpreadResults8)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8))

##GDPpc is significant, increases r^2 and changes defspread back to positive. will be included.

EquEUDefSpreadResults9 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_left_right, EuropeDefSpread)
summary(EquEUDefSpreadResults9)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8, EquEUDefSpreadResults9))

##cab_left_right is not statistically significant, doesn't change adjusted r^2 and only increases defspread by .01, will not be included.

EquEUDefSpreadResults10 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, EuropeDefSpread)
summary(EquEUDefSpreadResults10)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8, EquEUDefSpreadResults10))

##Cab_liberty_authority is not statistically significant (p = .1014), increases adjusted r^2 by .01, increases defspread by .01. it will not be included

EquEUDefSpreadResults11 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, EuropeDefSpread)
summary(EquEUDefSpreadResults11)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8, EquEUDefSpreadResults11))

##cab_eu_anti_pro is not statistically significant, it does not incerase r^2 and it doesn't affect the other variables' coefficients. it will not be included

EquEUDefSpreadResults12 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + left_right_ls_spread, EuropeDefSpread)
summary(EquEUDefSpreadResults12)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8, EquEUDefSpreadResults12))

##lef_right_ls_spread does not increase r^2 nor is it significant nor does it change the other variables significantly. it will not be included.

EquEUDefSpreadResults13 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + liberty_authority_ls_spread, EuropeDefSpread)
summary(EquEUDefSpreadResults13)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8, EquEUDefSpreadResults13))

##liberty_authority_ls_spread is significant even though it's pvalue is .0464, it increases adjusted r^2 by .01 and decreases log(GDPCap) by .09, increase population by .02 and inAt by .01. it will be included in the model. 

EquEUDefSpreadResults14 <- lm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + liberty_authority_ls_spread + eu_anti_pro_ls_spread, EuropeDefSpread)
summary(EquEUDefSpreadResults14)
screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8, EquEUDefSpreadResults13, EquEUDefSpreadResults14))

##eu_anti_pro_ls_spread isn't significant doesn't increase r^2 or the other variables and decreases significane of liberty_authority_ls_spread, will not be included. 

##model will be 13

```

**interaction with liberty_authority_ls_spread and Defspread**
``` {r EquSpreadInteractiontTest, echo = TRUE}

##center variables

EuropeDefSpread$liberty_authority_ls_spreadC <- ((EuropeDefSpread$liberty_authority_ls_spread) - mean(EuropeDefSpread$liberty_authority_ls_spread))

EuropeDefSpread$DefSpreadC <- ((EuropeDefSpread$DefSpread) - mean(EuropeDefSpread$DefSpread))

EuropeDefSpread$intrGAL_ls_DefSread <- (EuropeDefSpread$DefSpreadC * EuropeDefSpread$liberty_authority_ls_spreadC)

EquEUDefSpreadResults15 <- lm(log(EquSpnd_lead) ~ DefSpreadC + IntAt + log(Population) + log(GDPpCap) + liberty_authority_ls_spreadC + intrGAL_ls_DefSread, EuropeDefSpread)
summary(EquEUDefSpreadResults15)

screenreg(list(EquEUDefSpreadResults1, EquEUDefSpreadResults3, EquEUDefSpreadResults7, EquEUDefSpreadResults8, EquEUDefSpreadResults13, EquEUDefSpreadResults15))

##interaction term is not significant and will not be used. 

```


**Tests of within and pooling** we should explain this in the appendix 

``` {r EquSpreadpooledTest, echo = TRUE}

EquDefSpreadPooled <- plm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + liberty_authority_ls_spread, data = EuropeDefSpread, model = "pooling")

EquDefSpreadWithin <- plm(log(EquSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + liberty_authority_ls_spread, data =  EuropeDefSpread, model = "within")

pFtest(EquDefSpreadWithin, EquDefSpreadPooled)
##p <2.2e^-16 thus reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```


**Fixed Effects model**
```{r EquSpreadFixed, echo = TRUE}

EquEUDefSpreadResultsA <- plm(log(EquSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap) + liberty_authority_ls_spread, data=EuropeDefSpread, index=c("Country", "Year"), model="within") 
summary(EquEUDefSpreadResultsA)

screenreg(list(EquEUDefSpreadResults13, EquEUDefSpreadResultsA))

EquEUDefSpreadResultsB <- plm(log(EquSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap), data=EuropeDefSpread, index=c("Country", "Year"), model="within") 
screenreg(list(EquEUDefSpreadResults8, EquEUDefSpreadResults13, EquEUDefSpreadResultsB, EquEUDefSpreadResultsA))

##should discuss what to do with liberty_authority_ls_spread. if keep it need to add it to next section

```


**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r EquDefSperadHausmanTest, echo = TRUE}

EquEUDefSpreadResultsWithin <- plm(log(EquSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap), data = EuropeDefSpread, index=c("Country", "Year"), model="within") 

EquEUDefSpreadResultsRandom <- plm(log(EquSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap), data = EuropeDefSpread, index=c("Country", "Year"), model="random")

phtest(EquEUDefSpreadResultsWithin, EquEUDefSpreadResultsRandom)
##p - 1.592e^-07, reject null and use fixed effects

```
