---
title: "Nato Essential and Defense Spending"
author: "src"
date: "October 16, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
setwd("H:/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}
debug(CompilePubOpDataOmnibus)
##load data using function
EuropeOmnibus <- CompilePubOpDataOmnibus()
debug(StandardizeCountries)
write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))

##get rird of 2002 and 2003
#EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002", "2003"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors

EuropeOmnibus$Cab_left_right <- as.character(EuropeOmnibus$Cab_left_right)
EuropeOmnibus$Cab_left_right <- as.numeric(EuropeOmnibus$Cab_left_right)

EuropeOmnibus$Cab_liberty_authority <- as.character(EuropeOmnibus$Cab_liberty_authority)
EuropeOmnibus$Cab_liberty_authority <- as.numeric(EuropeOmnibus$Cab_liberty_authority)

EuropeOmnibus$Cab_eu_anti_pro <- as.character(EuropeOmnibus$Cab_eu_anti_pro)
EuropeOmnibus$Cab_eu_anti_pro <- as.numeric(EuropeOmnibus$Cab_eu_anti_pro)

EuropeOmnibus$left_right_ls_spread <- as.character(EuropeOmnibus$left_right_ls_spread)
EuropeOmnibus$left_right_ls_spread <- as.numeric(EuropeOmnibus$left_right_ls_spread)

EuropeOmnibus$liberty_authority_ls_spread <- as.character(EuropeOmnibus$liberty_authority_ls_spread)
EuropeOmnibus$liberty_authority_ls_spread <- as.numeric(EuropeOmnibus$liberty_authority_ls_spread)

EuropeOmnibus$eu_anti_pro_ls_spread <- as.character(EuropeOmnibus$eu_anti_pro_ls_spread)
EuropeOmnibus$eu_anti_pro_ls_spread <- as.numeric(EuropeOmnibus$eu_anti_pro_ls_spread)

```



###Is NATO still essential to your country's secuirty? and Defense Spending

**load and subset the data we want**
```{r NATOEssentialNarrowing, echo = TRUE, results = "hide"}

##load data using function
NATOEssentialSpread <- subset(EuropeOmnibus, !is.na(NATOessSpread))


##get rid of 2014
NATOEssentialSpread <- subset(NATOEssentialSpread, !Year %in% c("2014"))



```



**OLS model adding each variable individually**
```{r NATOEssentialRegressions, echo = TRUE}
NATOEssResults1 <- lm(log(DefSpnd_lead) ~ NATOessSpread, NATOEssentialSpread) 
summary(NATOEssResults1)
screenreg(list(NATOEssResults1))

NATOEssResults2 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt, NATOEssentialSpread)
summary(NATOEssResults2)
screenreg(list(NATOEssResults1, NATOEssResults2))

##intAT increases r^2 and is significant will be included

NATOEssResults3 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + DomAt, NATOEssentialSpread)
summary(NATOEssResults3)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults3))

##DomAt isn't significant, doesn't have an effect and doesn't increase r^2, will not be included

NATOEssResults4 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + CivilWar, NATOEssentialSpread)
summary(NATOEssResults4)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults4))

##doesn't increase r^2, isn't significant and doesn't have an effect on the other variables, will not be included

NATOEssResults5 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + IntlCnf, NATOEssentialSpread)
summary(NATOEssResults5)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults5))

##doesn't increase r^2, isn't significant and doesn't have an effect on the other variables, will not be included

NATOEssResults6 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population), NATOEssentialSpread)
summary(NATOEssResults6)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6))

##Increases r^2, is significant, increases the value and significance on NATOessspread, decreases value and significance of IntAt, will be included


NATOEssResults7 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap), NATOEssentialSpread)
summary(NATOEssResults7)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6, NATOEssResults7))

##increases significance, is significant, increase r^2, will be included


NATOEssResults8 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, NATOEssentialSpread)
summary(NATOEssResults8)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6, NATOEssResults7, NATOEssResults8))

##Cab_liberty_authority is significant, increaser r^2 by one and changes other coefficients, will be inlcuded.

NATOEssResults9 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + Cab_left_right, NATOEssentialSpread)
summary(NATOEssResults9)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6, NATOEssResults7, NATOEssResults8, NATOEssResults9))

##cab_left_right decreases significance of cablibauth, is not significant and deosn't increaser r^2, will not be included 

NATOEssResults10 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + Cab_eu_anti_pro, NATOEssentialSpread)
summary(NATOEssResults10)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6, NATOEssResults7, NATOEssResults8, NATOEssResults10))

##Cab_eu_anti_pro isn't significant and doesn't increaser r^2, will not be included. 

NATOEssResults11 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + left_right_ls_spread, NATOEssentialSpread)
summary(NATOEssResults11)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6, NATOEssResults7, NATOEssResults8, NATOEssResults11))

#left_right_ls is not significant nor does it affect other parts, will not be incldued

NATOEssResults12 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + liberty_authority_ls_spread, NATOEssentialSpread)
summary(NATOEssResults12)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6, NATOEssResults7, NATOEssResults8, NATOEssResults12))

##libertyauthority-Ls isn't significant, doesn't affect adjustd r^2 - will not be included

NATOEssResults13 <- lm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + eu_anti_pro_ls_spread, NATOEssentialSpread)
summary(NATOEssResults13)
screenreg(list(NATOEssResults1, NATOEssResults2, NATOEssResults6, NATOEssResults7, NATOEssResults8, NATOEssResults13))

##eu_anti_prolsspread isn't significant and doesn't affect model, will not be included

```


**Tests of within and pooling** we should explain this in the appendix 

``` {r NATOEssentialpooledTest, echo = TRUE}

NATOEssResultsPooled <- plm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, data = NATOEssentialSpread, model = "pooling")

NATOEssResultsWithin <- plm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, data =  NATOEssentialSpread, model = "within")
summary(NATOEssResultsWithin)

pFtest(NATOEssResultsWithin, NATOEssResultsPooled)
##p <2.2e^-16 thus reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```



**Fixed Effects model**
```{r NATOEssentialFixed, echo = TRUE}

NATOEssResultsA <- plm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, NATOEssentialSpread, index=c("Country", "Year"), model="within") 
summary(NATOEssResultsA)

screenreg(list(NATOEssResults7, NATOEssResultsA))

pFtest(NATOEssResultssA, NATOEssResults7)

```


**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r NATOEssentialHausmanTest, echo = TRUE}

NATOEssResultsWithin <- plm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, data = NATOEssentialSpread, index=c("Country", "Year"), model="within") 

NATOEssResultsRandom <- plm(log(DefSpnd_lead) ~ NATOessSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, data = NATOEssentialSpread, index=c("Country", "Year"), model="random")

screenreg(list(NATOEssResultsWithin, NATOEssResultsRandom))

phtest(NATOEssResultsWithin, NATOEssResultsRandom)
##p - .00436, possibly accept null and use random effects because random effects yields better results..

```

