---
title: "Desirable that EU exert Strong Leadership in World Affairs and Defense Spending"
author: "src"
date: "October 16, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
setwd("H:/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}
debug(CompilePubOpDataOmnibus)
##load data using function
EuropeOmnibus <- CompilePubOpDataOmnibus()
debug(StandardizeCountries)
write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))

##get rird of 2002 and 2003
#EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002", "2003"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors

EuropeOmnibus$Cab_left_right <- as.character(EuropeOmnibus$Cab_left_right)
EuropeOmnibus$Cab_left_right <- as.numeric(EuropeOmnibus$Cab_left_right)

EuropeOmnibus$Cab_liberty_authority <- as.character(EuropeOmnibus$Cab_liberty_authority)
EuropeOmnibus$Cab_liberty_authority <- as.numeric(EuropeOmnibus$Cab_liberty_authority)

EuropeOmnibus$Cab_eu_anti_pro <- as.character(EuropeOmnibus$Cab_eu_anti_pro)
EuropeOmnibus$Cab_eu_anti_pro <- as.numeric(EuropeOmnibus$Cab_eu_anti_pro)

EuropeOmnibus$left_right_ls_spread <- as.character(EuropeOmnibus$left_right_ls_spread)
EuropeOmnibus$left_right_ls_spread <- as.numeric(EuropeOmnibus$left_right_ls_spread)

EuropeOmnibus$liberty_authority_ls_spread <- as.character(EuropeOmnibus$liberty_authority_ls_spread)
EuropeOmnibus$liberty_authority_ls_spread <- as.numeric(EuropeOmnibus$liberty_authority_ls_spread)

EuropeOmnibus$eu_anti_pro_ls_spread <- as.character(EuropeOmnibus$eu_anti_pro_ls_spread)
EuropeOmnibus$eu_anti_pro_ls_spread <- as.numeric(EuropeOmnibus$eu_anti_pro_ls_spread)

```




###How desirable is it that the EU exert strong leadership in world affairs? Defense Spending

**load and subset the data we want**
```{r EULeadershipNarrowing, echo = TRUE, results = "hide"}

##load data using function
EULeadershipSpread <- subset(EuropeOmnibus, !is.na(EUldrSpread))


##get rid of Serbia
EULeadershipSpread <- subset(EULeadershipSpread, !Country %in% c("Serbia"))


```



**OLS model adding each variable individually**
```{r EULeadershipspreadRegressions, echo = TRUE}
EULeadershipResults1 <- lm(log(DefSpnd_lead) ~ EUldrSpread, EULeadershipSpread) 
summary(EULeadershipResults1)
screenreg(list(EULeadershipResults1))

EULeadershipResults2 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt, EULeadershipSpread)
summary(EULeadershipResults2)
screenreg(list(EULeadershipResults1, EULeadershipResults2))

##increases r^2 is signficant will be kept

EULeadershipResults3 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + DomAt, EULeadershipSpread)
summary(EULeadershipResults3)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults3))

#not significnat will not include

EULeadershipResults4 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + CivilWar, EULeadershipSpread)
summary(EULeadershipResults4)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults4))

##not significant or impacting, will not include

EULeadershipResults5 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + IntlCnf, EULeadershipSpread)
summary(EULeadershipResults5)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults5))

##no effect not significant, will not include 

EULeadershipResults6 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population), EULeadershipSpread)
summary(EULeadershipResults6)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6))

##significant, increases r^2 will be included

EULeadershipResults8 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap), EULeadershipSpread)
summary(EULeadershipResults8)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8))

##significant, increases r^2 but, it decreases significance and sign on EUldrSpread, should be discussed.....


EULeadershipResults9 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_left_right, EULeadershipSpread)
summary(EULeadershipResults9)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults9))

##cab_left_right isn't significant and doesn't increase r^2

EULeadershipResults10 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, EULeadershipSpread)
summary(EULeadershipResults10)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults10))

##not significant and doesn't change r^2 significantly (only by .01), will not be included

EULeadershipResults11 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, EULeadershipSpread)
summary(EULeadershipResults11)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults11))

##Cab_eu_anti_pro increases r^2 by .01 and is statistically significant, will be included in the model. 

EULeadershipResults12 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + left_right_ls_spread, EULeadershipSpread)
summary(EULeadershipResults12)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults11, EULeadershipResults12))

##left_right_ls_spread is not significant and will not be included in the model. 

EULeadershipResults13 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread, EULeadershipSpread)
summary(EULeadershipResults13)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults11, EULeadershipResults13))

##liberty_authority_ls_spread is very significant and increases adjusted r^2 by .02. will be included in the model. 

EULeadershipResults14 <- lm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread + eu_anti_pro_ls_spread, EULeadershipSpread)
summary(EULeadershipResults14)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults11, EULeadershipResults13, EULeadershipResults14))

##eu_anti_pro_ls_spread is statistically significant, doens't increase r^2, but will be included in the model. 

```

**interaction terms**

``` {r EUldrspreadInteractionTests, echo = TRUE}

##center variables: liberty_authority_ls_spread + eu_anti_pro_ls_spread

#liberty_authority_ls_spread
EULeadershipSpread$liberty_authority_ls_spreadC <- ((EULeadershipSpread$liberty_authority_ls_spread) - mean(EULeadershipSpread$liberty_authority_ls_spread))

EULeadershipSpread$EUldrSpreadC <- ((EULeadershipSpread$EUldrSpread) - mean(EULeadershipSpread$EUldrSpread))

EULeadershipSpread$Inter_GALTANls_EUldr <- ((EULeadershipSpread$EUldrSpreadC) * (EULeadershipSpread$liberty_authority_ls_spreadC))

##regression with interaction 

EULeadershipResults15 <- lm(log(DefSpnd_lead) ~ EUldrSpreadC + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spreadC + eu_anti_pro_ls_spread + Inter_GALTANls_EUldr, EULeadershipSpread)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults11, EULeadershipResults13, EULeadershipResults14, EULeadershipResults15))

##interaction for liberty_authority_ls_spreadC is not significant thus will not be included. 

##eu_anti_pro_ls_spread
EULeadershipSpread$eu_anti_pro_ls_spreadC <- ((EULeadershipSpread$eu_anti_pro_ls_spread) - mean(EULeadershipSpread$eu_anti_pro_ls_spread))

EULeadershipSpread$Inter_euAntiProls_EUldr <- ((EULeadershipSpread$EUldrSpreadC) * (EULeadershipSpread$eu_anti_pro_ls_sp))

EULeadershipResults16 <- lm(log(DefSpnd_lead) ~ EUldrSpreadC + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread + eu_anti_pro_ls_spreadC + Inter_euAntiProls_EUldr, EULeadershipSpread)
screenreg(list(EULeadershipResults1, EULeadershipResults2, EULeadershipResults6, EULeadershipResults8, EULeadershipResults11, EULeadershipResults13, EULeadershipResults14, EULeadershipResults16))

##interaction term is not significant, thus it will not be included. 

```


**Tests of within and pooling** we should explain this in the appendix 

``` {r EULeadershippooledTest, echo = TRUE}

EULeadershipPooled <- plm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread + eu_anti_pro_ls_spread, data = EULeadershipSpread, model = "pooling")

EULeadershipWithin <- plm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread + eu_anti_pro_ls_spread, data =  EULeadershipSpread, model = "within")

pFtest(EULeadershipWithin, EULeadershipPooled )
##p <2.2e^-16 thus reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```


**Fixed Effects model**
```{r EULeadershipspreadFixed, echo = TRUE}

EULeadershipResultsA <- plm(log(DefSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread + eu_anti_pro_ls_spread, EULeadershipSpread, index=c("Country", "Year"), model="within") 
summary(EULeadershipResultsA)

screenreg(list(EULeadershipResults14, EULeadershipResultsA))

pFtest(EULeadershipResultsA, EULeadershipResults14)

```


**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r EquDefSperadHausmanTest, echo = TRUE}

EULeadershipResultsWithin <- plm(log(DefSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread + eu_anti_pro_ls_spread, data = EULeadershipSpread, index=c("Country", "Year"), model="within") 

EULeadershipResultsRandom <- plm(log(DefSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread + eu_anti_pro_ls_spread, data = EULeadershipSpread, index=c("Country", "Year"), model="random")

phtest(EULeadershipResultsWithin, EULeadershipResultsRandom)
##p - 3.439e^-08, reject null and use fixed effects

```
