---
title: "Country spending too much or too little on defense data and defense spending"
author: "src"
date: "October 16, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)
require(stat)


#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
# setwd("H:/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}


# debug(CompilePubOpDataOmnibus)
##load data using function
debug(CompilePubOpDataOmnibus)
EuropeOmnibus <- CompilePubOpDataOmnibus()




write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))

##get rird of 2002 and 2003
#EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002", "2003"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors



```


###Country spending too much or too little on defense data --> defense spending

**load and subset the data we want**
```{r DefSpreadNarrowing, echo = FALSE, results = "hide"}

##load data using function
EuropeDefSpread <- subset(EuropeOmnibus, !is.na(DefSpread))


##get rid of sweden
EuropeDefSpread <- subset(EuropeDefSpread, !Country %in% c("Sweden", "Serbia", "Bulgaria", "Romania"))


```


**OLS model adding each variable individually**
```{r DefSpreadOLS, echo = TRUE}

VariableList<-subset(EuropeDefSpread,select=c(Population,GDPpCap))
VariableList$population<-log(VariableList$Population)
VariableList$GDPpCap<-log(VariableList$GDPpCap)
 rcorr(as.matrix(VariableList))


EUDefSpreadResults1 <- lm(log(DefSpnd_lead) ~ DefSpread, EuropeDefSpread) 
screenreg(list(EUDefSpreadResults1),
          custom.model.names=c("PubOp"))


EUDefSpreadResults7 <- lm(log(DefSpnd_lead) ~ DefSpread + log(Population), EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults7),
          custom.model.names=c("PubOp","lPop"))
summary(EUDefSpreadResults7)

##log population increases r^2 by .8, is significant at the .001% level, keeps def spread -0.00. will be included in the model. 


EUDefSpreadResults8 <- lm(log(DefSpnd_lead) ~ DefSpread + log(Population) + log(GDPpCap), EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults7, EUDefSpreadResults8),
          custom.model.names=c("PubOp","lPop","lGDPpCap"))
summary(EUDefSpreadResults8)

##gdppc increases r^2 by .21, which can potentially be bad. it also increases the significant of def spread to .05%, is itself significant at the .001 level. will be included. 


EUDefSpreadResults3 <- lm(log(DefSpnd_lead) ~ DefSpread + log(Population) + log(GDPpCap) + IntAt, EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults3),
          custom.model.names=c("PubOp","lPop","lGDPpCap","IntAt"))

##IntAt increases r^2 to .09 and is significant 95% will be included



EUDefSpreadResults4 <- lm(log(DefSpnd_lead) ~ DefSpread + log(Population) + log(GDPpCap) + IntAt + DomAt, EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults3, EUDefSpreadResults4),
          custom.model.names=c("PubOp","lPop","lGDPpCap","IntAt","DomAt"))

screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults4))
summary(EUDefSpreadResults4)

##DomAt is not statistically significant nor does it change r^2 or any other coefficient. will not be included


EUDefSpreadResults5 <- lm(log(DefSpnd_lead) ~ DefSpread + log(Population) + log(GDPpCap) + IntAt + DomAt + CivilWar, EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults3, EUDefSpreadResults4,EUDefSpreadResults5),
          custom.model.names=c("PubOp","lPop","lGDPpCap","IntAt","DomAt","CivilWar"))
summary(EUDefSpreadResults5)

##CivilWar is not significant, doesn't affect other coefficients (except sign of defspread but that's zero so doesn't matter really) and increases r^2 by .01 - will not be included.


EUDefSpreadResults6 <- lm(log(DefSpnd_lead) ~ DefSpread + log(Population) + log(GDPpCap) + IntAt + DomAt + CivilWar + IntlCnf, EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults3, EUDefSpreadResults4,EUDefSpreadResults5, EUDefSpreadResults6),
          custom.model.names=c("PubOp","lPop","lGDPpCap","IntAt","DomAt","CivilWar","IntlCnf"))
summary(EUDefSpreadResults6)


##IntlCnf does not change other coefficients significantly, is not significant and only increases r^2 by .01 - will not be included.



##log population increases r^2 by .5, is significant at the .001% level, keeps def spread -0.00. will be included in the model. 


EUDefSpreadResults8 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap), EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8))
summary(EUDefSpreadResults8)

##gdppc increases r^2 by .21, which can potentially be bad. it also increases the significant of def spread to .05%, is itself significant at the .001 level. will be included. 


EUDefSpreadResults6 <- lm(log(DefSpnd_lead) ~ DefSpread + log(Population) + log(GDPpCap) + IntAt + DomAt + CivilWar + IntlCnf, EuropeDefSpread)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults3, EUDefSpreadResults4,EUDefSpreadResults5, EUDefSpreadResults6),
          custom.model.names=c("PubOp","lPop","lGDPpCap","IntAt","DomAt","CivilWar","IntlCnf"))
summary(EUDefSpreadResults6)


EUDefSpreadResults9 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_left_right, EuropeDefSpread)
summary(EUDefSpreadResults9)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults9))

##Cablefright doesn't affect r^2 or the other variables and is not statistically significant thus will not be included

EUDefSpreadResults10 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, EuropeDefSpread)
summary(EUDefSpreadResults10)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults10))

##CabLibertyAuthority doesn't change r^2, it is not significant but it does increase DefSpread by .01 which is kind of interesting. before it was .0047
#Will try CabLibertyAuthority as an interaction with DefSpread

EUDefSpreadResults11 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, EuropeDefSpread)
summary(EUDefSpreadResults11)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults11))

##cab_eu_anti_pro is statisticaly significant and increases r^2 by .01

EUDefSpreadResults12 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority + Cab_eu_anti_pro, EuropeDefSpread)
summary(EUDefSpreadResults12)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults11, EUDefSpreadResults12))

##not sure if we should include  both cabinet_eu_anti_pro and cabinit_liberty_authority.......since the coefficient on defspread with libertyauthority is actually only .0050, increased from .0047, I think that cabinet_liberty_authority doesn't really have an effect. 

EUDefSpreadResults13 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + left_right_ls_spread, EuropeDefSpread)
summary(EUDefSpreadResults13)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults11, EUDefSpreadResults13))

##left_right_spread is not significant and does not increase r^2/change other variables

EUDefSpreadResults14 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + liberty_authority_ls_spread, EuropeDefSpread)
summary(EUDefSpreadResults14)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults11, EUDefSpreadResults14))

##liberty_authority_lsspread is not significant, doesn't increase r squared and doens't affect the other veriables significantly

EUDefSpreadResults15 <- lm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + eu_anti_pro_ls_spread, EuropeDefSpread)
summary(EUDefSpreadResults15)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults11, EUDefSpreadResults15))

##EU_anti_pro_ls_spreard is statistically significant, it increases the significance and value of defspread and it increases r^2 by .01, will be included and the interaction between defspread and euantiprolsspread will be tested. 

```


**Interactions in government variables:**
``` {r DefSpreadInteraction, echo=TRUE}


##center variables

EuropeDefSpread$eu_anti_pro_ls_spreadC <- ((EuropeDefSpread$eu_anti_pro_ls_spread) - mean(EuropeDefSpread$eu_anti_pro_ls_spread))

EuropeDefSpread$DefSpreadC <- ((EuropeDefSpread$DefSpread) - mean(EuropeDefSpread$DefSpread))

EuropeDefSpread$intrDefSpr_EuAntiPro <- (EuropeDefSpread$DefSpreadC * EuropeDefSpread$eu_anti_pro_ls_spreadC)

EUDefSpreadResults16 <- lm(log(DefSpnd_lead) ~ DefSpreadC + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro + eu_anti_pro_ls_spreadC + intrDefSpr_EuAntiPro, EuropeDefSpread)
summary(EUDefSpreadResults16)
screenreg(list(EUDefSpreadResults1, EUDefSpreadResults3, EUDefSpreadResults7, EUDefSpreadResults8, EUDefSpreadResults11, EUDefSpreadResults15, EUDefSpreadResults16))

#the interaction term is not significant and does not increase r^2, thus it will not be included in the model. 

```


**Tests of within and pooling** we should explain this in the 

``` {r defspreadPoolingTest, echo = TRUE}

DefSpreadPooled <- plm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, data = EuropeDefSpread, model = "pooling")

DefSpreadWithin <- plm(log(DefSpnd_lead) ~ DefSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, data =  EuropeDefSpread, model = "within")

pFtest(DefSpreadWithin, DefSpreadPooled)
##p <2.2e^-16 thus reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```


**Fixed Effects model** (within)
```{r DefSpreadFixed, echo = TRUE}

EUDefSpreadResultsA <- plm(log(DefSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, data = EuropeDefSpread, index=c("Country", "Year"), model="within") 
summary(EUDefSpreadResultsA)

screenreg(list(EUDefSpreadResults11, EUDefSpreadResultsA))

#testing for fixed effects
pFtest(EUDefSpreadResultsA, EUDefSpreadResults9)

```

**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r DefSperadHausmanTest, echo = TRUE}

EUDefSpreadResultsWithin <- plm(log(DefSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, data = EuropeDefSpread, index=c("Country", "Year"), model="within") 

EUDefSpreadResultsRandom <- plm(log(DefSpnd_lead) ~ DefSpread +  IntAt + log(Population) + log(GDPpCap) +Cab_eu_anti_pro, data = EuropeDefSpread, index=c("Country", "Year"), model="random")

phtest(EUDefSpreadResultsWithin, EUDefSpreadResultsRandom)
##p - 4.45e^-6, reject null and use fixed effects

```

