---
title: "SSI Regression"
author: "src"
date: "Wednesday, April 08, 2015"
output:
  html_document:
    keep_md: yes
---

###Analysis for estimating the effect EU Public opinion has on Defense Spending 

###U.S. Leadership Data


**Load packages and support files**
```{r Setup, echo = TRUE }
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)

# setwd("C:/Users/scohen/My Documents/Europe/Europe/Europe/") #Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
setwd("C:/Users/Greg Sanders/Documents/Development/Europe")

source("SSIRegression.R")

```


**Load and subset the data we want**
```{r Input, echo = TRUE}



##load data using function
EuropeOmnibus <- CompilePubOpDataOmnibus(lag = 1)

#Exclude US and aggregate categories
EuropeOmnibus <- subset(EuropeOmnibus,!Country %in% c("EU 10","EU 7","USA","EU 11","EU 9","EU 12","EU 8","Russian"))
  

##create data frame to test for correlation between variables
EuropeOmnibus<-data.frame(
  Dspend  = EuropeOmnibus$DefSpnd,
  ThrtR = EuropeOmnibus$ThreatRatio,
  IntAt = EuropeOmnibus$IntAt,
  DomAt = EuropeOmnibus$DomAt,
  CivWr = EuropeOmnibus$CivilWar,
  IntWr=  EuropeOmnibus$IntlCnf,
  Pop =  EuropeOmnibus$Population,
  GDPpC = EuropeOmnibus$GDPpCap,
  Dem = EuropeOmnibus$democ,
  NATO = EuropeOmnibus$NATOally,
  USldrSpread =EuropeOmnibus$USldrSpread,
  DefSpread =EuropeOmnibus$DefSpread,
  Country = EuropeOmnibus$Country,
  Year = EuropeOmnibus$Year
  
)


```

**create data frame to test correlation between variables and test them**
```{r USldrSpread, Correlation, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EuropeOmnibus,select=-c(Country,Year))))

```

**Summary statistics of data**
```{r USldrSpreadSummary, echo = TRUE}

summary(EuropeOmnibus)

```

**OLS models, adding each variable individually**
```{r USldrSpreadBuilding, echo = TRUE}
USleadResults1 <- lm(log(Dspend) ~ USldrSpread, EuropeOmnibus) 
USleadResults2 <- lm(log(Dspend) ~ USldrSpread + ThrtR, EuropeOmnibus)
USleadResults3 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt, EuropeOmnibus)
USleadResults4 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt, EuropeOmnibus)
USleadResults5 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr, EuropeOmnibus)
USleadResults6 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr, EuropeOmnibus)
USleadResults7 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop), EuropeOmnibus)
USleadResults8 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC), EuropeOmnibus)
USleadResults9 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem, EuropeOmnibus)
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
screenreg(list(USleadResults1, USleadResults2, USleadResults3, USleadResults4, USleadResults5, USleadResults6, USleadResults7, USleadResults8, USleadResults9, USleadResults10))

plot(USleadResults10)

```


**Hypothesis testing**

2. F-tests
``` {r USleadSpreadFtest, echo = TRUE}
##For IntAt and GDPpc
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults1 <- lm(log(Dspend) ~ USldrSpread + ThrtR + DomAt + CivWr + IntWr + log(Pop) + Dem + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults1, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivWr and DomAt
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults2 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults2, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivWr and IntWr
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults3 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults3, ratio = .000001, conf.level = 0.95)
##Cannot reject the null hypothesis that both of these coefficients are equal to zero.  
#because of this, IntWr might not merit inclusion in the model becasue it's statistics show that by itself, it's not statistically significant at the 95% level, it only increases r^2 by .01 when added and it doesn't affect the other variables.  (except a decrease in ThrtR by .05 but this is not a huge impact on the model)
#IntWr will also be tested with Dem so that f-test will determine it's merit in the model.
##Civil war is already significant with other variables.

##For CivWr and Dem
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults4 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + IntWr + log(Pop) + log(GDPpC) + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults4, ratio = .000001, conf.level = 0.95)
##Reject the null hypothesis that both of these coefficients are equal to zero

##For IntWr and Dem
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults5 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + log(Pop) + log(GDPpC) + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that  both of these coefficients are equal to zero
##This means that the above uncertainty of IntWr's significance is resolved and IntWr should stay in the model

##For GDPpc and CivWr
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults6 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + IntWr + log(Pop) + Dem + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults6, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##For Dem and USldrSpread
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults7 <- lm(log(Dspend) ~ ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults7, ratio = .000001, conf.level = .95)
## reject the null hypothesis that both of these coefficients are equal to zero

##For CivWr and USldrSpread
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults8 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults8, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##for USldrSpread and Dem
USleadResults10 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
Hresults9 <- lm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeOmnibus)
var.test(USleadResults10, Hresults9, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero


```


**Fixed effects model**
```{r USleaderFixedEffects, echo = TRUE}

Bresults1 <- plm(log(Dspend) ~ USldrSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, data=EuropeOmnibus, index=c("Country", "Year"), model="within")
summary(Bresults1)
#plot(Bresults1)

#Comparing OLS to fixed effects model and testing for fixed effects
screenreg(list(USleadResults10, Bresults1))

pFtest(Bresults1, USleadResults10)


```


###Country spending too much or too little on defense data

**load and subset the data we want**
```{r DefSpreadNarrowing, echo = TRUE}

##load data using function
EuropeDefSpread <- subset(EuropeOmnibus, !is.na(DefSpread))

EuropeDefSpread <- subset(EuropeDefSpread,!Country %in% c("EU 10","EU 7","USA","EU 11","EU 9","EU 12","EU 8"))


```

**summary statistics of data**
```{r DefSpreadSummary, echo = TRUE}

##Summary statistics of data
summary(EuropeDefSpread)

```

**create data frame to test correlation between variables and test them**
```{r DefSpreadCorr, echo = TRUE}

##Test correlation between variables 
rcorr(as.matrix(subset(EuropeDefSpread,select=-c(Country,Year))))

```



**OLS model adding each variable individually**
```{r DefSpreadBuilding, echo = TRUE}
Cresults1 <- lm(log(Dspend) ~ DefSpread, EuropeDefSpread) 
Cresults2 <- lm(log(Dspend) ~ DefSpread + ThrtR, EuropeDefSpread)
Cresults3 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt, EuropeDefSpread)
Cresults4 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt, EuropeDefSpread)
Cresults5 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr, EuropeDefSpread)
Cresults6 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr, EuropeDefSpread)
Cresults7 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop), EuropeDefSpread)
Cresults8 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC), EuropeDefSpread)
Cresults9 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem, EuropeDefSpread)
Cresults10 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
screenreg(list(Cresults1, Cresults2, Cresults3, Cresults4, Cresults5, Cresults6, Cresults7, Cresults8, Cresults9, Cresults10))

plot(Cresults10)

```


**Hypothesis testing**
``` {r DefSpreadTesting, echo = TRUE}

##ThrtR and GDPpc
Cresults10 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
Gresults1 <- lm(log(Dspend) ~ DefSpread + IntAt + DomAt + CivWr + IntWr + log(Pop) + Dem + NATO, EuropeDefSpread)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivWr
Cresults10 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
Gresults2 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivWr and IntWr
Cresults10 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
Gresults3 <-  lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivWr and Dem
Cresults10 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
Gresults4 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + IntWr + log(Pop) + log(GDPpC) + NATO, EuropeDefSpread)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
Gresults5 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + log(Pop) + log(GDPpC) + NATO, EuropeDefSpread)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Pop and Dem
Cresults10 <- lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
Gresults6 <-  lm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, EuropeDefSpread)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r DefSpreadFixed, echo = TRUE}

Dresults1 <- plm(log(Dspend) ~ DefSpread + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, data=EuropeDefSpread, index=c("Country", "Year"), model="within")
summary(Dresults1)

screenreg(list(Cresults10, Dresults1))

pFtest(Dresults1, Cresults10)

```
