---
title: "SSI Regression"
author: "src"
date: "Wednesday, April 08, 2015"
output:
  html_document:
    keep_md: yes
---

###Analysis for estimating the effect EU Public opinion has on Defense Spending 

**Load the Data**
```{r echo=TRUE}
# setwd("C:/Users/scohen/My Documents/Europe/Europe/Data/") #Your working directory here!
# #source("SSIRegression.R")
# 
# CompilePubOpData <- function(filename, lag = 1) {
#   ## start by setting up some items for later use.
#   ## Namely, loading needed packages and setting a path to my files.
#   require(plm)
#   require(plyr)
#   require(reshape)
#   
#   ## Set this path to the folder into which your git hub will download data
#   #path <- "K:/Development/Europe/" #     path <- "C:/Users/MRiley/My Documents/Europe/"
#   #path <- "C:/Users/scohen/My Documents/Europe/Europe/"
#   path <- ("C:/Users/MRiley/My Documents/Europe/")
#   
#   ## Next I'm going to load all of my data. The data: in order is..
#   ## public opinion, governance data from PolityIV, Terrorism data from GTD,
#   ## data we compiled on conflicts and a country's membership to the EU,
#   ## GDP per capita data (constant 2005 $), GDP data (also Const 2005 $), population data,
#   ## data on NATO membership, spending data, and neighbor spending data.
#   data.a <- read.csv(paste(path, filename, sep =""), header = TRUE) 
#   data.gov <- read.csv(paste(path, "SSI_Govern.csv", sep =""), header = TRUE)
#   data.ter <- read.csv(paste(path, "Terrorism Data.csv", sep =""), header = TRUE)
#   data.intlcnf <- read.csv(paste(path, "SSI_IntlConfl.csv", sep =""), header = TRUE)
#   data.cvlwr <- read.csv(paste(path, "SSI_CivilWar.csv", sep =""), header = TRUE)
#   data.gdppc <- read.csv(paste(path, "SSI_GDPperCAP.csv", sep =""), header = TRUE)
#   data.gdp <- read.csv(paste(path, "SSI_Const05_GDP.csv", sep =""), header = TRUE)
#   data.pop <- read.csv(paste(path, "SSI_Population.csv", sep =""), header = TRUE)
#   data.nato <- read.csv(paste(path, "SSI_NATO.csv", sep =""), header = TRUE)
#   data.euds <- read.csv(paste(path, "EUDefenseSpending_EUROS.csv", sep =""), header = TRUE)
#   data.nghspnd <- read.csv(paste(path, "SSIMilSpendingData.CSV", sep=""), header = TRUE, na.strings = "#VALUE!")
#   
#   
#   #### This next section is where we change the column names of the data sets that don't need
#   #### to be reshaped. 
#   
#   ## Then we change the coloumnames to make them more universal
#   colnames(data.gov)[colnames(data.gov)=="country"] <- "Country"
#   colnames(data.gov)[colnames(data.gov)=="year"] <- "Year"
#   colnames(data.ter)[colnames(data.ter)=="country_txt"] <- "Country"
#   colnames(data.ter)[colnames(data.ter)=="iyear"] <- "Year"
#   data.euds <- rename(data.euds, c("X2001"="2001", "X2002"="2002", "X2003"="2003", "X2004"="2004", "X2005"="2005", "X2006"="2006", "X2007"="2007", "X2008"="2008", "X2009"="2009", "X2010"="2010", "X2011"="2011", "X2012"="2012", "X2013"="2013"))
#   colnames(data.gdppc)[colnames(data.gdppc)=="United.Kingdom"] <- "UK"
#   colnames(data.gdppc)[colnames(data.gdppc)=="Slovak.Republic"] <- "Slovakia"
#   colnames(data.gdppc)[colnames(data.gdppc)=="Russian.Federation"] <- "Russia"
#   colnames(data.pop)[colnames(data.pop)=="United.Kingdom"] <- "UK"
#   colnames(data.pop)[colnames(data.pop)=="Slovak.Republic"] <- "Slovakia"
#   colnames(data.pop)[colnames(data.pop)=="Russian.Federation"] <- "Russia"   
#   colnames(data.euds)[colnames(data.euds)=="United Kingdom"] <- "UK"
#   colnames(data.euds)[colnames(data.euds)=="Slovak.Republic"] <- "Slovakia"
#   colnames(data.euds)[colnames(data.euds)=="Russian.Federation"] <- "Russia"
#   colnames(data.euds)[colnames(data.euds)=="Serbia*"] <- "Serbia"
#   
#   #### In this next component, we will reshape and fit data so we can synthesize with with
#   #### the other data that we have.
#   
#   ##reshaping EU defense spending data and creating log(GDP)
#   data.gdp <- rename(data.gdp, c("X2000"="2000", "X2001"="2001", "X2002"="2002", "X2003"="2003", "X2004"="2004", "X2005"="2005", "X2006"     ="2006", "X2007"="2007", "X2008"="2008", "X2009"="2009", "X2010"="2010", "X2011"="2011", "X2012"="2012", "X2013"="2013"))
#   data.gdp <- melt(data.gdp, id="Country")
#   data.gdp <- rename(data.gdp, c("variable"="Year", "value"="GDP2005usd"))
#   data.gdp <- arrange(data.gdp, Country)
#   #data.gdp$logGDP <- log(data.gdp$GDP2005usd)
#   
#   ##reshaping neighbor spending data 
#   data.nghspnd <- rename(data.nghspnd, c("X2000"="2000", "X2001"="2001", "X2002"="2002", "X2003"="2003", "X2004"="2004", "X2005"="2005", "X2006"     ="2006", "X2007"="2007", "X2008"="2008", "X2009"="2009", "X2010"="2010", "X2011"="2011", "X2012"="2012", "X2013"="2013"))
#   data.nghspnd <- melt(data.nghspnd, id=c("COUNTRY", "Country.List"))
#   data.nghspnd <- rename(data.nghspnd, c("COUNTRY"="Country","variable"="Year", "value"="neighspend"))
#   data.nghspnd <- data.nghspnd[,c(1,3,4)]
#   data.nghspnd <- arrange(data.nghspnd, Country)
#   
#   ## Reshaping international conflict data
#   data.intlcnf <- melt(data.intlcnf, id = "Year")
#   data.intlcnf <- rename(data.intlcnf, c("variable"="Country", "value"="IntlCnf"))
#   data.intlcnf$Year <- as.integer(as.character(data.intlcnf$Year))
#   data.intlcnf$Country <- as.character(data.intlcnf$Country)
#   
#   ## Reshaping civil war data
#   data.cvlwr <- data.cvlwr[-c(1), ]
#   data.cvlwr <- melt(data.cvlwr, id = "Year")
#   data.cvlwr <- rename(data.cvlwr, c("variable"="Country", "value"="CivilWar"))
#   data.cvlwr$Year <- as.integer(as.character(data.cvlwr$Year))
#   data.cvlwr$Country <- as.character(data.cvlwr$Country)
#   
#   ## Combining Neighbor Spending and GDP data to create a threat ratio variable
#   ## The value of this variable is NghSpnd/GDP
#   threatvariable <- as.data.frame(NULL)
#   c.loop <- unique(data.nghspnd$Country)
#   t.loop <- unique(data.nghspnd$Year)
#   
#   for (i in c.loop) {
#     numerator <- data.nghspnd[data.nghspnd$Country == i,]
#     denominator <- data.gdp[data.gdp$Country == i,]
#     for (t in t.loop){
#       numerator.use <- numerator[numerator$Year == t,]
#       denominator.use <- denominator[denominator$Year == t,]
#       x <- (numerator.use[1,3]/denominator.use[1,3])*1000000
#       row <- data.frame(Country = i, Year = t, ThreatRatio = x)
#       threatvariable <- rbind(threatvariable, row)
#     }
#     
#   }
#   threatvariable$Country <- as.character(threatvariable$Country)
#   threatvariable$Year <- as.integer(as.character(threatvariable$Year))
#   
#   
#   ## We need to reshape and rename the EU defense spending data
#   data.euds <- rename(data.euds, c("X2001"="2001", "X2002"="2002", "X2003"="2003", "X2004"="2004", "X2005"="2005", "X2006"     ="2006", "X2007"="2007", "X2008"="2008", "X2009"="2009", "X2010"="2010", "X2011"="2011", "X2012"="2012", "X2013"="2013"))
#   data.euds <- melt(data.euds, id=c("Country", "Unit.Currency"))
#   data.euds <- rename(data.euds, c("variable"="Year", "value"="DefSpnd"))
#   data.euds <- data.euds[,c(1,3,4)]
#   data.euds[,2] <- as.integer(as.character(data.euds[,2]))
#   data.euds$Year <- data.euds$Year-lag
#   data.euds[,3] <- data.euds[,3]*1000000
#   data.euds$Country <- as.character(data.euds$Country)
#   data.euds$Country[data.euds$Country == "United Kingdom"] <- "UK" 
#   #data.euds$logeuds <- log(data.euds$DefSpnd)
#   
#   ## Also need to reshape the GDP per Capita data, and then rename some of the columns
#   data.pcap <- melt(data.gdppc, id = "Year")
#   colnames(data.pcap)[colnames(data.pcap)=="year"] <- "Year"
#   colnames(data.pcap)[colnames(data.pcap)=="variable"] <- "Country"
#   colnames(data.pcap)[colnames(data.pcap)=="value"] <- "GDPpCap"
#   
#   ## Also need to reshape the population data, and then rename some of the columns
#   data.population <- melt(data.pop, id = "Year")
#   colnames(data.population)[colnames(data.population)=="year"] <- "Year"
#   colnames(data.population)[colnames(data.population)=="variable"] <- "Country"
#   colnames(data.population)[colnames(data.population)=="value"] <- "Population"
#   
#   ## Also need to reshape the in NATO/alliance, and then rename some of the columns
#   data.ally <- melt(data.nato, id = "Year")
#   colnames(data.ally)[colnames(data.ally)=="year"] <- "Year"
#   colnames(data.ally)[colnames(data.ally)=="variable"] <- "Country"
#   colnames(data.ally)[colnames(data.ally)=="value"] <- "NATOally"
#   
#   
#   ## This next section is to synthesize the terrorism data and boil it down to
#   ## the information we actually need for the regression
#   data.ter <- data.ter[,2:6]
#   data.ter$Country <- as.character(data.ter$Country)
#   data.ter$Country[data.ter$Country == "Great Britain"] <- "UK"
#   countryloop <- sort(unique(data.ter$Country))
#   timeloop <- sort(unique(data.ter$Year), decreasing= FALSE)
#   attacks <- as.data.frame(NULL)
#   
#   
#   complete.ter <- data.ter[complete.cases(data.ter),]
#   #     ddply(complete.ter,
#   #           .(Country,Year),
#   #           summarise,
#   #           attacks=nrow())
#   for (i in countryloop){ 
#     use <- complete.ter[complete.ter$Country== i,]
#     for (t in timeloop){
#       use.a <- use[use$Year == t,]
#       row <- data.frame(Country=i, Year=t, Attacks=nrow(use.a), IntAt=sum(use.a[,3]), DomAt=nrow(use.a)-sum(use.a[,3]))
#       attacks <- rbind(attacks, row)
#       
#     }
#     
#   }
#   
#   
#   ## Now it is time to start combining the data, so we can run the regression
#   out1 <- plyr::join(data.a, data.gov, by = c("Country", "Year"))
#   out2 <- plyr::join(out1, data.intlcnf, by = c("Country", "Year"))
#   out3 <- plyr::join(out2, data.cvlwr, by = c("Country", "Year"))
#   out4 <- plyr::join(out3, data.pcap, by = c("Country", "Year"))
#   out5 <- plyr::join(out4, data.population, by = c("Country", "Year"))
#   out6 <- plyr::join(out5, data.ally, by = c("Country", "Year"))
#   out7 <- plyr::join(out6, attacks, by = c("Country", "Year"))
#   out8 <- plyr::join(out7, data.euds, by = c("Country", "Year"))
#   output <- plyr::join(out8, threatvariable, by = c("Country", "Year"))
#   
#   View(output)  
#   
#   output
#   
#   
# } 
# 
# 
# require("Hmisc")
# require(texreg)
# require(plm)
# require(ggplot)
```


###U.S. Leadership Data

**load and subset the data we want**
```{r echo = TRUE}
source("SSIRegression.R")

setwd("C:/Users/scohen/My Documents/Europe/Europe/Europe/") #Your working directory here!

uslead.1lag <- CompilePubOpData("SSI_US_Leader_Data.csv", lag = 1)

regdat <- uslead.1lag[34:152,]

regdat$NATOally[is.na(regdat$NATOally)]<-0

```

**load packages**
```{r echo = TRUE}

require(Hmisc)
require(texreg)
require(plm)
require(ggplot)

```


**Re-label data**
```{r echo = TRUE}
Dspend  <- regdat$DefSpnd
ThrtR <- regdat$ThreatRatio
IntAt <- regdat$IntAt
DomAt <- regdat$DomAt
CivWr <- regdat$CivilWar
IntWr <- regdat$IntlCnf
Pop <- regdat$Population
GDPpC <- regdat$GDPpCap
Dem <- regdat$democ
NATO <- regdat$NATOally
PubOp <- regdat$Spread

```

**create data frame to test correlation between variables and test them**
```{r echo = TRUE}
reg_df<-data.frame(
  Dspend  = regdat$DefSpnd,
  ThrtR = regdat$ThreatRatio,
  IntAt = regdat$IntAt,
  DomAt = regdat$DomAt,
  CivWr = regdat$CivilWar,
  IntWr=  regdat$IntlCnf,
  Pop =  regdat$Population,
  GDPpC = regdat$GDPpCap,
  Dem = regdat$democ,
  NATO = regdat$NATOally,
  PubOp =regdat$Spread
)

rcorr(as.matrix(reg_df))

```

**Summary statistics of data**
```{r echo = TRUE}

summary(regdat)

```

**OLS models, adding each variable individually**
```{r echo = TRUE}
Aresults1 <- lm(log(Dspend) ~ PubOp, regdat) 
Aresults2 <- lm(log(Dspend) ~ PubOp + ThrtR, regdat)
Aresults3 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt, regdat)
Aresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt, regdat)
Aresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr, regdat)
Aresults6 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr, regdat)
Aresults7 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop), regdat)
Aresults8 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC), regdat)
Aresults9 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem, regdat)
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
screenreg(list(Aresults1, Aresults2, Aresults3, Aresults4, Aresults5, Aresults6, Aresults7, Aresults8, Aresults9, Aresults10))

plot(Aresults10)

```


**Hypothesis testing**

2. F-tests
``` {r echo = TRUE}
##For IntAt and GDPpc
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults1 <- lm(log(Dspend) ~ PubOp + ThrtR + DomAt + CivWr + IntWr + log(Pop) + Dem + NATO, regdat)
var.test(Aresults10, Hresults1, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivWr and DomAt
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults2 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults2, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivWr and IntWr
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults3 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults3, ratio = .000001, conf.level = 0.95)
##Cannot reject the null hypothesis that both of these coefficients are equal to zero.  
#because of this, IntWr might not merit inclusion in the model becasue it's statistics show that by itself, it's not statistically significant at the 95% level, it only increases r^2 by .01 when added and it doesn't affect the other variables.  (except a decrease in ThrtR by .05 but this is not a huge impact on the model)
#IntWr will also be tested with Dem so that f-test will determine it's merit in the model.
##Civil war is already significant with other variables.

##For CivWr and Dem
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + IntWr + log(Pop) + log(GDPpC) + NATO, regdat)
var.test(Aresults10, Hresults4, ratio = .000001, conf.level = 0.95)
##Reject the null hypothesis that both of these coefficients are equal to zero

##For IntWr and Dem
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + log(Pop) + log(GDPpC) + NATO, regdat)
var.test(Aresults10, Hresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that  both of these coefficients are equal to zero
##This means that the above uncertainty of IntWr's significance is resolved and IntWr should stay in the model

##For GDPpc and CivWr
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults6 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + IntWr + log(Pop) + Dem + NATO, regdat)
var.test(Aresults10, Hresults6, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##For Dem and PubOp
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults7 <- lm(log(Dspend) ~ ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + NATO, regdat)
var.test(Aresults10, Hresults7, ratio = .000001, conf.level = .95)
## reject the null hypothesis that both of these coefficients are equal to zero

##For CivWr and PubOp
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults8 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults8, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##for PubOp and Dem
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults9 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults9, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero


```


**Fixed effects model**
```{r echo = TRUE}

Bresults1 <- plm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, data=regdat, index=c("Country", "Year"), model="within")
summary(Bresults1)
#plot(Bresults1)

```

**Comparing OLS to fixed effects model and testing for fixed effects**
```{r echo = TRUE}

screenreg(list(Aresults10, Bresults1))

pFtest(Bresults1, Aresults10)

```


###Country spending too much or too little on defense data

**load and subset the data we want**
```{r echo = TRUE}

DefSpnd_IncDec_Data.1lag <- CompilePubOpData("SSI_DefSpnd_IncDec.csv", lag = 1)

regdat1 <- DefSpnd_IncDec_Data.1lag[34:152,]

regdat1$NATOally[is.na(regdat$NATOally)]<-0

```

**summary statistics of data**
```{r echo = TRUE}

summary(regdat1)

```

**create data frame to test correlation between variables and test them**
```{r echo = TRUE}
reg_df1<-data.frame(
  Dspend  = regdat1$DefSpnd,
  ThrtR = regdat1$ThreatRatio,
  IntAt = regdat1$IntAt,
  DomAt = regdat1$DomAt,
  CivWr = regdat1$CivilWar,
  IntWr=  regdat1$IntlCnf,
  Pop =  regdat1$Population,
  GDPpC = regdat1$GDPpCap,
  Dem = regdat1$democ,
  NATO = regdat1$NATOally,
  PubOp =regdat1$Spread
)

rcorr(as.matrix(reg_df1))

```



**OLS model adding each variable individually**
```{r echo = TRUE}
Cresults1 <- lm(log(Dspend) ~ PubOp, regdat1) 
Cresults2 <- lm(log(Dspend) ~ PubOp + ThrtR, regdat1)
Cresults3 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt, regdat1)
Cresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt, regdat)
Cresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr, regdat1)
Cresults6 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr, regdat1)
Cresults7 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop), regdat1)
Cresults8 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC), regdat1)
Cresults9 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem, regdat1)
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
screenreg(list(Cresults1, Cresults2, Cresults3, Cresults4, Cresults5, Cresults6, Cresults7, Cresults8, Cresults9, Cresults10))

plot(Cresults10)

```


**Hypothesis testing**
``` {r echo = TRUE}

##ThrtR and GDPpc
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults1 <- lm(log(Dspend) ~ PubOp + IntAt + DomAt + CivWr + IntWr + log(Pop) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivWr
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults2 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivWr and IntWr
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults3 <-  lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivWr and Dem
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + IntWr + log(Pop) + log(GDPpC) + NATO, regdat1)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + log(Pop) + log(GDPpC) + NATO, regdat1)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Pop and Dem
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults6 <-  lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r echo = TRUE}

Dresults1 <- plm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, data=regdat1, index=c("Country", "Year"), model="within")
summary(Dresults1)


```

**comparing OLS to Fixed Effects model**
```{r echo = TRUE}
screenreg(list(Cresults10, Dresults1))

pFtest(Dresults1, Cresults10)

```


