---
title: "SSI Regression"
author: "src"
date: "Wednesday, April 08, 2015"
output:
  html_document:
    keep_md: yes
---

###Analysis for estimating the effect EU Public opinion has on Defense Spending 

###U.S. Leadership Data

**load and subset the data we want**
```{r echo = TRUE}
source("SSIRegression.R")

# setwd("C:/Users/scohen/My Documents/Europe/Europe/Europe/") #Your working directory here!

uslead.1lag <- CompilePubOpData("SSI_US_Leader_Data.csv", lag = 1)

regdat <- uslead.1lag[34:152,]

regdat$NATOally[is.na(regdat$NATOally)]<-0

```

**load packages**
```{r echo = TRUE}

require(Hmisc)
require(texreg)
require(plm)
require(ggplot)

```


**Re-label data**
```{r echo = TRUE}
Dspend  <- regdat$DefSpnd
ThrtR <- regdat$ThreatRatio
IntAt <- regdat$IntAt
DomAt <- regdat$DomAt
CivWr <- regdat$CivilWar
IntWr <- regdat$IntlCnf
Pop <- regdat$Population
GDPpC <- regdat$GDPpCap
Dem <- regdat$democ
NATO <- regdat$NATOally
PubOp <- regdat$Spread

```

**create data frame to test correlation between variables and test them**
```{r echo = TRUE}
reg_df<-data.frame(
  Dspend  = regdat$DefSpnd,
  ThrtR = regdat$ThreatRatio,
  IntAt = regdat$IntAt,
  DomAt = regdat$DomAt,
  CivWr = regdat$CivilWar,
  IntWr=  regdat$IntlCnf,
  Pop =  regdat$Population,
  GDPpC = regdat$GDPpCap,
  Dem = regdat$democ,
  NATO = regdat$NATOally,
  PubOp =regdat$Spread
)

rcorr(as.matrix(reg_df))

```

**Summary statistics of data**
```{r echo = TRUE}

summary(regdat)

```

**OLS models, adding each variable individually**
```{r echo = TRUE}
Aresults1 <- lm(log(Dspend) ~ PubOp, regdat) 
Aresults2 <- lm(log(Dspend) ~ PubOp + ThrtR, regdat)
Aresults3 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt, regdat)
Aresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt, regdat)
Aresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr, regdat)
Aresults6 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr, regdat)
Aresults7 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop), regdat)
Aresults8 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC), regdat)
Aresults9 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem, regdat)
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
screenreg(list(Aresults1, Aresults2, Aresults3, Aresults4, Aresults5, Aresults6, Aresults7, Aresults8, Aresults9, Aresults10))

plot(Aresults10)

```


**Hypothesis testing**

2. F-tests
``` {r echo = TRUE}
##For IntAt and GDPpc
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults1 <- lm(log(Dspend) ~ PubOp + ThrtR + DomAt + CivWr + IntWr + log(Pop) + Dem + NATO, regdat)
var.test(Aresults10, Hresults1, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivWr and DomAt
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults2 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults2, ratio = .000001, conf.level = 0.95)
##reject null hypothesis that both of these coefficients are equal to zero

##For CivWr and IntWr
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults3 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults3, ratio = .000001, conf.level = 0.95)
##Cannot reject the null hypothesis that both of these coefficients are equal to zero.  
#because of this, IntWr might not merit inclusion in the model becasue it's statistics show that by itself, it's not statistically significant at the 95% level, it only increases r^2 by .01 when added and it doesn't affect the other variables.  (except a decrease in ThrtR by .05 but this is not a huge impact on the model)
#IntWr will also be tested with Dem so that f-test will determine it's merit in the model.
##Civil war is already significant with other variables.

##For CivWr and Dem
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + IntWr + log(Pop) + log(GDPpC) + NATO, regdat)
var.test(Aresults10, Hresults4, ratio = .000001, conf.level = 0.95)
##Reject the null hypothesis that both of these coefficients are equal to zero

##For IntWr and Dem
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + log(Pop) + log(GDPpC) + NATO, regdat)
var.test(Aresults10, Hresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that  both of these coefficients are equal to zero
##This means that the above uncertainty of IntWr's significance is resolved and IntWr should stay in the model

##For GDPpc and CivWr
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults6 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + IntWr + log(Pop) + Dem + NATO, regdat)
var.test(Aresults10, Hresults6, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##For Dem and PubOp
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults7 <- lm(log(Dspend) ~ ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + NATO, regdat)
var.test(Aresults10, Hresults7, ratio = .000001, conf.level = .95)
## reject the null hypothesis that both of these coefficients are equal to zero

##For CivWr and PubOp
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults8 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults8, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero

##for PubOp and Dem
Aresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
Hresults9 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat)
var.test(Aresults10, Hresults9, ratio = .000001, conf.level = .95)
##reject the null hypothesis that both of these coefficients are equal to zero


```


**Fixed effects model**
```{r echo = TRUE}

Bresults1 <- plm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, data=regdat, index=c("Country", "Year"), model="within")
summary(Bresults1)
#plot(Bresults1)

```

**Comparing OLS to fixed effects model and testing for fixed effects**
```{r echo = TRUE}

screenreg(list(Aresults10, Bresults1))

pFtest(Bresults1, Aresults10)

```


###Country spending too much or too little on defense data

**load and subset the data we want**
```{r echo = TRUE}

DefSpnd_IncDec_Data.1lag <- CompilePubOpData("SSI_DefSpnd_IncDec.csv", lag = 1)

regdat1 <- DefSpnd_IncDec_Data.1lag[34:152,]

regdat1$NATOally[is.na(regdat$NATOally)]<-0

```

**summary statistics of data**
```{r echo = TRUE}

summary(regdat1)

```

**create data frame to test correlation between variables and test them**
```{r echo = TRUE}
reg_df1<-data.frame(
  Dspend  = regdat1$DefSpnd,
  ThrtR = regdat1$ThreatRatio,
  IntAt = regdat1$IntAt,
  DomAt = regdat1$DomAt,
  CivWr = regdat1$CivilWar,
  IntWr=  regdat1$IntlCnf,
  Pop =  regdat1$Population,
  GDPpC = regdat1$GDPpCap,
  Dem = regdat1$democ,
  NATO = regdat1$NATOally,
  PubOp =regdat1$Spread
)

rcorr(as.matrix(reg_df1))

```



**OLS model adding each variable individually**
```{r echo = TRUE}
Cresults1 <- lm(log(Dspend) ~ PubOp, regdat1) 
Cresults2 <- lm(log(Dspend) ~ PubOp + ThrtR, regdat1)
Cresults3 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt, regdat1)
Cresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt, regdat)
Cresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr, regdat1)
Cresults6 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr, regdat1)
Cresults7 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop), regdat1)
Cresults8 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC), regdat1)
Cresults9 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem, regdat1)
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
screenreg(list(Cresults1, Cresults2, Cresults3, Cresults4, Cresults5, Cresults6, Cresults7, Cresults8, Cresults9, Cresults10))

plot(Cresults10)

```


**Hypothesis testing**
``` {r echo = TRUE}

##ThrtR and GDPpc
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults1 <- lm(log(Dspend) ~ PubOp + IntAt + DomAt + CivWr + IntWr + log(Pop) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults1, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##DomAt and CivWr
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults2 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults2, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivWr and IntWr
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults3 <-  lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults3, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##CivWr and Dem
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults4 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + IntWr + log(Pop) + log(GDPpC) + NATO, regdat1)
var.test(Cresults10, Gresults4, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##IntrWr and Dem
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults5 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + log(Pop) + log(GDPpC) + NATO, regdat1)
var.test(Cresults10, Gresults5, ratio = .000001, conf.level = .95)
##reject the null hypothesis that these two variables are equal to zero

##Pop and Dem
Cresults10 <- lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
Gresults6 <-  lm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, regdat1)
var.test(Cresults10, Gresults6, ratio = .000001, conf.level = .95)

##reject the null hypothesis that these two variables are equal to zero. 
```



**Fixed Effects model**
```{r echo = TRUE}

Dresults1 <- plm(log(Dspend) ~ PubOp + ThrtR + IntAt + DomAt + CivWr + IntWr + log(Pop) + log(GDPpC) + Dem + NATO, data=regdat1, index=c("Country", "Year"), model="within")
summary(Dresults1)


```

**comparing OLS to Fixed Effects model**
```{r echo = TRUE}
screenreg(list(Cresults10, Dresults1))

pFtest(Dresults1, Cresults10)

```