---
title: "DesirabilityForEUExertStrongLeadership_InvestmentSpending"
author: "src"
date: "October 16, 2015"
output: html_document
---

**Load packages and support files**
```{r Setup, echo = FALSE, results = "hide"}
options(error=recover)
options(warn=1)

require(Hmisc)
require(texreg)
require(plm)
require(ggplot2)

#Your working directory here!
# setwd("K:/Development/Europe") #Your working directory here!
# setwd("C:/Users/Greg Sanders/Documents/Development/Europe")
setwd("H:/Europe/Git/Europe")

source("EuropeInput.R")

```


**Load and subset the data we want**
```{r Input, echo = FALSE, results = "hide"}
debug(CompilePubOpDataOmnibus)
##load data using function
EuropeOmnibus <- CompilePubOpDataOmnibus()
debug(StandardizeCountries)
write.table(EuropeOmnibus
            ,file=paste("data\\EuropeOmnibus.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )


#get rid of greece
EuropeOmnibus <- subset(EuropeOmnibus, !Country %in% c("Greece", "Sweden","Russia","United States"))

##get rird of 2002 and 2003
#EuropeOmnibus <- subset(EuropeOmnibus, !Year %in% c("2002", "2003"))


write.table(EuropeOmnibus
            ,file=paste("data\\EuropeSample.csv"
                        ,sep=""
                        )
            #   ,header=TRUE
            , sep=","
            , row.names=FALSE
            , append=FALSE
            )
##change government spreads from factors

EuropeOmnibus$Cab_left_right <- as.character(EuropeOmnibus$Cab_left_right)
EuropeOmnibus$Cab_left_right <- as.numeric(EuropeOmnibus$Cab_left_right)

EuropeOmnibus$Cab_liberty_authority <- as.character(EuropeOmnibus$Cab_liberty_authority)
EuropeOmnibus$Cab_liberty_authority <- as.numeric(EuropeOmnibus$Cab_liberty_authority)

EuropeOmnibus$Cab_eu_anti_pro <- as.character(EuropeOmnibus$Cab_eu_anti_pro)
EuropeOmnibus$Cab_eu_anti_pro <- as.numeric(EuropeOmnibus$Cab_eu_anti_pro)

EuropeOmnibus$left_right_ls_spread <- as.character(EuropeOmnibus$left_right_ls_spread)
EuropeOmnibus$left_right_ls_spread <- as.numeric(EuropeOmnibus$left_right_ls_spread)

EuropeOmnibus$liberty_authority_ls_spread <- as.character(EuropeOmnibus$liberty_authority_ls_spread)
EuropeOmnibus$liberty_authority_ls_spread <- as.numeric(EuropeOmnibus$liberty_authority_ls_spread)

EuropeOmnibus$eu_anti_pro_ls_spread <- as.character(EuropeOmnibus$eu_anti_pro_ls_spread)
EuropeOmnibus$eu_anti_pro_ls_spread <- as.numeric(EuropeOmnibus$eu_anti_pro_ls_spread)

```



###How desirable is it that the EU exert strong leadership in world affairs? Investment SPending

**load and subset the data we want**
```{r EUleadEquipNarrowing, echo = TRUE, results = "hide"}

##load data using function
EULeadershipSpread <- subset(EuropeOmnibus, !is.na(EUldrSpread))


##get rid of Serbia
EULeadershipSpread <- subset(EULeadershipSpread, !Country %in% c("Serbia"))



```

**summary statistics of data**
```{r EUleadEquipSpreadSummary, echo = TRUE, results = "hide"}

##Summary statistics of data
summary(EULeadershipSpread)

```



**OLS model adding each variable individually**
```{r EUleadEquipspreadRegressions, echo = TRUE}
InvEULeadResults1 <- lm(log(EquSpnd_lead) ~ EUldrSpread, EULeadershipSpread) 
summary(InvEULeadResults1)
screenreg(list(InvEULeadResults1))

InvEULeadResults2 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt, EULeadershipSpread)
summary(InvEULeadResults2)
screenreg(list(InvEULeadResults1, InvEULeadResults2))

InvEULeadResults3 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + DomAt, EULeadershipSpread)
summary(InvEULeadResults3)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults3))

##DomAt is not significant nor does it change r^2 will not be included.

InvEULeadResults4 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + CivilWar, EULeadershipSpread)
summary(InvEULeadResults4)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults4))

##CivilWar doesn't increase r^2 and is not significant, will not be included

InvEULeadResults5 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + IntlCnf, EULeadershipSpread)
summary(InvEULeadResults5)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults5))

##IntlCnf is not significant and doesn't increase r^2, thus will not be included

InvEULeadResults6 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population), EULeadershipSpread)
summary(InvEULeadResults6)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6))

##log(Population) is significant and increases r^2, will be included

InvEULeadResults7 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap), EULeadershipSpread)
summary(InvEULeadResults7)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6, InvEULeadResults7))

## increases r^2, is significant and increases significance on other variables, will be included


InvEULeadResults8 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_left_right, EULeadershipSpread)
summary(InvEULeadResults8)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6, InvEULeadResults7, InvEULeadResults8))


##Cab_left_right is not significant and doesn't change the other coefficients and doen'st increase adjusted r^2

InvEULeadResults9 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_liberty_authority, EULeadershipSpread)
summary(InvEULeadResults9)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6, InvEULeadResults7, InvEULeadResults9))

##Cab_liberty_authority isn't significant, doesn't increase adjusted r^2 and doesn't affect the other variables

InvEULeadResults10 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + Cab_eu_anti_pro, EULeadershipSpread)
summary(InvEULeadResults10)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6, InvEULeadResults7, InvEULeadResults10))

##Cab_eu_anti_pro doesn't increase r^2, isn't significant, wont be included in the model

InvEULeadResults11 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + left_right_ls_spread, EULeadershipSpread)
summary(InvEULeadResults11)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6, InvEULeadResults7, InvEULeadResults11))

##left_right_ls_spread doens't affect model and isn't significant, will not be included

InvEULeadResults12 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + liberty_authority_ls_spread, EULeadershipSpread)
summary(InvEULeadResults12)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6, InvEULeadResults7, InvEULeadResults12))

##liberty_authority_ls_spread isn't significant and doesn't affect other parts of the model, will not be included

InvEULeadResults13 <- lm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap) + eu_anti_pro_ls_spread, EULeadershipSpread)
summary(InvEULeadResults13)
screenreg(list(InvEULeadResults1, InvEULeadResults2, InvEULeadResults6, InvEULeadResults7, InvEULeadResults13))

##eu_anti_pro_ls_spread doesn't increase r^2 and isn't significnat

```


**Tests of within and pooling** we should explain this in the appendix 

``` {r EULeadershipEquipspreadpooledTest, echo = TRUE}

InvEULeadPooled <- plm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap), data = EULeadershipSpread, model = "pooling")

InvEULeadWithin <- plm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap), data =  EULeadershipSpread, model = "within")
summary(InvEULeadWithin)

pFtest(InvEULeadWithin, InvEULeadPooled)
##p <2.2e^-16 thus reject the null that the pooling model works (all coefficients for each country should be the same ever year)

```


**Fixed Effects model**
```{r EUleadEquipspreadFixed, echo = TRUE}

InvEULeadResultsA <- plm(log(EquSpnd_lead) ~ EUldrSpread + IntAt + log(Population) + log(GDPpCap), EULeadershipSpread, index=c("Country", "Year"), model="within") 
summary(InvEULeadResultsA)

screenreg(list(InvEULeadResults7, InvEULeadResultsA))


```


**Compare Fixed effects to random effects** Hausman test. null hypothesist that RE model is true, alternative = fixed effects model
``` {r EquDefSperadHausmanTest, echo = TRUE}

InvEULeadWithin <- plm(log(EquSpnd_lead) ~ EUldrSpread +  IntAt + log(Population) + log(GDPpCap), data = EULeadershipSpread, index=c("Country", "Year"), model="within") 

InvEULeadRandom <- plm(log(EquSpnd_lead) ~ EUldrSpread +  IntAt + log(Population) + log(GDPpCap), data = EULeadershipSpread, index=c("Country", "Year"), model="random")

screenreg(list(InvEULeadWithin, InvEULeadRandom))

phtest(InvEULeadWithin, InvEULeadRandom)
##p - 9.25ee^-05, reject null and use fixed effects

```
